### 2023-10-07 Introduction.qmd

---
title: Introduction
---

# Variability and Generalization

## Introduction

The factors that influence how learning generalizes are of great interest to those seeking to understanding the nature of the human learning system, and to those attempting to improve the efficacy of teaching and training programs. Such factors are likely legion, but the present effort will limit itself to two characteristics of the learning input - variability and frequency skew. Variability manipulations control either the number of unique instances learners are trained on, or how dispersed or spread out the instances are. Much less research has examined skew manipulations, which hold the number of instances constant, and instead vary the relative frequencies with which the training items occur. Both factors have been linked to generalization in past research. The present essay with will provide an extensive review and discussion of both topics. The review of the expansive literature on the effect of attempts to include at least some examples from each of the many relevant domains, but much of the work is biased towards the background of the author. Conversely, the literature on skew proved to be far sparser than anticipated, as well as being overrepresented in a region of the language learning literature that is somewhat distant from more conventional cognitive psychology.

## The study of variability

In their most generic form, studies assessing the "benefits of variability" hypothesis assign participants to either a constant or varied group for the training stage of the experiment. Then, subjects in both groups complete an identical testing stage which often consists items/conditions seen during training, and novel items/conditions. If the varied group performs better in the testing stage, this is taken for evidence of the benefits of variability hypothesis. Even in this relatively simple between-groups design, researchers must make a number of crucial decisions which will now be highlighted.

1)  What is varied. Unless the task is unidimensional, there will often be many variables that could be varied or held fixed, and the experiment will need to decide the dimension(s) along which the variation will occur. For instance, in a projectile throwing accuracy task -- researchers might vary the distance from the target, the size of the target, the weight of the projectile. They might also vary a contextual variable not directly relevant to the task, but which will still be encoded by the subject on a trial by trial basis, e.g. the background color.

2)  The amount of variation, relative to the control condition. The simplest comparison would be to compare a constant group who trains with 1 example/condition, against a varied group that trains from 2 examples/conditions. However, it is not uncommon in the literature for the varied condition to train from 3 or 4 conditions. For example, Catalano and Kleiner train varied subjects from 4 different velocities in their coincident timing task, and Goode et al. (2008) have varied subjects' practice with 3 different variants (i.e. different letter scrambles of the same word) of an anagram for a given word, while their constant participants view the same variant 3 different times. Alternatively, rather than a constant vs. varied comparison, subjects in all conditions might experience a variety of training items, but with one group experiencing a greater number of unique items (Nosofsky et al. 2018).

3)  Locations of the varied examples within the task-space. For tasks in which the stimuli or conditions exist in some continuous metric space, the experimenter must decide whether the varied instances are relatively close together (e.g. throwing a ball from a distance of 4 feet and 5 feet), or far apart (throwing from 4 feet and 20 feet). Spreading the varied training items further apart may be beneficial in terms of providing a more representative sample of the task space to the learner, however large distances may also result in significant differences in difficulty between the training examples, which can be a common confound in variability studies.

4)  The proximity of the testing conditions to the training conditions of the varied and constant groups. Intuitively, the fairest form of comparison is to include testing conditions that are of an equivalent distance from both the varied and constant groups. However researchers might also attempt to demonstrate the benefits of variation as being sufficiently powerful to outperform constant training, even in cases where the constant group trained from a closer proximity to the testing conditions, or whose training conditions are identical to the testing conditions (Goode et al., 2008; Kerr & Booth, 1978).

Variability Literature Review

An early and influential work on the influence of variability on category learning is that of Posner and Keele (1968). In an ambitious attempt to address the question of how category information is represented, the authors trained participants to categorize artificial dot patterns, manipulating whether learners were exposed to examples clustered close to the category prototypes (e.g. a low variability condition), or spread further away from the prototype (the varied-training group). It should be noted that both groups in this study were trained with the same number of unique instances and the manipulated difference was how spread out the instances were. The authors claim based on prior experiments using the same stimuli, that the training stimuli for the varied group were at least as far away from the testing stimuli as the training stimuli of the less-varied group. The authors interpreted their findings as evidence for the extraction of an abstraction or schema that is extracted and stored, and then over time becomes more likely to be the reference point from which generalization occurs, given that specific instances are thought to decay at a faster rate than prototypes or schema. The Posner and Keele study have been extremely influential and continues to be cited in contemporary research as clear evidence that schema abstraction underlies the benefits of varied training. It's also referenced as a key influence in the development of Schema Theory of Motor Learning (Richard A. Schmidt, 1975), which in turn influenced decades of investigations on the potential benefits of varied training in motor skill learning. However, the classic Posner & Keele study despite being far more carefully designed than many subsequent studies, and despite being a relative rarity in explicitly discussing and attempting to control for potential confounds of similarity between groups, may nevertheless be emblematic of a common issue in many investigations of the effects of varied training on learning. The problem with Posner & Keele's conclusion was demonstrated clearly almost 3 decades later (Palmeri & Nosofsky, 2001), when researchers conducting a near replication of the original study also collected similarity judgements following training and performed multidimensional scaling analysis. Rather than being in the middle of the training stimuli as was the case in the physical stimuli space, the psychological representation of the prototype was shown reside at an extreme point, and generalization patterns by participants that would have seemed to warrant the learning of a prototype were then easily accounted for with only the assumption that the participants encoded instances. One of the primary concerns of the present paper is that many of the studies which purport to explain the benefits of variation via prototypes, schemas, or other abstractions, are often overlooking the potential of instance based similarity accounts.

## Motor skill learning

Training variation has also been shown to promote transfer in motor learning. Much of this research has been influenced by the work of Schmidt (1975), who proposed a schema-based account of motor learning as an attempt to address the longstanding problem of how novel movements are produced. Schema theory presumes a priori that learners possess general motor programs for classes of movements, such as an underhand throw. When called up for use, such programs must be parameterized, as well as schema rules that determine how a motor program is parameterized or scaled for a particular movement. Schema theory predicts that varied training results in the formation of a more general schema-rule, which can allow for transfer to novel movements within a given movement class, such as an underhand throw (though it is agnostic to the development of the movement classes themselves). Experiments that test this hypothesis are often designed to compare the transfer performance of a constant-trained group against that of a varied-trained group. Both groups train on the same task, but the varied group practices with multiple instances along some task-relevant dimension that remains invariant for the constant group. For example, investigators might train two groups of participants to throw a projectile at a target, with a constant group that throws from a single location, and a varied group that throws from multiple locations. Both groups are then tested from novel locations.

One of the earliest, and still often cited investigations of Schmidt's benefits of variability hypothesis was the work of Kerr and Booth (1978). Two groups of children, aged 8 and 12, were assigned to either constant or varied training of a bean bag throwing task. The constant group practiced throwing a bean-bag at a small target placed 3 feet in front of them, and the varied group practiced throwing from a distance of both 2 feet and 4 feet. Participants were blindfolded and unable to see the target while making each throw but would receive feedback by looking at where the beanbag had landed in between each training trial. 12 weeks later, all of the children were given a final test from a distance of 3 feet which was novel for the varied participants and repeated for the constant participants. Participants were also blind folded for testing and did not receive trial by trial feedback in this stage. However, at the halfway point of the testing stage they were allowed to see the landing location of the 4 beanbags they had thrown, and then completed the final 4 testing throws. In both age groups, participants performed significantly better in the varied condition than the constant condition, though the effect was larger for the younger, 8-year-old children. Although this design does not directly assess the hypothesis of varied training producing superior generalization to constant training (since the constant group is not tested from a novel position), it nevertheless offers a compelling example of the merits of varied practice.

On occasion the Kerr and Booth design may be nested within a larger experimental design. One such study that used a movement timing task, wherein subjects had to move their hand from a starting location, to a target location, attempting to arrive at the target location at specific time following the onset of a cue (Wrisberg et al., 1987). This study utilized 4 different constant groups, and 3 varied groups, with one of the constant groups training under conditions identical to the testing conditions, and which were not trained on by any of the varied groups, e.g. the design of Kerr and Booth. However, in this case the varied group did not outperform the constant group. A more recent study attempting a slightly more direct replication of the original Kerr & Booth study (Willey & Liu, 2018), having subjects throw beanbags at a target, with the varied group training from positions (5 and 9 feet) on either side of the constant group (7 feet). However this study diverged from the original in that the participants were adults; they faced away from the target and threw the beanbag backwards over their bodies; they alternated using their right and left hands every 6 trials; and underwent a relatively extreme amount of training (20 sessions with 60 practice trials each, spread out over 5-7 weeks). Like Wrisberg et al. (1987), this study did not find a varied advantage from the constant training position, though the varied group did perform better at distances novel to both groups.

Some support for the Kerr and Booth findings was found with a relatively less common experimental task of training participants in hitting a projectile at a target with the use of a racket (Green et al., 1995). Varied participants trained with tennis, squash, badminton, and short-tennis rackets were compared against constant subjects trained with only a tennis racket. One of the testing conditions had subjects repeat the use of the tennis racket, which had been used on all 128 training trials for the constant group, and only 32 training trials for the varied group. Nevertheless, the varied group outperformed the constant group when using the tennis racket at testing, and also performed better in conditions with several novel racket lengths. Of course, this finding is less surprising than that of Kerr & Booth, given that varied subjects did have some prior exposure to the constant groups condition. This highlights an issue rarely discussed in the literature, of how much practice from an additional position might be necessary to induce benefits. Experimenters almost uniformly have varied participants train with an equivalent number of trials from each of their conditions

One of the few studies that has replicated the surprising result of varied outperforming constant, from the constant training condition, did so in the relatively distant domain of verbal manipulation (Goode et al., 2008). All participants trained to solve anagrams of 40 different words ranging in length from 5 to 11 letters, with an anagram of each word repeated 3 times throughout training, for a total of 120 training trials. Although subjects in all conditions were exposed to the same 40 unique words (i.e. the solution to an anagram), participants in the varied group saw 3 different arrangements for each solution-word, such as DOLOF, FOLOD, and OOFLD for the solution word FLOOD, whereas constant subjects would train on three repetitions of LDOOF (spread evenly across training). Two different constant groups were used. Both constant groups trained with three repetitions of the same word scramble, but for constant group A, the testing phase consisted of the identical letter arrangement to that seen during training (e.g. LDOOF), whereas for constant group B, the testing phase consisted of a arrangement they had not seen during training, thus presenting them with a testing situation similar situation to the varied group. At the testing stage, the varied group outperformed both constant groups, a particularly impressive result, given that constant group A had 3 prior exposures to the word arrangement (i.e. the particular permutation of letters) which the varied group had not explicitly seen. However varied subjects in this study did not exhibit the typical decrement in the training phase typical of other varied manipulations in the literature, and actually achieved higher levels of anagram solving accuracy by the end of training than either of the constant groups -- solving 2 more anagrams on average than the constant group. This might suggest that for tasks of this nature where the learner can simply get stuck with a particular word scramble, repeated exposure to the identical scramble might be less helpful towards finding the solution than being given a different arrangement of the same letters. This contention is supported by the fact that constant group A, who was tested on the identical arrangement as they experienced during training, performed no better at testing than did constant group B, who had trained on a different arrangement of the same word solution -- further suggesting that there may not have been a strong identity advantage in this task.

Pitting varied against constant practice against each other on the home turf of the constant group provides a compelling argument for the benefits of varied training, as well as an interesting challenge for theoretical accounts that posit generalization to occur as some function of distance. However, despite its appeal this particular contrast is relatively uncommon in the literature. It is unclear whether this may be cause for concern over publication bias, or just researchers feeling the design is too risky. A far more common design is to have separate constant groups that each train exclusively from each of the conditions that the varied group encounters (Catalano & Kleiner, 1984; Chua et al., 2019; McCracken & Stelmach, 1977; Moxley, 1979; Newell & Shapiro, 1976), or for a single constant group to train from just one of the conditions experienced by the varied participants (Pigott & Shapiro, 1984; Roller et al., 2001; Wrisberg & McLean, 1984; Wrisberg & Mead, 1981). A less common contrast places the constant group training in a region of the task space outside of the range of examples experienced by the varied group, but distinct from the transfer condition (Wrisberg et al., 1987; Wulf, 1991).

Of particular relevant to the current essay is the early work of Catalano and Kleiner (1984), as theirs was one of the earliest studies to investigate the influence of varied vs. constant training on multiple testing locations of graded distance from the training condition. Participants were trained on coincident timing task, in which subjects observe a series of lightbulbs turning on sequentially at a consistent rate and attempt to time a button response with the onset of the final bulb. The constant groups trained with a single velocity of either 5,7,9, or 11 mph, while the varied group trained from all 4 of these velocities. Participants were then assigned to one of four possible generalization conditions, all of which fell outside of the range of the varied training conditions -- 1, 3, 13 or 15 mph. As is often the case, the varied group performed worse during the training phase. In the testing phase, the general pattern was for all participants to perform worse as the testing conditions became further away from the training conditions, but since the drop off in performance as a function of distance was far less steep for the varied group, the authors suggested that varied training induced a decremented generalization gradient, such that the varied participants were less affected by the change between training and testing conditions.

In the category learning literature, the constant vs. varied comparison is much less suitable. Instead, researchers tend to compare a condition with many repetitions of a few items against condition with fewer repetitions of a wider array of exemplars. Much of the earlier work in this sub-area trained subjects on artificial categories, such as dot patterns (Homa & Vosburgh, 1976; Posner & Keele, 1968), where more varied or distorted training examples were often shown to produce superior generalize when categorizing novel exemplars. More recently, researchers have also begun to utilize more realistic stimuli in their experiments. Wahlheim et al. (2012) conducted one such study. In a within-participants design, participants were trained on bird categories with either high repetitions of a few exemplars, or few repetitions of many exemplars. Across four different experiments, which were conducted to address on unrelated question on metacognitive judgements, the researchers consistently found that participants generalized better to novel species following training with more unique exemplars (i.e. higher variability), while high repetition training produced significantly better performance categorizing the specific species they had trained on. A variability advantage was also found in the relatively complex domain of rock categorization (Nosofsky et al., 2018). For 10 different rock categories, participants were trained with either many repetitions of 3 unique examples of each category, or few repetitions of 9 unique examples, with an equal number of total training trials in each group (the design also included 2 other conditions less amenable to considering the impact of variation). The high-variability group, trained with 9 unique examples, showed significantly better generalization performance than the other conditions. Moreover, the pattern of results in this study could be nicely accounted for by an extended version of the Generalized Context Model.

The studies described thus far have studied the benefits of variability by exposing participants to a greater or lesser number of distinct examples during training. A distinct sub-literature within the category learning domain has focused much less on benefits derived from varied training, instead emphasizing how increased variability during the learning of a novel category influences how far the category boundary will then be generalized. The general approach is to train participants on examples from two categories, with the examples from one of the categories being more dispersed than the other. Participants are then tested with novel items located within ambiguous regions of the task space which allow the experimenters to assess whether the difference in variability influences how far participants generalize the category boundaries.

Cohen et al. (2001) trained subjects on two categories, one with much more variability than the other. In experiment 1, a low variability category composed of 1 instance was compared against a high-variability category of 2 instances in one condition, and 7 instances in another. In experiment 2 both categories were composed of 3 instances, but for the low-variability group the instances were clustered close to each other, whereas the high-variability groups instances were spread much further apart. Participants were tested on an ambiguous novel instance that was located in between the two trained categories. Both experiments provided evidence that participants were much more likely to categorize the novel middle stimulus into a category with greater variation. Moreover, this effect was at odds of the predications of the baseline version of the GCM, thus providing some evidence that training variation may at least sometimes induce effects that cannot be entirely accounted for by exemplar-similarity accounts.

Further observations consonant with the results of Cohen et al. (2001) have since been observed in numerous investigations (Hahn et al., 2005; Hsu & Grifﬁths, 2010; Perlman et al., 2012; Sakamoto et al., 2008). The results of Sakamoto et al. (2008) are noteworthy. They first reproduced the basic finding of participants being more likely to categorize an unknown middle stimulus into a training category with higher variability. In a second experiment, they held the variability between the two training categories constant and instead manipulated the training sequence, such that the examples of one category appeared in an ordered fashion, with very small changes from one example to the other (the stimuli were lines that varied only in length), whereas examples in the alternate category were shown in a random order and thus included larger jumps in the stimulus space from trial to trial. They found that the middle stimulus was more likely to be categorized into the category that had been learned with a random sequence, which was attributed to an increased perception of variability which resulted from the larger trial to trial discrepancies.

The study by Hahn et al. (2005), is also of particular interest to the present discussion. Their experimental design was similar to previous studies, but they included a larger set of testing items which were used to assess generalization both between the two training categories as well as novel items located in the outer edges of the training categories. During generalization testing, participants were given the option to respond with "neither", in addition to responses to the two training categories. The "neither" response was included to test how far away in the stimulus space participants would continue to categorize novel items as belonging to a trained category. Consistent with prior findings, high-variability training resulted in an increased probability of categorizing items in between the training categories as belong to the high variability category. Additionally, participants trained with higher variability also extended the category boundary further out into the periphery than participants trained with a lower variability category were willing to do. The authors then used the standard GCM framework to compare a variety of similarity-based models to account for their results. Of particular interest are their evaluations of a category response bias parameter, and a similarity scaling parameter. A model fit improvement when the response bias parameter is allowed to vary between the high-variability and low-variability trained groups is taken to suggest a simple bias for responding with one of the trained categories over the other. Alternatively, an improvement in fit due to a separate similarity scaling parameter may reflect the groups being differentially sensitive to the distances between stimuli. No improvement in model fit was found by allowing the response-bias parameter to differ between groups, however the model performance did improvement significantly when the similarity scaling parameter was fit separately. The best fitting similarity-scaling parameters were such that the high-variability group was less sensitive to the distances between stimuli, resulting in greater similarity values between their training items and testing items -- accounting for their extended generalization gradients, but also for their poor performance in a recognition condition. Additional model comparisons suggested that this similarity rescaling applied across the entire stimulus space, rather than to the high variability category in particular.

Variability effects have also been examined in the higher-level domain of how learners acquire novel concepts, and then instantiate (rather than merely recognize) that concept in untrained contexts (Braithwaite & Goldstone, 2015). This study trained participants on problems involving the concept of sampling with replacement (SWR). Training consisted of examples that were either highly similar in their semantic context (e.g. all involving people selecting objects) or in which the surface features were varied between examples (e.g. people choosing objects AND objects selected in a sequence). The experimenters also surveyed how much prior knowledge each participant had with SWR. They found that whether variation was beneficial depended on the prior knowledge of the participants -- such that participants with some prior knowledge benefited from varied training, whereas participants with minimal prior knowledge performed better after training with similar examples. The authors hypothesized that in order to benefit from varied examples, participants must be able to detect the structure common to the diverse examples, and that participants with prior knowledge are more likely to be sensitive to such structure, and thus to benefit from varied training. To test this hypothesis more directly, the authors conducted a 2nd experiment, wherein they controlled prior knowledge by exposing some subjects to a short graphical or verbal pre-training lesson, designed to increase sensitivity to the training examples. Consistent with their hypothesis, participants exposed to the structural sensitivity pre-training benefited more from varied training than the controls participants who benefited more from training with similar examples.

Variability has also been examined within the realm of language learning. A particularly impressive study is that of Perry et al. (2010). In nine training sessions spread out over nine weeks infants were trained on object labels in a naturalistic play setting. All infants were introduced to three novel objects of the same category, with participants in the tight condition being exposed to three similar exemplars of the category, and participants in the varied condition being exposed to three dissimilar objects of the same category. Importantly, the similarity of the objects was carefully controlled for by having a separate group of adult subjects provide pairwise similarity judgements of the category objects prior to the study onset. Multidimensional scaling was then performed to obtain the coordinates of the objects psychological space, and out of the 10 objects for each category, the 3 most similar objects were selected for the tight group and the three least similar objects for the varied group, with the leftover four objects being retained for testing. By the end of the nine weeks, all of the infants had learned the labels of the training objects. The varied group demonstrated superior ability to correctly generalize the object labels to untrained exemplars of the same category, a pattern consistent with much of the existing literature. More interesting was the superior performance of the varied group on a higher order generalization task -- such that they were able to appropriately generalize the bias they had learned during training for attending to the shape of objects to novel solid objects, but not to non-solids. The tight training group, on the other hand, tended to overgeneralize the shape bias, leading the researchers to suggest that the varied training induced a more context-sensitive understanding of when to apply their knowledge.

## Some complications to the influence of variability

### Sequence Effects

A necessary consequence of varied training is that participants will have the experience of switching from one task condition to another. The number of switches can vary greatly, with the two extremes being varied participants completing all of their training trials in one before switching to the next condition (blocked sequencing), or if alternate between conditions on a trial by trial basis (random/intermixed/interleaved sequencing). Not long after the initial influx of schema-theory inspired studies testing the benefits of variability hypothesis it was shown that the influence of varied training might interact with the type of training sequence chosen by the experimenter (J. B. Shea & Morgan, 1979). In this seminal study, both groups of training subjects trained with the same number of trials of three separate movement patterns. A blocked group that completed all of their trials with one sequence before beginning the next sequence, and a random group that trained with all three movement patterns interspersed throughout the course of training. Participants were also randomly assigned to retention testing under either blocked or random sequence conditions, thus resulting in all four training-testing combinations of blocked-blocked; blocked-random; random-blocked; random-random. There was some effect of sequence context, such that both groups performed better when the testing sequence matched their training sequence. However, the main finding of interest was the advantage random-training, which resulted in superior testing performance than blocked training regardless of whether the testing stage had a blocked or random sequence, an effect observed both immediately after training, and in a follow up test ten days after the end of training.

Prior to the influential Shea and Morgan (1979) study, studies investigating the benefits of variability hypothesis had utilized both blocked and random training schedules, often without comment or justification. It was later observed (Lee et al., 1985) that positive evidence for benefits of varied training seemed more likely to occur for studies that utilized random schedules. The theoretical basis of such studies was invariably an appeal to Schmidt's schema theory; however schema theory made no clear predictions of an effect of study sequence on retention or generalization, thus prompting the need for alternate accounts. One such account, the elaborative processing account (J. B. Shea & Zimny, 1983), draws on the earlier work of Battig (Battig, 1966) and argues that randomly sequencing conditions during training promotes comparison and contrastive processes between those conditions, which result in a deeper understanding of the training task than could arise via blocked sequencing. Supporting evidence for elaborative processing comes in the form of random-sequence trained subjects self-reporting more nuanced mental representations of movement patterns following training (J. B. Shea & Zimny, 1983), and by manipulating whether subjects are able to perform comparisons during training (Wright et al., 1992). An alternative, though not incompatible account suggests that the benefits of random-sequencing are a result of such sequences forcing the learner to continually reconstruct the relevant motor task in working memory (Lee & Magill, 1985). Blocked training, on the other hand, allows the learner to maintain the same motor task in short term memory without decay for much of the training which facilitates training performance, but hinders ability to retrieve the appropriate motor memory in a later testing context. A much more recent study (Chua et al., 2019), replicates the standard findings of an advantage of varied training over constant training (expt 1, bean-bag throwing task), and of random training over blocked training (expt 2 & 3, bean-bag throwing & golf putting). The novelty of this study is that the experimenters queried subjects about their attentional focus throughout the training stage. In all three experiments varied or random trained-subjects reported significantly greater external attention (e.g. attending to the target distance), and constant or blocked subjects reported more internal attention (e.g. posture or hand position). The authors argue that the benefits of varied/random training may be mediated by changes in attentional focus, however the claims made in the paper seem to go far beyond what can be justified by the analyses reported -- e.g. the increased external focus could be a simple byproduct of varied training. A stronger form of evidence that was not provided may have been to use multiple regression analyses to show that the testing advantage of the varied/random groups over the constant/blocked groups could be accounted for by the differences in self-reported attentional focus.

Other task and participant effects

Of course, the effects of varied training, and different training sequences, are likely to be far more complex than simply more varied training being better than less, or random training being better than blocked. Null effects of both manipulations have been reported (Magill & Hall, 1990; Van Rossum, 1990 for reviews), and a variety of moderators have emerged. In one of the earlier examples of the complex relationship between study sequence and learning (Del Rey et al., 1982), experimenters recruited participants who self-reported either large amounts, or very little experience with athletic actives, and then trained participants on a coincident timing task under with either a single constant training velocity, or with four training velocities with either blocked, or random training sequence conditions - resulting in 6 experimental conditions: (athlete vs. non-athlete) x (constant vs. blocked vs. random training). Athlete participants had superior performance during training, regardless of sequence condition, and training performance was superior for all subjects in the constant group, followed by blocked training in the middle, and then random training resulting in the worst training performance. Of greater interest is the pattern of testing results for novel transfer conditions. Among the athlete-participants, transfer performance was best for those who received random training, followed by blocked, and then constant training. Non-athletes showed an almost opposite pattern, with superior performance for those who had constant training. A similar pattern was later observed in a golf-putting training study, wherein participants who had some prior golf experience benefited most from random-sequenced training, and participants with no golf experience benefited most from blocked training (Guadagnoli et al., 1999). More recently, the same pattern was observed in the concept learning literature (Braithwaite & Goldstone, 2015, expt 1). This study trained participants on a mathematical concept and found that participants who self-reported some prior experience with the concept improved more from pre-test to post-test after training with varied examples, while participants who reported no prior experience showed greater gains following training with highly similar examples.

In addition to the influence of prior experiences described above, ample evidence also suggests that numerous aspects of the experiment may also interact with the influence of variation. One important study examined the impact of the amount of training completed in a force production task (C. H. Shea et al., 1990). This study employed a typical blocked vs. random training procedure, but with the additional manipulation of separate groups receiving 50, 200, or 400 total training trials. For the group that received only 50 training trials retention was best when training had been blocked. However, for the conditions that received 200 or 400 training trials the pattern was reversed, with random training resulting in superior retention than blocked training. These results were taken to suggest that the benefits of randomization may take time to emerge. Another experimental factor shown to interact with training sequence is the complexity of the training task (Albaret & Thon, 1998). In addition to random or blocked training, participants in this study were assigned to train on a drawing task at one of three different levels of complexity (reproducing from memory shapes composed of 2, 3 or 4 components). On a transfer task 48 hours after the completion of training, only participants trained at the lower levels of task complexity (2 or 3 components) showed superior performance to the blocked condition. The authors suggest that the benefits of random sequencing, thought to arise from more elaborate cognitive processing, or the necessity of continually recalling task information from long term into short term memory, are more likely to be obscured as the complexity of the task forces the blocked participants to also engage in such processes.

A final important influence of particular relevance to the practice-sequence literature concerns the exact structure of "random" sequencing. Although the term random is commonly used for convenience, experimenters do not typically leave the order of training entirely up to chance. Rather, the training sequence is often constrained such that each condition must occur a minimum number of times in each quartile of the training phase, thus resulting in an even distribution the conditions throughout training. While the assurance of the conditions being evenly spread throughout training is consistent across studies, other aspects of the sequence structure are a bit more idiosyncratic. Some researchers report setting a maximum number of consecutive repetitions, e.g. no more than 2 consecutive trials of the same condition (Del Rey et al., 1982; J. B. Shea & Morgan, 1979), or structure the random trials such that the same condition never occurs consecutively (Wulf, 1991). Also common is to structure experiments such that random condition really consists of many small blocks, where participants do a few trials of one condition consecutively and then switch to another condition (Chua et al., 2019; Willey & Liu, 2018; Wrisberg et al., 1987), resulting in much more switches than would arise if training was perfectly blocked. The question of whether such differences in the structure of random sequencing are consequential has been addressed experimentally a few times, in all cases consisting of a 1) a no-repeat random condition; 2) a blocked random condition (typically 3 or 4 repeats before a switch); and 3) a standard fully-blocked condition. Blocked-random training resulted in better performance than either repeat-random, or fully - blocked training in both a bean-bag throwing (Pigott & Shapiro, 1984), and basketball shot training study (Landin & Hebert, 1997), and in a replication plus extension of the seminal Shea and Morgan (1979) study, blocked-random training was equally effective as no-repeat random training, with both random structures leading to better performance than the fully-blocked training condition. Consequences on different study schedules have also been repeatedly observed in the category learning literature (Carvalho & Goldstone, 2014, 2017). This line of research has revealed that the effects of blocking vs. interleaving can depend on the structure of the category being learned, and also that the different schedules can result in the participants requiring different representations. A fruitful line of inquiry in the motor skill learning literature may be to attempt to identify whether structural aspects of the motor task interact with different training sequences in a reliable manner.

Numerous researchers have attempted to provide coherent frameworks to account for the full range of influences of training variation and sequencing described above (along with many other effects not discussed). Such accounts are generally quite similar, invoking ideas of desirable levels of difficulty(Bjork & Bjork, 2011; R.A. Schmidt & Bjork, 1992), or optimal challenge points (Guadagnoli & Lee, 2004). They tend to start by describing the dissociation between acquisition performance (performance during training) and testing performance (delayed retention and/or transfer), most strikingly observed as varied/random training participants performing worse than their constant/blocked counterparts during the training stage of the study, but then outperforming the constant/blocked comparisons at a later retention or transfer stage. This observation is then used to justify the idea that the most enduring and generalizable learning occurs by training at an optimal level of training difficulty, with difficulty being some function of the experience of the learner, and the cognitive or visuomotor processing demands of the task. It then follows that the factors that tend to make training more difficult (i.e. increased variability or randomization), are more likely to be beneficial when the learner has some experience, or when the processing demands of the task are not too extreme (which may only occur after some experience with the task). Such frameworks may be helpful heuristics in some cases, but they also seem to be overly flexible such that any null result of some intervention might be accounted for by a suboptimal amount of training trials, or by suggesting the training task was too difficult. The development of computational models that can account for how changes in the parameters of the motor-skill task scale with difficulty, would be a great step forward.