### 2023-10-17 Manuscript/pre_sections/pre_html.qmd
fix html toc

<div style="text-align: center; margin-top: 50px; font-size: 24px; font-weight: bold;">
  Variability and Generalization - A Similarity-Based Account
</div>

<div style="text-align: center; margin-top: 300px; font-size: 32px;">
  Thomas E. Gorman
</div>

<!-- Second Page and so on... -->
<div style="text-align: justify; margin-top: 700px;">
  Submitted to the faculty of the University Graduate School in partial fulfillment of the
  requirements for the degree Doctor of Philosophy in the Department of Psychology and Brain
  Sciences and the Cognitive Science Program, Indiana University
  Indiana University
</div>

<div style="margin-top: 400px;">
  <!-- Your page numbering if needed -->
</div>

<div style="page-break-before: always;"></div>

<!-- Committee Members Page -->
<div style="text-align: center; margin-top: 300px;">
  Accepted by the Graduate Faculty, Indiana University, in partial fulfillment of the
  requirements for the degree of Doctor of Philosophy.
</div>

<div style="text-align: center; margin-top: 300px;">
  _____________________________  Robert L. Goldstone<br><br>
  _____________________________  Robert Nosofsky<br><br>
  _____________________________  Peter Todd<br><br>
  _____________________________  Mike Jones
</div>

<div style="page-break-before: always;"></div>

<div style="page-break-before: always;"></div>

<!-- Acknowledgements -->
<div style="text-align: center; font-weight: bold; font-size: 24px;">
  Acknowledgements
</div>

<div style="page-break-before: always;"></div>

## Table of contents
<div id="html-toc"></div>
```{=html}
<script>
<!-- document.addEventListener("DOMContentLoaded", function(){ -->
<!--   var toc = document.getElementById("TOC"); -->
<!--   var placeholder = document.getElementById("toc-placeholder"); -->
<!--   if (toc && placeholder) { -->
<!--     placeholder.appendChild(toc); -->
<!--   } -->
<!-- }); -->

<!-- document.addEventListener("DOMContentLoaded", function(){ -->
<!--     var tocDiv = document.getElementById("html-toc"); -->
<!--     var headings = document.querySelectorAll('h1, h2, h3'); // adjust if you need more depth -->
<!--     var tocList = document.createElement("ul"); -->

<!--     headings.forEach(function(heading) { -->
<!--         var listItem = document.createElement("li"); -->
<!--         var link = document.createElement("a"); -->
<!--         link.textContent = heading.textContent; -->
<!--         link.href = "#" + heading.id; -->
<!--         listItem.appendChild(link); -->

<!--         if (heading.tagName == "H2") { -->
<!--             listItem.style.marginLeft = "20px"; -->
<!--         } else if (heading.tagName == "H3") { -->
<!--             listItem.style.marginLeft = "40px"; -->
<!--         } -->
<!--         tocList.appendChild(listItem); -->
<!--     }); -->

<!--     tocDiv.appendChild(tocList); -->
<!-- }); -->

document.addEventListener("DOMContentLoaded", function(){
    var tocDiv = document.getElementById("html-toc");
    var headings = document.querySelectorAll('h1, h2, h3'); // adjust if you need more depth
    var tocList = document.createElement("ul");

    headings.forEach(function(heading, index) {
        // Create or get the ID for the heading
        if (!heading.id) {
            heading.id = heading.textContent.replace(/[^a-z0-9]+/gi, '-').toLowerCase() + "-" + index;
        }
        var listItem = document.createElement("li");
        var link = document.createElement("a");
        link.textContent = heading.textContent;
        link.href = "#" + heading.id;
        listItem.appendChild(link);

        if (heading.tagName == "H2") {
            listItem.style.marginLeft = "20px";
        } else if (heading.tagName == "H3") {
            listItem.style.marginLeft = "40px";
        }
        tocList.appendChild(listItem);
    });

    tocDiv.appendChild(tocList);
});

</script>
```