### 2023-10-19 paper.qmd

---
title: Dissertation Manuscript
toc: false
page-layout: full
---

```{=html}
<p class="text-center">
  <a class="btn btn-success btn-sm mb-1" href="/Manuscript/output/manuscript.html" target="manuscript">
  <iconify-icon inline="" icon="fa6-solid:file-lines"></iconify-icon> &ensp;HTML
  </a>

  <a class="btn btn-success btn-sm mb-1" href="/Manuscript/output/manuscript.html" target="_blank">
  <iconify-icon inline="" icon="fa6-solid:arrow-up-right-from-square"></iconify-icon> &ensp;HTML <small>(new window)</small>
  </a>

  <a class="btn btn-warning btn-sm mb-1" href="/Manuscript/output/TG_Full_Dissertation.pdf" target="manuscript">
  <iconify-icon inline="" icon="fa6-solid:file-pdf"></iconify-icon> &ensp;PDF
  </a>

<a class="btn btn-warning btn-sm mb-1" target="_blank" href="/Manuscript/output/TG_Full_Dissertation.pdf">
  <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;PDF <small>(download)</small>
  </a>

    <a class="btn btn-info btn-sm mb-1" target="_blank" href="/Manuscript/output/TG_Full_Dissertation.docx">
  <iconify-icon inline="" icon="fa6-solid:file-word"></iconify-icon> &ensp;DOCX <small>(download)</small>
  </a>

  <a class="btn btn-info btn-sm mb-1" target="_blank" href="/Manuscript/manuscript.html.md">
  <iconify-icon inline="" icon="fa6-solid:file-word"></iconify-icon> &ensp;md <small>(download)</small>
  </a>

  <!--  <a class="btn btn-warning btn-sm mb-1" href="/Manuscript/manuscript.pdf.md" target="manuscript"> -->
  <!-- <iconify-icon inline="" icon="fa6-solid:file-pdf"></iconify-icon> &ensp;Manuscript-md -->
  <!-- </a>  -->

</p>

<iframe src="/Manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:95%;width:100%;border:2px solid #ccc;" height="95%" width="100%" name="manuscript"></iframe>
```