### 2023-10-28 Appendix/IGAS-Supplemental.qmd

---
# title: "R Notebook"
# output: html_notebook
execute: 
  warning: false
  eval: true
---

```{r, include=FALSE, echo=FALSE}
source(here::here('Functions/IGAS_ProcessFunctions.R'))
library(patchwork)
#rm(list=ls())
###### data loading and organization #######

# load the processed data from experiment 1 and 2
e1 <- readRDS(here::here("data/igas_e1_cleanedData-final.rds")) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity)
e2<- readRDS(here::here('data/igas_e2_cleanedData-final.rds')) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity)
# load subject similarity data - computed with the IGAS model in 'IGAS-SimModel.R'

options(contrasts = c("contr.sum", "contr.poly"))
defaultContrasts = options()$contrasts
theme_set(theme_classic())

dodge <- position_dodge(width = 0.9)
e2GrpPos <- c("400","500","625","675","800","900")
e2Grp <- paste("Constant","Constant", "Constant","Constant","Constant","Constant", "Varied")
e2Labels <- paste(c("400\n Constant","500\n Constant","625\n Constant","675\n Constant",
                   "800\n Constant","900\n Constant","500-800\n Varied"),sep="")

e1Pos <- c("610","760","835","910")
e1Var <- paste("Varied Train Position","Constant Train Position", "Novel Position", "Varied Training Position")
e1Labels<- paste(c("610\n Varied Trained","760\n Constant Trained","835\n Novel Location","910\n Varied Trained"),sep="")

transfer <- filter(e2, stage=="Transfer") %>% droplevels() %>% select(-trainHalf,-initialVelocityY,ThrowPosition2)%>% ungroup()
transfer <- transfer %>% group_by(positionX) %>% mutate(globalAvg=mean(AbsDistFromCenter),globalSd=sd(AbsDistFromCenter)) %>% 
  group_by(sbjCode,positionX) %>% mutate(scaledDev = scaleVar(globalAvg,globalSd,AbsDistFromCenter)) %>%ungroup()
transfer <- transfer %>% group_by(sbjCode,positionX) %>% mutate(ind=1,testPosIndex=cumsum(ind),posN=max(testPosIndex)) %>%
  select(-ind) %>% mutate(testHalf = case_when(testPosIndex<15 ~"1st Half",testPosIndex>=15 ~"2nd Half")) %>% convert_as_factor(testHalf)
variedTest <- transfer %>% filter(condit==7) %>% mutate(extrapolate=ifelse(positionX=="900" | positionX=="400","extrapolation","interpolation")) 
constantTest <- transfer %>% filter(condit!=7) %>% mutate(extrapolate=ifelse(distFromTrain==0,"interpolation","extrapolation"))
transfer <- rbind(variedTest,constantTest)
transfer<- transfer %>% mutate(novel=ifelse(distFromTrain3==0,"trainedLocation","novelLocation"))%>% convert_as_factor(novel,extrapolate)
transfer <- transfer %>% relocate(sbjCode,condit2,Group,conditType2,stage,trial,novel,extrapolate,positionX,
                                  AbsDistFromCenter,globalAvg,globalSd,scaledDev,distFromTrain3) %>% ungroup()

transfer <- transfer %>% group_by(sbjCode,positionX) %>% mutate(halfInd=floor(max(testPosIndex)/2)) %>%
    mutate(testHalfSbj = case_when(testPosIndex<halfInd ~"1st-Half",testPosIndex>=halfInd ~"2nd-Half")) %>% convert_as_factor(testHalfSbj)

```

Reviewer #2: This study addresses a question that is important both theoretically and practically. However, the authors need to rule out the following, less interesting alternative. Namely, the results could be due to task practice effect, as follows.

Since there was no pre-training test, and no practice trials (as far as I can tell), and since the task was an online motor task that participants could not rely on their prior motor experience, trying to launch the ball to the target could only be done via trial and error. For the varied training group, they got to practice at two distances. Therefore, they had a better "calibration" in terms of the relationship between launching speed and target distance. This was likely beneficial both in Exp.1 when both transfer distances were interpolations from the two trained distances, and in Exp.2 when two transfer distances were interpolations and two were extrapolations but the latter two were immediately next to the training distances.

In comparison, since the constant group trained at only a single distance, any transfer distance (or at least the first transfer distance tested) was extrapolation even if this transfer distance was shorter than the trained, because the participants did not know beforehand how to shoot the ball to the shortest distance due to the existence of the barrier. If the transfer distance was longer, for sure that was extrapolation.

Regardless, the above analysis suggests that the constant group would always be a step behind the varied group. The number of trials at each transfer distance may not be sufficient for them to catch up the varied group either (whether there was learning during testing should be checked). If such disadvantage for the constant group is indeed due to the lack of tryout opportunities, then the authors should verify whether the same results still hold if all groups were provided opportunities to practice, or if pre-training tests across all distances were offered.

### exponential learning models fit to individual subjects
```{r}

e2TestFits <- readRDS(here::here('data/IGAS-e2TestFits-April4.rds'))
e2TestFits <- e2TestFits %>% mutate(Asymptote.Minus.Start=pAsym-pStart)
exp2.fit2 <- e2TestFits %>% ungroup() %>% group_by(sbjCode,conditType) %>%
  summarise(MeanAsym=mean(pAsym),MeanStart=mean(pStart),
            MeanRate=mean(pRate),
            asymMinusStart=mean(Asymptote.Minus.Start),.groups="keep") %>% 
  ungroup() %>% as.data.frame()

```

### Group comparison of learning rate fits
```{r}
# mr1=e2TestFits %>% anova_test(dv=pRate,between=conditType,wid=sbjCode,within=positionX,type=3);show(mr1)
# mr2=exp2.fit2 %>% anova_test(dv=MeanRate,between=conditType,wid=sbjCode,type=3);show(mr2)

h4<-e2TestFits %>% ggplot(aes(x=positionX,y=pRate,fill=conditType))+geom_bar(stat="summary",position=dodge,fun="mean")+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ggtitle("Learning rates ")+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+ylab("Inverse Learning Rate")+xlab("Testing Location")
h4

# mr1=e2TestFits %>% filter(converged==TRUE)%>% anova_test(dv=pRate,between=conditType,wid=sbjCode,within=positionX,type=3);show(mr1)
# mr2=exp2.fit2 %>% anova_test(dv=MeanRate,between=conditType,wid=sbjCode,type=3);show(mr2)

h4<-e2TestFits %>%filter(Pval<.4)%>% ggplot(aes(x=positionX,y=pRate,fill=conditType))+geom_bar(stat="summary",position=dodge,fun="mean")+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ggtitle("Learning rates ")+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+ylab("Inverse Learning Rate")+xlab("Testing Location")
h4

```

### First vs. second half of testing stage
```{r,fig.height=7,fig.width=8}

testSplit <- transfer %>%
  group_by(sbjCode, conditType, positionX, testHalfSbj) %>%
  summarise(
    MeanTargetDistance = mean(AbsDistFromCenter),
    MeanScaledDev =mean(scaledDev,trim=.05),
    .groups = "keep"
  ) %>% 
  as.data.frame()

testSplit2 <- transfer %>%
  group_by(sbjCode, conditType, testHalfSbj) %>%
  summarise(
    MeanTargetDistance = mean(AbsDistFromCenter, trim = .01),
    MeanScaledDev =
      mean(scaledDev, trim = .05),
    .groups = "keep"
  ) %>% as.data.frame()

tsw <- testSplit %>% ungroup() %>% 
  pivot_wider(names_from = testHalfSbj,values_from=c(MeanTargetDistance,MeanScaledDev)) %>%
  mutate(endMinusStart = `MeanTargetDistance_2nd-Half` - `MeanTargetDistance_1st-Half`,
    endMinusStartScaled = `MeanScaledDev_2nd-Half` - `MeanScaledDev_1st-Half`) %>% 
    as.data.frame()

tsw2 <- tsw %>% 
  group_by(sbjCode,conditType) %>% summarise(endMinusStart=mean(endMinusStart),endMinusStartScaled=mean(endMinusStartScaled)) %>% as.data.frame()

# testSplit %>% ggplot(aes(x=testHalfSbj,y=MeanTargetDistance))+
#   geom_bar(aes(group=conditType,fill=conditType),stat="summary",position=dodge)+
#   facet_wrap(~positionX,ncol=2)+
#   stat_summary(aes(x=testHalfSbj,group=conditType),fun.data=mean_se,geom="errorbar",position=dodge)

h1=testSplit %>% filter(testHalfSbj=="1st-Half") %>% ggplot(aes(x=positionX,y=MeanTargetDistance))+
 geom_bar(aes(group=conditType,fill=conditType),stat="summary",position=dodge,fun="mean")+stat_summary(aes(x=positionX,group=conditType),fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+
 scale_y_continuous(name="Mean Absolute Deviation From Target",limits=c(0,400))+ 
 ggtitle("Testing Performance - 1st half")+
  ylab("Mean Absolute Deviation From Target")+xlab("Testing Location")+theme(legend.position="none")

h2=testSplit %>% filter(testHalfSbj=="2nd-Half") %>% ggplot(aes(x=positionX,y=MeanTargetDistance))+
 geom_bar(aes(group=conditType,fill=conditType),stat="summary",position=dodge,fun="mean")+stat_summary(aes(x=positionX,group=conditType),fun.data=mean_se,geom="errorbar",position=dodge,width=.8) + 
 scale_y_continuous(name="",limits=c(0,400))+ 
 ggtitle("Testing Performance - 2nd half")+
  xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+
  guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)

#egg::ggarrange(h1,h2,ncol=2)

h3 <- tsw %>% ggplot(aes(x=positionX,y=endMinusStart))+
  geom_bar(aes(group=conditType,fill=conditType),stat="summary",position=dodge,fun="mean")+
  stat_summary(aes(x=positionX,group=conditType),fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+
  ylab("2nd Half Deviation - 1st Half Deviation")+ 
  ggtitle("Improvement Per Testing Location")+
  xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+
  guides(fill=guide_legend(title="Testing Location"))+theme(legend.position="none")

#ggarrange(h1,h2,h3,h4,ncol=2)

(h1 + h2) / (h3 + h4)

# 
 #testSplit2 %>% anova_test(dv=MeanScaledDev,between=conditType,within = c(testHalfSbj),wid=sbjCode,type=3)
 #testSplit %>% anova_test(dv=MeanTargetDistance,between=conditType,covariate=testHalfSbj,within=(positionX),wid=sbjCode)
# testSplit %>% lmer(MeanTargetDistance ~ conditType*testHalfSbj + (1|sbjCode)+(1+testHalfSbj|positionX),data=.) %>% summary()

# testSplit %>% anova_test(dv=MeanTargetDistance,between=conditType,within=positionX,wid=sbjCode,type=3)
# 
# print("Testing End - Testing Start Comparison")
# tsw %>% anova_test(dv=endMinusStart,between=conditType,within=positionX,wid=sbjCode,type=3)
# tsw2 %>% anova_test(dv=endMinusStartScaled,between=conditType,wid=sbjCode,type=3)
# 
# print("2nd Half of testing only comparison")
# testSplit %>% filter(testHalfSbj=="2nd-Half")%>%  anova_test(dv=MeanTargetDistance,between=conditType,within=(positionX),wid=sbjCode)
# testSplit2 %>% filter(testHalfSbj=="2nd-Half")%>%  anova_test(dv=MeanScaledDev,between=conditType,wid=sbjCode,type=3)
# 
# testSplit2 %>% filter(testHalfSbj=="2nd-Half")%>% group_by(conditType) %>% get_summary_stats(type = "mean_sd")

```

### Group Comparison for asymptote-starting performance
```{r}

# ma1=e2TestFits %>% filter()%>% anova_test(dv=Asymptote.Minus.Start,between=conditType,wid=sbjCode,within=positionX,type=3);show(ma1)
# ma2=exp2.fit2 %>% anova_test(dv=asymMinusStart,between=conditType,wid=sbjCode,type=3);show(ma2)

e2TestFits %>% ggplot(aes(x=conditType,y=Asymptote.Minus.Start,fill=conditType))+geom_bar(stat="summary",position=dodge,fun="mean")+ 
  stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+facet_grid(~positionX)+ggtitle("e2 testing (asymptote - start performance)  ")
```

### Relative distance and under/overshooting

Reviewer 3
Absolute versus relative distance: From a methodological standpoint, I understand the need to differentiate these two types of distance. However, from a theoretical perspective there may be some issue in differentiating these two concepts. Schema theory relies on relative (or invariant) information to inform the motor program. However, both distances would be important to an instance or exemplar representation. You may want to consider commenting on this issue.

Reviewer 2
For the same reason, the plots showing improvement during training could be due to participants learning the task, rather than fine motor skills. Although task learning and motor learning are impossible to separate cleanly, the common practice in the field is indeed to offer practice trials to reduce the task learning aspects. The authors should address this.

In addition to absolute errors (which is related to variance), the authors should also provide other measures of performance, e.g., the mean of the signed errors, so that readers have a better idea whether there was any meaningful over- or undershooting.

#### experiment 1 training - relative distances
```{r,echo=FALSE,warning=FALSE}

# possible that scaling required loading special package from devtools
exp1Train <- e1 %>% filter(stage!="Transfer",mode==1) %>% group_by(positionX) %>%mutate(scaleDev=scale_this(AbsDistFromCenter)) %>%ungroup() %>% group_by(sbjCode,trainHalf,conditType,devianceDirection,positionX)
exp1Train = exp1Train %>% summarise(MeanTargetDistance=mean(AbsDistFromCenter,trim=.05),scaledDist=mean(scaleDev,trim=.05),meanRelativeDist=mean(relativeDeviance,trim=.05))

# ggplot(data = exp1Train, aes(x=stage, y=scaledDist)) + geom_boxplot(aes(fill=conditType),position=position_dodge(1)) +ggtitle("Training Performance - Experiment 1")

# manuscript plot - original
ggplot(data = exp1Train, aes(x=trainHalf, y=MeanTargetDistance)) + geom_boxplot(aes(fill=conditType),position=position_dodge(1)) +ylab("Mean Distance From Center Of Target") +xlab("Training Stage")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.5)+theme_classic() +ggtitle("Experiment 1 Training - Absolute Distance")

exp1Train %>% ggplot(aes(x=trainHalf,y=MeanTargetDistance))+
  geom_bar(aes(group=positionX,fill=positionX),stat="summary",fun=mean,position=dodge)+
  facet_wrap(~devianceDirection,ncol=2)+
  stat_summary(aes(x=trainHalf,group=positionX),fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+
  ylab("Mean Distance From Center Of Target")+
  xlab("Training Stage")+theme(plot.title = element_text(hjust = 0.5))+
  guides(fill=guide_legend(title="Throw Location"))+theme(legend.title.align=.25)

ggplot(data = exp1Train, aes(x=trainHalf, y=meanRelativeDist)) + geom_boxplot(aes(fill=conditType),position=position_dodge(1)) +ylab("Mean Relative Distance From Center Of Target") +xlab("Training Stage")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.5)+theme_classic()+ggtitle("Experiment 1 Training - Relative Distance ")+facet_grid(.~devianceDirection)

ggplot(data = exp1Train, aes(x=trainHalf, y=MeanTargetDistance)) + geom_boxplot(aes(fill=conditType),position=position_dodge(1)) +ylab("Mean Distance From Center Of Target") +xlab("Training Stage")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.5)+theme_classic() +ggtitle("Experiment 1 Training - Absolute Distance")+facet_grid(.~devianceDirection)

t= e1 %>% filter(stage!="Transfer",mode==1)  %>% group_by(conditType,positionX,devianceDirection) %>% 
  summarise(Mean=round(mean(AbsDistFromCenter),2),sd=round(sd(AbsDistFromCenter),2),msd=paste(Mean,"(",sd,")",sep=""),.groups="keep") %>%
   select(-Mean,-sd) %>% spread(positionX,msd) %>% as.data.frame() %>%stargazer(.,type="text",summary=FALSE,rownames = FALSE)

t= e1 %>% filter(stage!="Transfer",mode==1)  %>% group_by(conditType,positionX) %>% 
  summarise(Mean=round(mean(relativeDeviance),2),sd=round(sd(AbsDistFromCenter),2),msd=paste(Mean,"(",sd,")",sep=""),.groups="keep") %>%
   select(-Mean,-sd) %>% spread(positionX,msd) %>% as.data.frame() %>%stargazer(.,type="text",summary=FALSE,rownames = FALSE)

t= e1 %>% filter(stage=="Transfer",mode==1)  %>% group_by(conditType,positionX) %>% 
  summarise(Mean=round(mean(relativeDeviance),2),sd=round(sd(AbsDistFromCenter),2),msd=paste(Mean,"(",sd,")",sep=""),.groups="keep") %>%
   select(-Mean,-sd) %>% spread(positionX,msd) %>% as.data.frame() %>%stargazer(.,type="text",summary=FALSE,rownames = FALSE)

# f=ftable(xtabs(MeanTargetDistance  ~ conditType+positionX+devianceDirection, data = exp1Train))
# stargazer(f,type="text",summary=FALSE,rownames = FALSE)

```

#### experiment 2 training - relative distances
```{r,echo=FALSE,warning=FALSE}
# possible that scaling required loading special package from devtools
exp2Train <- e2 %>% filter(stage!="Transfer",mode==1) %>% group_by(Group2,sbjCode,conditType,Group) %>%mutate(scaleDev=scale_this(AbsDistFromCenter)) %>%ungroup() %>% group_by(Group2,sbjCode,stage,conditType,devianceDirection)
exp2Train = exp2Train %>% summarise(MeanTargetDistance=mean(AbsDistFromCenter),scaledDist=mean(scaleDev),meanRelativeDist=mean(relativeDeviance,trim=.05))
exp2Train$stage <- factor(exp2Train$stage, levels = c("Beginning", "Middle", "End")) #in case the levels get out of order

#ggplot(data = exp2Train, aes(x=stage, y=scaledDist)) + geom_boxplot(aes(fill=Group2),position=position_dodge(1)) +ggtitle("")

# manuscript plots

par(mfrow=c(1,2))
ggplot(data = exp2Train, aes(x=stage, y=meanRelativeDist)) + geom_boxplot(aes(fill=conditType),position=position_dodge(1)) +ylab("Mean Relative Distance From Center Of Target") +xlab("Training Stage")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Group"))+theme(legend.title.align=.25)+theme_classic()+ggtitle("Experiment 2 Training - Relative Distance")+facet_grid(.~devianceDirection)

ggplot(data = exp2Train, aes(x=stage, y=meanRelativeDist)) + geom_boxplot(aes(fill=Group2),position=position_dodge(1)) +ylab("Mean Distance From Center Of Target") +xlab("Training Stage")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+theme_classic()+ggtitle("Experiment 2 Training - Relative Distance")+facet_grid(.~devianceDirection)

```

#### Experiment 1 Testing - relative distances
```{r,echo=FALSE,warning=FALSE,fig.width=10,fig.height=7}
exp1Transfer3 <- e1 %>% filter(stage=="Transfer") %>% group_by(conditType2,sbjCode,positionX,ThrowPosition,devianceDirection) %>% summarise(MeanTargetDeviance = mean(AbsDistFromCenter),meanRelativeDist=mean(relativeDeviance))

exp1Transfer3_raw <- e1 %>% filter(stage=="Transfer",mode==1,AbsDistFromCenter<1500,trialType!=44)

#manuscript plot
dodge <- position_dodge(width = 0.9)
e1Pos <- c("610","760","835","910")
e1Var <- paste("Varied Train Position","Constant Train Position", "Novel Position", "Varied Training Position")
e1Labels<- paste(c("610\n Varied Trained","760\n Constant Trained","835\n Novel Location","910\n Varied Trained"),sep="")

# exp1Transfer3 %>% ggplot(aes(x=positionX,y=MeanTargetDeviance,group=conditType2,fill=conditType2))+geom_bar(stat="summary",position=dodge)+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+theme_classic()+scale_x_discrete(name="Testing Location",breaks=e1Pos,labels=e1Labels)

# exp1Transfer3 %>% ggplot(aes(x=positionX,y=meanRelativeDist,group=conditType2,fill=conditType2))+geom_bar(stat="summary",position=dodge)+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+theme_classic()+scale_x_discrete(name="Testing Location",breaks=e1Pos,labels=e1Labels)+ggtitle("Experiment 1 Test -Relative Distance")
# 
# 
# 
# exp1Transfer3 %>% ggplot(aes(x=positionX,y=MeanTargetDeviance,group=conditType2,fill=conditType2))+geom_bar(stat="summary",position=dodge)+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+theme_classic()+scale_x_discrete(name="Testing Location",breaks=e1Pos,labels=e1Labels)+facet_grid(.~devianceDirection)

exp1Transfer3 %>% ggplot(aes(x=positionX,y=meanRelativeDist,group=conditType2,fill=conditType2))+geom_bar(stat="summary",position=dodge)+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+ylab("Mean Signed Deviation From Target Center") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+scale_x_discrete(name="Testing Location",breaks=e1Pos,labels=e1Labels)+facet_grid(.~devianceDirection)

t= exp1Transfer3  %>% 
  group_by(conditType2,positionX) %>% 
  summarise(Mean.unsigned=round(mean(MeanTargetDeviance),2),
            sd.unsigned=round(sd(MeanTargetDeviance),2),
            msdu=paste(Mean.unsigned,"(",sd.unsigned,")",sep=""),
            Mean.signed=round(mean(meanRelativeDist),2),
            sd.signed=round(sd(meanRelativeDist),2),
            msds=paste(Mean.signed,"(",sd.signed,")",sep=""),
            .groups="keep") %>% ungroup() %>%
   select(-Mean.unsigned,-sd.unsigned,-Mean.signed,-sd.signed) %>% 
  pivot_wider(names_from = positionX,values_from=c(msdu,msds)) %>% as.data.frame() %>%stargazer(.,type="text",summary=FALSE,rownames = FALSE)

t= exp1Transfer3  %>% 
  mutate(Condition=conditType2) %>%
  group_by(Condition,positionX) %>% 
  summarise(Mean.signed=round(mean(meanRelativeDist),2),
            sd.signed=round(sd(meanRelativeDist),2),
            msds=paste(Mean.signed,"(",sd.signed,")",sep=""),
            .groups="keep") %>% ungroup() %>%
   select(-Mean.signed,-sd.signed) %>% 
  pivot_wider(names_from = positionX,values_from=c(msds)) %>% as.data.frame() %>%stargazer(.,type="text",summary=FALSE,rownames = FALSE)

```

#### Experiment 2 Testing - relative distances
```{r,echo=FALSE,warning=FALSE,fig.width=8,fig.height=8}

exp2Transfer3 <- e2 %>% filter(stage=="Transfer") %>% group_by(conditType2,sbjCode,positionX,devianceDirection,Group) %>% summarise(MeanTargetDeviance = mean(AbsDistFromCenter,trim=.05),meanRelativeDist=mean(relativeDeviance)) %>% ungroup()
exp2Transfer3$ThrowPosition <- exp2Transfer3$positionX

exp2Transfer4 <- e2 %>% filter(stage=="Transfer",mode==1,AbsDistFromCenter<1500,trialType!=44) %>% group_by(Group2,sbjCode,positionX,devianceDirection,Group) %>% summarise(MeanTargetDeviance = mean(AbsDistFromCenter),meanRelativeDist=mean(relativeDeviance))

# exp2Transfer3 %>% ggplot(aes(x=ThrowPosition,y=meanRelativeDist,group=conditType2,fill=conditType2))+geom_bar(stat="summary",position=dodge,fun="mean")+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+theme_classic()+ggtitle("Experiment 2 Test - Relative Distance")
# 
# exp2Transfer3 %>% ggplot(aes(x=ThrowPosition,y=MeanTargetDeviance,group=conditType2,fill=conditType2))+geom_bar(stat="summary",position=dodge,fun="mean")+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+theme_classic()+facet_grid(.~devianceDirection)+ggtitle("Absolute Dist - Split by Over/Under")

exp2Transfer3 %>% ggplot(aes(x=ThrowPosition,y=meanRelativeDist,group=conditType2,fill=conditType2))+geom_bar(stat="summary",position=dodge,fun="mean")+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ylab("Mean Signed Deviation From Target Center") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+facet_grid(.~devianceDirection)

# 
# exp2Transfer4 %>% ggplot(aes(x=positionX,y=MeanTargetDeviance,group=Group2,fill=Group2))+geom_bar(stat="summary",position=position_dodge(),fun="mean")+ stat_summary(fun.data=mean_se,geom="errorbar",position=position_dodge())+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Position"))+theme(legend.title.align=.25)+theme_classic()+facet_grid(.~devianceDirection)+ggtitle("Absolute Dist - Split by Over/Under")
# 
# 
# exp2Transfer4 %>% ggplot(aes(x=positionX,y=meanRelativeDist,group=Group2,fill=Group2))+geom_bar(stat="summary",position=position_dodge(),fun="mean")+ stat_summary(fun.data=mean_se,geom="errorbar",position=position_dodge())+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Position"))+theme(legend.title.align=.25)+theme_classic()+facet_grid(.~devianceDirection)+ggtitle("Experiment 2 Test - Relative Distance")

# 
# t= exp2Transfer3  %>% 
#   mutate(Condition=conditType2) %>%
#   group_by(Condition,positionX) %>% 
#   summarise(Mean.signed=round(mean(meanRelativeDist),2),
#             sd.signed=round(sd(meanRelativeDist),2),
#             msds=paste(Mean.signed,"(",sd.signed,")",sep=""),
#             .groups="keep") %>% ungroup() %>%
#    select(-Mean.signed,-sd.signed) %>% 
#   pivot_wider(names_from = positionX,values_from=c(msds)) %>% as.data.frame() %>%stargazer(.,type="text",summary=FALSE,rownames = FALSE)
#

```

#### Experimenet 1 - intermittent testing
```{r,fig.width=9}
# first 3 trials - 1 throw from 610, 760 and 910
# interleaved every 20 trials, 2 throws from 610, 760, 910
# 10 full intermittent tests (6 trials each), plus 3 at beginning = 63 total per subject. 
#[1] 0  2  3  5  6  8  9  11 12 14 15 17 18 20 21 23 24 26 27 29 30
# First final testing trial is 264. So 1-263 = training and intermittent testing. 
# 
# intTest.third <- exp1.inter %>% filter(intStage!="int1")%>% group_by(sbjCode,conditType,stage,intStage,positionX) %>%
#   summarise(MeanTargetDistance=mean(AbsDistFromCenter,trim=.05)) %>% group_by(sbjCode,conditType,stage,positionX) %>%
#   summarise(MeanTargetDistance=mean(MeanTargetDistance))
# 
# intTest.third %>% ggplot(aes(x=positionX,y=MeanTargetDistance))+
#   geom_bar(aes(group=stage,fill=stage),stat="summary",fun=mean,position=dodge)+
#   facet_wrap(~conditType,ncol=2)+
#   stat_summary(aes(x=positionX,group=stage),fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+
#   ylab("Mean Distance From Center Of Target")+
#   xlab("Throw Location")+theme(plot.title = element_text(hjust = 0.5))+
#   guides(fill=guide_legend(title="Training Stage"))+theme(legend.title.align=.25)
# 
# intTest.half %>% anova_test(dv=MeanTargetDistance,between=conditType,within=trainHalf,covariate=positionX,wid=sbjCode)
# intTest.half %>% anova_test(dv=MeanTargetDistance,between=conditType,within=c(trainHalf,positionX),wid=sbjCode)

intTest.half <- readRDS(here::here("data/e1_intTest.rds"))

# intTest.half <- exp1.inter %>%filter(AbsDistFromCenter<1550)%>% 
#   group_by(sbjCode,conditType,trainHalf,positionX) %>%
#   summarise(MeanTargetDistance=mean(AbsDistFromCenter,trim=.01))

intTest.half %>% ggplot(aes(x=positionX,y=MeanTargetDistance))+
  geom_bar(aes(group=trainHalf,fill=trainHalf),stat="summary",fun=mean,position=dodge)+
  facet_wrap(~conditType,ncol=2)+
  stat_summary(aes(x=positionX,group=trainHalf),fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+
  ylab("Mean Distance From Center Of Target")+
  xlab("Intermittent Testing Throw Location")+theme(plot.title = element_text(hjust = 0.5))+
  guides(fill=guide_legend(title="Training Stage"))+theme(legend.title.align=.25)

cnames=c("Condition","610_First Half","760_First Half","910_First Half","610_Second Half","760_Second Half","910_Second Half")
test= intTest.half %>% rename(Condition="conditType") %>% group_by(Condition,trainHalf,positionX) %>% 
  summarise(Mean=round(mean(MeanTargetDistance),2),sd=round(sd(MeanTargetDistance),2)) 
test=test %>% group_by(Condition) %>% mutate(msd=paste(Mean,"(",sd,")",sep="")) %>%
 select(Condition,positionX,trainHalf,msd)%>% pivot_wider(names_from = c(positionX,trainHalf),values_from=c(msd))
test=test %>% as.data.frame()
colnames(test) <- cnames
stargazer(test,type="text",summary=FALSE,rownames=FALSE)

```

### Training plots - Experiment 1
```{r}
#| warning: false

# possible that scaling required loading special package from devtools
exp1Train <- e1 %>% filter(stage!="Transfer",mode==1) %>% group_by(Group,sbjCode) %>%mutate(scaleDev=scale_this(AbsDistFromCenter)) %>%ungroup() %>% group_by(Group,sbjCode,stage,conditType)
exp1Train = exp1Train %>% summarise(MeanTargetDistance=mean(AbsDistFromCenter),scaledDist=mean(scaleDev,trim=.05))
exp1Train$stage <- factor(exp1Train$stage, levels = c("Beginning", "Middle", "End")) #in case the levels get out of order
exp1TrainTrials <- e1 %>% filter(stage!="Transfer",mode==1,trialType!=44) %>% group_by(Group,sbjCode,positionX) %>% mutate(scaleDev=scale_this(AbsDistFromCenter),ind=1,trainIndex=cumsum(ind)) %>%ungroup() %>% group_by(Group,sbjCode,stage,conditType)

# exp1TrainTrials %>% filter(AbsDistFromCenter<1000) %>% ggplot(aes(x=trainIndex,y=AbsDistFromCenter,group=conditType2)) +geom_point()+geom_smooth(method="lm",se=FALSE) +facet_grid(~positionX)

# manuscript plot - original
ggplot(data = exp1Train, aes(x=stage, y=MeanTargetDistance)) + geom_boxplot(aes(fill=conditType),position=position_dodge(1))+stat_summary(fun="mean",aes(group=conditType),position=position_dodge(1))+
ylab("Mean Distance From Center Of Target") +xlab("Training Stage")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.5)+theme_classic()

lineplot.CI(data=exp1Train,x.factor=stage,group=conditType,response=scaledDist,xlab="Training Stage",x.leg=2,legend=TRUE,ylab="Distance from Target (scaled)",main="Training Performance - Experiment 1",col=c("red","black"))

lineplot.CI(data=exp1Train,x.factor=stage,group=conditType,response=MeanTargetDistance,xlab="Training Stage",x.leg=2,legend=TRUE,ylab="Distance From Target",main="Training Performance - Experiment 1",col=c("red","blue"))

```

#### Not in manuscript #####

# fit to testing performance averaged across positions
```{r}
e2Fits.AggPos <- readRDS(here::here('data/IGAS-e2Fits.AggPos-April_12.rds'))

e2Fits.AggPos %>% ggplot(aes(x=Group,y=pStart,fill=Group))+geom_bar(stat="summary",position=dodge,fun="mean")+ 
  stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ggtitle("experiment 2 - starting performance per position")

e2Fits.AggPos %>% ggplot(aes(x=Group,y=pAsym,fill=Group))+geom_bar(stat="summary",position=dodge,fun="mean")+ 
  stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ggtitle("e2 testing performance asymptote per position ")

e2Fits.AggPos %>% ggplot(aes(x=Group,y=pRate,fill=Group))+geom_bar(stat="summary",position=dodge,fun="mean")+ 
  stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ggtitle("e2 testing performance asymptote per position ")

#mr1=e2Fits.AggPos %>% anova_test(dv=pRate,between=conditType,wid=sbjCode,type=3);show(mr1)

```

#### statistical tests for starting performance
```{r}

ms1=e2TestFits %>% filter(converged==TRUE) %>% anova_test(dv=pStart,between=conditType,wid=sbjCode,within=positionX,type=3);
ms2=exp2.fit2 %>% anova_test(dv=MeanStart,between=conditType,wid=sbjCode,type=3);
#ms1
ms2

e2TestFits %>% ggplot(aes(x=conditType,y=pStart,fill=conditType))+geom_bar(stat="summary",position=dodge,fun="mean")+ 
  stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+facet_grid(~positionX)+ggtitle("experiment 2 - starting performance per position")

```

#### statistical tests for asymptote
```{r}

ma1=e2TestFits %>% filter(converged==TRUE)%>% anova_test(dv=pAsym,between=conditType,wid=sbjCode,within=positionX,type=3);
ma2=exp2.fit2 %>% anova_test(dv=MeanAsym,between=conditType,wid=sbjCode,type=3);
ma2

e2TestFits %>% ggplot(aes(x=conditType,y=pAsym,fill=conditType))+geom_bar(stat="summary",position=dodge,fun="mean")+ 
  stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+facet_grid(~positionX)+ggtitle("e2 testing performance asymptote per position ")

```