### 2024-02-28 index.qmd

---
title: "Dissertation"
author: Thomas E. Gorman
comments: false
date-format: 'MMM YYYY'
format:
  html:
    grid:
      sidebar-width: 250px
      body-width: 1400px
      margin-width: 170px
      gutter-width: 1.5rem
    self-contained: false
---

## Project 1

## Project 2

[Link to HTW project page](https://tegorman13.github.io/htw/){target="_blank"} \

[Working Draft of HTW Manuscript](https://tegorman13.github.io/htw/paper.html){target="_blank"}