### 2024-04-05 Sections/Discussion.qmd
fix minor formatting error in IGAS manuscript

---
title: General Discussion
date: today
---

Optional: - emphasize that both tasks involve extrapolation - it's just emphasized more in Project 2, and project 2 is designed to more closely mirror tasks used in the function learning literature.

## Project comparison

To facilitate ease of comparison between the two projects and their respective tasks, we'll now refer to project 1 as Hit The Target (HTT) and project 2 as Hit The Wall (HTW).

Task Structure - hitting target vs. hitting wall with correct force

Feedback - relation between feedback and correction - usefulness of feedback was part of what makes HTW different from HTT. And in general in visuomotor learning tasks, feedback usually is not quite as precise as it is in HTW. - consistency of HTW finding across continuous vs. ordinal makes it seem like this wasn't the sole determinative factor.

\ \ \ \ \

## General Discussion

To facilitate ease of comparison between the two projects and their respective tasks, we'll now refer to project 1 as Hit The Target (HTT) and project 2 as Hit The Wall (HTW).

### Summary

Across both projects, we investigated the influence of training variability on learning and generalization in computerized visuomotor skill learning, and function learning tasks. In project 1 (HTT), experiments 1 and 2 demonstrated that varied training led to superior testing performance compared to constant training. In Experiment 1, the varied group even outperformed the constant group even when testing from the constant groups trained position. In contrast, Project 2 (HTW) found the opposite pattern - the varied training groups exhibited poorer performance than the constant groups, both in terms of final training accuracy and the ability to accurately extrapolate to novel velocity bands during testing. This detrimental effect of variability was observed across three experiments, with the exception of the reverse order condition in Experiment 3, where the varied group was able to match the constant group's performance.

Both projects also contributed computational modeling componenents.

### Differences between the two Projects

Although the HTT and HTW tasks are similiar in requiring participants to learn how to correctly launch projectiles,  they nevertheless differ across numerous dimensions that may be relevant to the opposing patterns observed in the two projects (see @tbl-task-diff provides for a detailed comparison of the two tasks).

In HTT, the salient perceptual elements of the task (i.e. the launching box, target and barrier) are subject to variation (i.e. different distances between the launching box and target), and the spatial layout of these perceptually variable elements are intrinsically linked to the task objective of striking the target. Conversely, the perceptual task elements in HTW are invariant across trials, and the task objective is specified by the target velocity value specified as a numeral at the top of the screen. If the benefits of training variation do arise from the formation and flexible retrieval of distinct memory traces, then the lack of perceptual salience between training instances in the HTW task may have limited any potential benefits of variability. Future work could investigate this possibility further employing a modified version of the HTW task wherein the correct velocity value is indicated by some perceptual feature of the task (e.g. the color of the wall, or size of the ball), rather than displaying the target velocity numerically.

The HTT and HTW tasks also differed in terms of general task complexity. The HTT task was designed to mimic projectile launching tasks commonly employed in visuomotor learning studies, and the parabolic trajectories necessary to strike the target in HTT were sensitive to both the x and y dimensions of the projectiles velocity (and to a lesser extent, the position within the launching box at which the ball was released). Conversely the HTW task was influenced to a greater extent by the tasks commonly utilized in the function learning literature, wherein the correct output respones are determined by a single input dimension. *In HTW, the task space is also almost perfectly smooth, at least for the continuous feedback subjects, if they throw 100 units too hard, they'll be told that they were 100 units too hard. Whereas in HTT,  it was possible to produce xy velocity combinations that were technically closer to the empirical solution space than other throws, but which resulted in worse feedback due to striking the barrier.*

::: {#tbl-task-diff}

| Dimension                                       | HTT (Project 1)                                                                                                                               | HTW (Project 2)                                                                                                |
| ----------------------------------------------- | --------------------------------------------------------------------------------------------------------------------------------------------- | -------------------------------------------------------------------------------------------------------------- |
| Task Description                                | Projectile launching to hit a target                                                                                                          | Projectile launching to hit wall at a specific velocity                                                        |
| Task Complexity                                 | More complex parabolic trajectory, both x and y velocities relevant to outcome                                                                | Simpler 1D mapping of force to outcome. Only x velocity is relevant.                                           |
| Task Space                                      | More complex: xy velocity combinations closer to the solution space may still result in worse feedback due to striking the barrier.           | Simpler: smooth, linear mapping between velocity and feedback.                                                 |
| Perceptual salience of Varied Conditions | Varied conditions (# of throwing distances) are perceptually distinct, i.e. salient differences in distance between launching box and target. | Varied conditions (# of velocity bands) are less salient - only difference is the numeral displayed on screen. |
| Testing Feedback                                | Testing always included feedback                                                                                                              | Primary testing stage had no feedback.                                                                         |
| Potential for Learning during Testing           | Limited potential for learning during testing due to feedback.                                                                                | Some potential for learning during no-feedback testing by observing ball trajectory.                           |
| Position/Velocity Experience                    | Varied group gets half as much experience on any one position as the constant group.                                                          | Varied group gets 1/3 as much experience on any one velocity band as the constant group.                       |
| Testing Structure                               | Random interleaving of trained/transfer testing distances.                                                                                    | Blocked structure, separately testing trained vs extrapolation testing bands.                                  |

Comparison of the tasks in Project 1 (HTT) and Project 2 (HTW).
:::

\ \ \ \

Rework the section on "Differences between the two Projects" include more speculation on how the perceptual distinctiveness of the task conditions may have memory or attentional consequences that relate to when training variation is useful. In particular discussing connections to ideas from instance based, or exemplar-based memory models of learning (e.g. works of Gordan Logan; Nosofsky; Crump, Hintzman, Estes etc.)

### Empirical Results and Modeling Summary

This summary encapsulates the empirical investigations and computational modeling efforts within the Hit The Target (HTT) and Hit The Wall (HTW) projects, emphasizing the nuanced role of training variability on visuomotor learning and generalization.

**Project 1 (HTT)** explored spatial training variability's effects. **Empirical results** from Experiments 1 and 2 consistently showed that varied training enhances transfer performance, despite initial training challenges. The **Instance-Based Generalization and Adaptation (IGAS) model** was developed to interpret these results, offering a computational perspective that underscores the importance of specific training experiences in shaping generalization. By quantifying similarity across varied training instances, the IGAS model successfully explained why varied training could lead to superior transfer outcomes, validating the hypothesis that broader training experiences foster a more adaptable skill set.

**Project 2 (HTW)** focused on velocity training variability and its influence on function learning tasks. **Empirical findings** across three experiments demonstrated varied training's differential effects on extrapolation capabilities, with notable variations depending on the specific conditions and feedback types. The **Associative Learning Model (ALM) and EXemplar-based Associative Memory (EXAM) models** were applied to these findings, elucidating the mechanisms behind the observed patterns. ALM highlighted the role of associative mechanisms in learning from varied velocity training, while EXAM offered insights into how exemplar-based memory processes might support extrapolation. Together, these models provided a comprehensive understanding of how varied training influences learning and extrapolation, showcasing the complexity of underlying cognitive processes.

**Comparative analysis** between HTT and HTW, supported by the IGAS, ALM, and EXAM models, revealed the multifaceted nature of training variability's impact. While empirical data underscored the beneficial effects of variability on learning and generalization, the models offered a deeper understanding of the cognitive and memory-related mechanisms at play. Specifically, they suggested that the effectiveness of varied training might depend on the task's perceptual features and the cognitive operations required for task performance.

**Conclusions** drawn from both the empirical and modeling efforts highlight the critical role of training variability in enhancing learning outcomes. The integration of computational models with empirical data not only validates the observed benefits of varied training but also advances our understanding of the cognitive foundations underlying these effects. Future research directions may include refining these models to capture additional nuances of learning processes and exploring their applicability to other learning domains. This holistic approach, combining empirical research with computational modeling, promises to yield rich insights into the complexities of learning and generalization.