### 2024-04-12 Sections/Introduction.tex
adding more citations to intro

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[Numbers=Proportional,Numbers=OldStyle]{Linux Libertine O}
  \setmathfont[]{Libertinus Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% -----------------------
% CUSTOM PREAMBLE STUFF
% -----------------------

% -----------------
% Typography tweaks
% -----------------
% Indent size
\setlength{\parindent}{0.5in}
\setlength{\leftmargin}{0.5in}

% Fix widows and orphans
\usepackage[all,defaultlines=2]{nowidow}

% List things
\usepackage{enumitem}
% Same document-level indentation for ordered and ordered lists
\setlist[1]{labelindent=\parindent}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}

% Wrap definition list terms
% https://tex.stackexchange.com/a/9763/11851
\setlist[description]{style=unboxed}

% For better TOCs
\usepackage{tocloft}

% Remove left margin in lists inside longtables
% https://tex.stackexchange.com/a/378190/11851
\AtBeginEnvironment{longtable}{\setlist[itemize]{nosep, wide=0pt, leftmargin=*, before=\vspace*{-\baselineskip}, after=\vspace*{-\baselineskip}}}

% For fancy ORCID links
\usepackage{orcidlink}

% Indent all first paragraphs because APA wants that
\usepackage{indentfirst}

% -----------------
% Title block stuff
% -----------------

% Abstract
\usepackage{abstract}
\renewcommand{\abstractnamefont}{\normalfont\normalsize\bfseries}
\renewcommand{\abstracttextfont}{\normalsize}
\setlength{\absleftindent}{\parindent}
\setlength{\absrightindent}{\parindent}

% Keywords
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}}#1}

% Title
\usepackage{titling}
\setlength{\droptitle}{2\baselineskip}
\pretitle{\begin{center}\normalfont\normalsize\bfseries}
\posttitle{\par\end{center}\vspace{2\baselineskip}}

% ------------------
% Section headings
% ------------------
\usepackage{titlesec}
\titleformat*{\section}{\center\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}
\titleformat*{\subsubsection}{\normalsize\bfseries\itshape}
\titleformat{\paragraph}[runin]{\bfseries}{\theparagraph.}{3pt}{\space}[.]
\titleformat{\subparagraph}[runin]{\bfseries\itshape}{\thesubparagraph.}{3pt}{\space}[.]

% \titlespacing{<command>}{<left>}{<before-sep>}{<after-sep>}
% Starred version removes indentation in following paragraph
\titlespacing*{\section}{0em}{2em}{0em}
\titlespacing*{\subsection}{0em}{1em}{0em}
\titlespacing*{\subsubsection}{0em}{0em}{0em}
\titlespacing*{\paragraph}{\parindent}{1ex}{1em}
\titlespacing*{\subparagraph}{\parindent}{1ex}{1em}

% -----------
% Footnotes
% -----------
% NB: footmisc has to come after biblatex because of conflicts
\usepackage[bottom]{footmisc}
\renewcommand*{\footnotelayout}{\footnotesize}

\addtolength{\skip\footins}{12pt}  % vertical space between rule and main text
\setlength{\footnotesep}{16pt}  % vertical space between footnotes

% ----------
% Captions
% ----------
\usepackage[font=normalsize]{caption}

% ---------------------------
% END CUSTOM PREAMBLE STUFF
% ---------------------------
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Introduction},
  colorlinks=true,
  linkcolor={DarkSlateBlue},
  filecolor={Maroon},
  citecolor={DarkSlateBlue},
  urlcolor={DarkSlateBlue},
  pdfcreator={LaTeX via pandoc}}

% -----------------------
% END-OF-PREAMBLE STUFF
% -----------------------

% For patching commands like \subtitle
\usepackage{etoolbox}

% -----------------
% Running headers
% -----------------

\usepackage{fancyhdr}
\setlength{\headheight}{0.25in}
\renewcommand{\headrulewidth}{0pt}  % Remove lines
\renewcommand{\footrulewidth}{0pt}

% SHORT TITLE           Page #
\fancypagestyle{normal}{
  \fancyhf{}
  \lhead{\uppercase{ Introduction }}
  \rhead{\thepage}
}

% Running head: SHORT TITLE        Page #
\fancypagestyle{title}{
  \fancyhf{}
  \lhead{Running head: \uppercase{ Introduction }}
  \rhead{}
}

% Use regular heading style
\pagestyle{normal}

% ----------
% BibLaTeX
% ----------
\usepackage[style=apa,backend=biber]{biblatex}

\setlength\bibitemsep{0pt}  % No space between bib entries
\setlength\bibhang{\parindent}  % Match document indentation

% Fix biblatex's odd preference for using In: by default.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{}\intitlepunct}}}

\addbibresource{../Assets/Bib/Dissertation.bib}

% Start bibliography on new page
\pretocmd{\printbibliography}{\clearpage}{}{}

% ---------------------- 
% Title block elements
% ---------------------- 
\usepackage{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{ and }
\renewcommand*{\Authands}{, and }
\renewcommand*{\Affilfont}{\normalsize}
\renewcommand*{\Authfont}{\normalsize}

\title{Introduction}

\date{}

% Typeset URLs in the same font as their parent environment
%
% This has to come at the end of the preamble, after any biblatex stuff because 
% some biblatex styles (like APA) define their own \urlstyle{}
\usepackage{url}
\urlstyle{same}

% ---------------------------
% END END-OF-PREAMBLE STUFF
% ---------------------------
\begin{document}
% ---------------
% TITLE SECTION
% ---------------

\begin{titlepage}
\center

% Don't include a newpage before \maketitle
{\let\newpage\relax\maketitle}

% Needs to come after \maketitle
\thispagestyle{title}

\vspace{0.25in}

\vspace{0.5in}

\begin{center}

\textbf{Author note}

\end{center}

\vfill  % Fill the rest of the page with whitespace

\end{titlepage}

\doublespacing

\begin{center}
\singlespacing
\textbf{Introduction}
\end{center}

% -------------------
% END TITLE SECTION
% -------------------

\setstretch{2}
\section{Introduction}\label{introduction}

The past century of research on human learning has produced ample
evidence that although learners can improve at almost any task, such
improvements are often specific to the trained task, with unreliable or
even nonexistent transfer or generalization to novel tasks or conditions
\autocite{barnettWhenWhereWe2002,dettermanCaseProsecutionTransfer1993}.
Such generalization challenges are of noteworthy practical relevance,
given that educators, trainers, and rehabilitators typically intend for
their students to be able to apply what they have learned to new
situations. It is therefore important to better understand the factors
that influence generalization, and to develop cognitive models that can
predict when generalization is likely to occur. Such characteristics
have included training difficulty, spacing, temporal order, feedback
schedules, and the primary focus of the current work - the variability
of training examples.

\subsection{The study of variability}\label{the-study-of-variability}

Varied training has been shown to influence learning in wide array of
different tasks and domains, including categorization
\autocite{hahnEffectsCategoryDiversity2005,maddoxStimulusRangeDiscontinuity2011,posnerGenesisAbstractIdeas1968,nosofskyModelguidedSearchOptimal2019,morgensternOneshotCategorizationNovel2019,plebanekEffectsFrequencyVariability2021},
language learning
\autocite{jonesDensityDistinctivenessEarly2020,perryLearnLocallyThink2010,twomeyAllRightNoises2018,wonnacottInputEffectsAcquisition2012,brekelmansDoesHighVariability2022},
anagram completion \autocite{goodeSuperiorityVariableRepeated2008},
trajectory extrapolation
\autocite{fulvioTaskSpecificResponseStrategy2014}, cognitive control
tasks
\autocite{sabahWhenLessMore2019,moshon-cohenStimulusVariabilityImproves2024},
associative learning
\autocite{leeEvidentialDiversityIncreases2019,reichmannVariabilityAbstractionEvaluative2023,fanStimulusDiversityIncreases2022,pradaExperiencedCategoryVariability2020},
visual search
\autocite{georgeStimulusVariabilityTask2021,gonzalezDiversityTrainingEnhances2011,kelleyLearningAttendEffects2009},
voice identity learning \autocite{lavanEffectsHighVariability2019}, face
recognition
\autocite{honigPerceptualSimilarityModulates2022,burtonIdentityVariationRepresentations2016,menonVariationPhotosSame2015},
the perception of social group heterogeneity
\autocite{linvilleExemplarAbstractionModels1993,konovalovaInformationSamplingExplanation2020,parkPerceptionVariabilityCategory1987,gershmanStructureLearningPrinciples2023}
, simple motor learning
\autocite{braunMotorTaskVariation2009,kerrSpecificVariedPractice1978,rollerVariablePracticeLenses2001,willeyLimitedGeneralizationVaried2018},
sports training
\autocite{greenPracticeVariabilityTransfer1995a,northEffectConsistentVaried2019,breslinConstantVariablePractice2012},
and complex skill learning
\autocite{seowTransferEffectsVaried2019,huetEducationAttentionExplanation2011,hacquesVisualControlClimbing2022}.
See \textcite{czyzVariabilityPracticeInformation2021} or
\textcite{ravivHowVariabilityShapes2022} for more detailed reviews.

Research on the effects of varied training typically manipulates
variability in one of two ways. In the first approach, a high
variability group is exposed to a greater number of unique instances
during training, while a low variability group receives fewer unique
instances with more repetitions. Alternatively, both groups may receive
the same number of unique instances, but the high variability group's
instances are more widely distributed or spread out in the relevant
psychological space, while the low variability group's instances are
clustered more tightly together. Researchers then compare the training
groups in terms of their performance during the training phase, as well
as their generalization performance during a testing phase. Researchers
will usually compare the performance of the two groups during both the
training phase and a subsequent testing phase. The primary theoretical
interest is often to assess the influence of training variability on
generalization to novel testing items or conditions. However, the test
may also include some or all of the items that were used during the
training stage, allowing for an assessment of whether the variability
manipulation influenced the learning of the trained items themselves, or
to make it easy to measure how much performance degrades as a function
of how far away testing items are from the training items.

The influence of training variation has received a large amount of
attention in the domain of sensorimotor skill learning. Much of this
research has been influenced by the work of
\textcite{schmidtSchemaTheoryDiscrete1975}, who proposed a schema-based
account of motor learning as an attempt to address the longstanding
problem of how novel movements are produced. Schema theory presumes that
learners possess general motor programs for a class of movements
(e.g.~an underhand throw). When called up for use motor programs are
parameterized by schema rules which determine how the motor program is
parameterized or scaled to the particular demands of the current task.
Schema theory predicts that variable training facilitates the formation
of more robust schemas, which will result in improved generalization or
transfer. Experiments that test this hypothesis are often designed to
compare the transfer performance of a constant-trained group against
that of a varied-trained group. Both groups train on the same task, but
the varied group practices with multiple instances along some
task-relevant dimension that remains invariant for the constant group.
For example, studies using a projectile throwing task might assign
participants to either constant training that practicing throwing from a
single location, and a varied group that throws from multiple locations.
Following training, both groups are then tested from novel throwing
locations
\autocite{pigottMotorSchemaStructure1984,willeyLimitedGeneralizationVaried2018,pachecoLearningSpecificIndividual2018,wulfEffectTypePractice1991}.

One of the earliest, and still often cited investigations of Schmidt's
benefits of variability hypothesis was the work of
\textcite{kerrSpecificVariedPractice1978}. Two groups of children, aged
8 and 12, were assigned to either constant or varied training of a bean
bag throwing task. The constant group practiced throwing a bean-bag at a
small target placed 3 feet in front of them, and the varied group
practiced throwing from a distance of both 2 feet and 4 feet.
Participants were blindfolded and unable to see the target while making
each throw but would receive feedback by looking at where the beanbag
had landed in between each training trial. 12 weeks later, all of the
children were given a final test from a distance of 3 feet which was
novel for the varied participants and repeated for the constant
participants. Participants were also blindfolded for testing and did not
receive trial by trial feedback in this stage. In both age groups,
participants performed significantly better in the varied condition than
the constant condition, though the effect was larger for the younger,
8-year-old children. This result offers a particularly compelling
example of the merits of varied practice, given that the varied group
was able to outperform the constant group even from the home turf
location where one may have expected the constatn group to have the
strongest advantage. A similar pattern of results was observed in
another study wherein varied participants trained with tennis, squash,
badminton, and short-tennis rackets were compared against constant
subjects trained with only a tennis racket
\autocite{greenPracticeVariabilityTransfer1995a}. One of the testing
conditions had subjects repeat the use of the tennis racket, which had
been used on all 128 training trials for the constant group, and only 32
training trials for the varied group. Nevertheless, the varied group
outperformed the constant group when using the tennis racket at testing,
and also performed better in conditions with several novel racket
lengths. However, as is the case with many of the patterns commonly
observed in the ``benefits of variability'' literature, the pattern
wherein the varied group outperfroms the constant group even from the
constants group's home turf has not been consistently replicated. One
recent study attempted a near replication of the Kerr \& Booth study
\autocite{willeyLongtermMotorLearning2018}, having subjects throw
beanbags at a target, with the varied group training from positions (5
and 9 feet) on either side of the constant group (7 feet). This study
did not find a varied advantage from the constant training position,
though the varied group did perform better at distances novel to both
groups. However, this study diverged from the original in that the
participants were adults; and the amount of training was much greater
(20 sessions with 60 practice trials each, spread out over 5-7 weeks).

Pitting varied against constant practice against each other on the home
turf of the constant group provides a compelling argument for the
benefits of varied training, as well as an interesting challenge for
theoretical accounts that posit generalization to occur as some function
of distance. However, despite its appeal this particular contrast is
relatively uncommon in the literature. It is unclear whether this may be
cause for concern over publication bias, or just researchers feeling the
design is too risky. A far more common design is to have separate
constant groups that each train exclusively from each of the conditions
that the varied group encounters
\autocite{catalanoDistantTransferCoincident1984a,chuaPracticeVariabilityPromotes2019,newellVariabilityPracticeTransfer1976,moxleySchemaVariabilityPractice1979,mccrackenTestSchemaTheory1977},
or for a single constant group to train from just one of the conditions
experienced by the varied participants
\autocite{pigottMotorSchemaStructure1984,rollerVariablePracticeLenses2001,wrisbergTrainingProductionNovel1984,wrisbergDevelopingCoincidentTiming1983}.
A less common contrast places the constant group training in a region of
the task space outside of the range of examples experienced by the
varied group, but distinct from the transfer condition
\autocite{wrisbergVariabilityPracticeHypothesis1987,wulfVariabilityPracticeImplicit1997}.
Of particular relevance to the current work is the early study of
\textcite{catalanoDistantTransferCoincident1984a}, as theirs was one of
the earliest studies to investigate the influence of varied vs.~constant
training on multiple testing locations of graded distance from the
training condition. Participants were trained on coincident timing task,
in which subjects observe a series of lightbulbs turning on sequentially
at a consistent rate and attempt to time a button response with the
onset of the final bulb. The constant groups trained with a single
velocity of either 5,7,9, or 11 mph, while the varied group trained from
all 4 of these velocities. Participants were then assigned to one of
four possible generalization conditions, all of which fell outside of
the range of the varied training conditions -- 1, 3, 13 or 15 mph. As is
often the case, the varied group performed worse during the training
phase. In the testing phase, the general pattern was for all
participants to perform worse as the testing conditions became further
away from the training conditions, but since the drop off in performance
as a function of distance was far less steep for the varied group, the
authors suggested that varied training induced a decremented
generalization gradient, such that the varied participants were less
affected by the change between training and testing conditions.

Benefits of varied training have also been observed in many studies
outside of the sensorimotor domain.
\textcite{goodeSuperiorityVariableRepeated2008} trained participants to
solve anagrams of 40 different words ranging in length from 5 to 11
letters, with an anagram of each word repeated 3 times throughout
training, for a total of 120 training trials. Although subjects in all
conditions were exposed to the same 40 unique words (i.e.~the solution
to an anagram), participants in the varied group saw 3 different
arrangements for each solution-word, such as DOLOF, FOLOD, and OOFLD for
the solution word FLOOD, whereas constant subjects would train on three
repetitions of LDOOF (spread evenly across training). Two different
constant groups were used. Both constant groups trained with three
repetitions of the same word scramble, but for constant group A, the
testing phase consisted of the identical letter arrangement to that seen
during training (e.g.~LDOOF), whereas for constant group B, the testing
phase consisted of a arrangement they had not seen during training, thus
presenting them with a testing situation similar situation to the varied
group. At the testing stage, the varied group outperformed both constant
groups, a particularly impressive result, given that constant group A
had 3 prior exposures to the word arrangement (i.e.~the particular
permutation of letters) which the varied group had not explicitly seen.
However varied subjects in this study did not exhibit the typical
decrement in the training phase typical of other varied manipulations in
the literature, and actually achieved higher levels of anagram solving
accuracy by the end of training than either of the constant groups --
solving 2 more anagrams on average than the constant group. This might
suggest that for tasks of this nature where the learner can simply get
stuck with a particular word scramble, repeated exposure to the
identical scramble might be less helpful towards finding the solution
than being given a different arrangement of the same letters. This
contention is supported by the fact that constant group A, who was
tested on the identical arrangement as they experienced during training,
performed no better at testing than did constant group B, who had
trained on a different arrangement of the same word solution -- further
suggesting that there may not have been a strong identity advantage in
this task.

In the domain of category learning, the constant vs.~varied comparison
is much less suitable. Instead, researchers will typically employ
designs where all training groups encounter numerous stimuli, but one
group experiences a greater number of unique exemplars
\autocite{wahlheimMetacognitiveJudgmentsRepetition2012,nosofskyModelguidedSearchOptimal2019,doyleMetacognitiveMonitoringCategory2016,hoschPriorExperienceVariability2023,brunsteinPreparingNoveltyDiverse2011},
or designs where the number of unique training exemplars is held
constant, but one group trains with items that are more dispersed, or
spread out across the category space
\autocite{posnerGenesisAbstractIdeas1968,homaCategoryBreadthAbstraction1976,huHighvariabilityTrainingDoes2024,bowmanTrainingSetCoherence2020,maddoxStimulusRangeDiscontinuity2011}.

Much of the earlier work in this sub-area trained subjects on artificial
categories, such as dot patterns
\autocite{homaCategoryBreadthAbstraction1976,posnerGenesisAbstractIdeas1968}.
A seminal study by \textcite{posnerGenesisAbstractIdeas1968} trained
participants to categorize artificial dot patterns, manipulating whether
learners were trained with low variability examples clustered close to
the category prototypes (i.e.~low distortion training patterns), or
higher-variability patterns spread further away from the prototype
(i.e.~high-distortion patterns). Participants that received training on
more highly-distorted items showed superior generalization to novel high
distortion patterns in the subsequent testing phase. It should be noted
that unlike the sensorimotor studies discussed earlier, the
\textcite{posnerGenesisAbstractIdeas1968} study did not present
low-varied and high-varied participants with an equal number of training
rathers, but instead had participants remain in the training stage of
the experiment until they reached a criterion level of performance. This
train-until-criterion procedure led to the high-variability condition
participants tending to complete a larger number of training trials
before switching to the testing stage. More recent work
\autocite{huHighvariabilityTrainingDoes2024}, also used dot pattern
categories, but matched the number of training trials across conditions.
Under this procedure, higher-variability participants tended to reach
lower levels of performance by the end of the training stage. The
results in the testing phase were the opposite of
\textcite{posnerGenesisAbstractIdeas1968}, with the low-variability
training group showing superior generalization to novel high-distortion
patterns (as well as generalization to novel patterns of low or medium
distortion levels). However, whether this discrepancy is solely a result
of the different training procedures is unclear, as the studies also
differed in the nature of the prototype patterns used.
\textcite{posnerGenesisAbstractIdeas1968} utilized simpler, recognizable
prototypes (e.g., a triangle, the letter M, the letter F), while
\textcite{huHighvariabilityTrainingDoes2024} employed random prototype
patterns.

Recent studies have also begun utilizing more complex or realistic
sitmuli when assessing the influence of variability on category
learning. \textcite{wahlheimMetacognitiveJudgmentsRepetition2012}
conducted one such study. In a within-participants design, participants
were trained on bird categories with either high repetitions of a few
exemplars, or few repetitions of many exemplars. Across four different
experiments, which were conducted to address an unrelated question on
metacognitive judgements, the researchers consistently found that
participants generalized better to novel species following training with
more unique exemplars (i.e.~higher variability), while high repetition
training produced significantly better performance categorizing the
specific species they had trained on. A variability advantage was also
found in the relatively complex domain of rock categorization
\autocite{nosofskyModelguidedSearchOptimal2019}. For 10 different rock
categories, participants were trained with either many repetitions of 3
unique examples of each category, or few repetitions of 9 unique
examples, with an equal number of total training trials in each group
(the design also included 2 other conditions less amenable to
considering the impact of variation). The high-variability group,
trained with 9 unique examples, showed significantly better
generalization performance than the other conditions.

A distinct sub-literature within the category learning domain has
examined how the variability or dispersion of the categories themselves
influences generalization to ambiguous regions of the category space
(e.g.~the region between the two categories). The general approach is to
train participants with examples from a high variability category and a
low variability category. Participants are then tested with novel items
located within ambiguous regions of the category space which allow the
experimenters to assess whether the difference in category variability
influenced how far participants generalize the category boundaries.
\textcite{cohenCategoryVariabilityExemplar2001} trained subjects on two
categories, one with much more variability than the other. In experiment
1, a low variability category composed of 1 instance was compared
against a high-variability category of 2 instances in one condition, and
7 instances in another. In experiment 2 both categories were composed of
3 instances, but for the low-variability group the instances were
clustered close to each other, whereas the high-variability groups
instances were spread much further apart. Participants were tested on an
ambiguous novel instance that was located in between the two trained
categories. Both experiments provided evidence that participants were
much more likely to categorize the novel middle stimulus into a category
with greater variation.

Further observations consonant with the results of
\textcite{cohenCategoryVariabilityExemplar2001} have since been observed
in numerous investigations
\autocites{hahnEffectsCategoryDiversity2005,perlmanFurtherAttemptsClarify2012,sakamotoPuttingPsychologyBack2008,hsuEffectsGenerativeDiscriminative2010}[but
see][ and
\textcite{yangCategoryVariabilityEffect2014}]{stewartEffectCategoryVariability2002}.
The results of \textcite{sakamotoPuttingPsychologyBack2008} are
noteworthy. They first reproduced the basic finding of participants
being more likely to categorize an unknown middle stimulus into a
training category with higher variability. In a second experiment, they
held the variability between the two training categories constant and
instead manipulated the training sequence, such that the examples of one
category appeared in an ordered fashion, with very small changes from
one example to the other (the stimuli were lines that varied only in
length), whereas examples in the alternate category were shown in a
random order and thus included larger jumps in the stimulus space from
trial to trial. They found that the middle stimulus was more likely to
be categorized into the category that had been learned with a random
sequence, which was attributed to an increased perception of variability
which resulted from the larger trial to trial discrepancies.

The work of \textcite{hahnEffectsCategoryDiversity2005}, is also of
particular interest to the present work. Their experimental design was
similar to previous studies, but they included a larger set of testing
items which were used to assess generalization both between the two
training categories as well as novel items located in the outer edges of
the training categories. During generalization testing, participants
were given the option to respond with ``neither'', in addition to
responses to the two training categories. The ``neither'' response was
included to test how far away in the stimulus space participants would
continue to categorize novel items as belonging to a trained category.
Consistent with prior findings, high-variability training resulted in an
increased probability of categorizing items in between the training
categories as belong to the high variability category. Additionally,
participants trained with higher variability also extended the category
boundary further out into the periphery than participants trained with a
lower variability category were willing to do. The author compared a
variety of similarity-based models based around the Generalized Context
Model
\autocite{nosofskyAttentionSimilarityIdentificationcategorization1986}
to account for their results, manipulating whether a response-bias or
similarity-scaling parameter was fit separately between variability
conditions. No improvement in model fit was found by allowing the
response-bias parameter to differ between groups, however the model
performance did improvement significantly when the similarity scaling
parameter was fit separately. The best fitting similarity-scaling
parameters were such that the high-variability group was less sensitive
to the distances between stimuli, resulting in greater similarity values
between their training items and testing items. This model accounted for
both the extended generalization gradients of the varied particpants,
and also for their poorer performance in a recognition condition.

Variability has also been examined in the learning of higher-order
linguistic categories \autocite{perryLearnLocallyThink2010}. In nine
training sessions spread out over nine weeks infants were trained on
object labels in a naturalistic play setting. All infants were
introduced to three novel objects of the same category, with
participants in the ``tight'' condition being exposed to three similar
exemplars of the category, and participants in the varied condition
being exposed to three dissimilar objects of the same category.
Importantly, the similarity of the objects was carefully controlled for
by having a separate group of adult subjects provide pairwise similarity
judgements of the category objects prior to the study onset.
Multidimensional scaling was then performed to obtain the coordinates of
the objects psychological space, and out of the 10 objects for each
category, the 3 most similar objects were selected for the tight group
and the three least similar objects for the varied group, with the
leftover four objects being retained for testing. By the end of the nine
weeks, all of the infants had learned the labels of the training
objects. In the testing phase, the varied group demonstrated superior
ability to correctly generalize the object labels to untrained exemplars
of the same category. More interesting was the superior performance of
the varied group on a higher order generalization task -- such that they
were able to appropriately generalize the bias they had learned during
training for attending to the shape of objects to novel solid objects,
but not to non-solids. The tight training group, on the other hand,
tended to overgeneralize the shape bias, leading the researchers to
suggest that the varied training induced a more context-sensitive
understanding of when to apply their knowledge.

Of course, the relationship between training variability and transfer is
unlikely to be a simple function wherein increased variation is always
beneficial. Numerous studies have found null, or in some cases negative
effects of training variation
\autocite{deloshExtrapolationSineQua1997,sinkeviciuteRoleInputVariability2019,wrisbergVariabilityPracticeHypothesis1987,vanrossumSchmidtSchemaTheory1990},
and many more have suggested that the benefits of variability may depend
on additional factors such as prior task experience, the order of
training trials, or the type of transfer being measured
\autocite{bernikerEffectsTrainingBreadth2014,braithwaiteEffectsVariationPrior2015,hahnEffectsCategoryDiversity2005,lavanEffectsHighVariability2019,northEffectConsistentVaried2019,sadakataIndividualAptitudeMandarin2014,zamanPerceptualVariabilityImplications2021}.

In an example of a more complex influence of training variation,
\autocite{braithwaiteEffectsVariationPrior2015} trained participants on
example problems involving the concept of sampling with replacement
(SWR). Training consisted of examples that were either highly similar in
their semantic context (e.g.~all involving people selecting objects) or
in which the surface features were varied between examples (e.g.~people
choosing objects AND objects selected in a sequence). The experimenters
also surveyed how much prior knowledge each participant had with SWR.
They found that whether variation was beneficial depended on the prior
knowledge of the participants -- such that participants with some prior
knowledge benefited from varied training, whereas participants with
minimal prior knowledge performed better after training with similar
examples. The authors hypothesized that in order to benefit from varied
examples, participants must be able to detect the structure common to
the diverse examples, and that participants with prior knowledge are
more likely to be sensitive to such structure, and thus to benefit from
varied training. To test this hypothesis more directly, the authors
conducted a 2nd experiment, wherein they controlled prior knowledge by
exposing some subjects to a short graphical or verbal pre-training
lesson, designed to increase sensitivity to the training examples.
Consistent with their hypothesis, participants exposed to the structural
sensitivity pre-training benefited more from varied training than the
controls participants who benefited more from training with similar
examples. Interactions between prior experience and the influence of
varied training have also been observed in sensorimotor learning
\autocite{guadagnoliRelationshipContextualInterference1999,delreyEffectsContextualInterference1982}.
\textcite{delreyEffectsContextualInterference1982} recruited
participants who self-reported either extensive, or very little
experience with athletic activities, and then trained participants on a
coincident timing task under with either a single constant training
velocity, with one of several varied training procedures.
Unsurprisingly, athlete participants had superior performance during
training, regardless of condition, and training performance was superior
for all subjects in the constant group. Of greater interest is the
pattern of testing results from novel transfer conditions. Among the
athlete-participants, transfer performance was best for those who
received variable training. Non-athletes showed the opposite pattern,
with superior performance for those who had constant training.

\subsubsection{The current work}\label{the-current-work}

The overarching purpose of this dissertation is to investigate the
effects of training variability on learning and generalization within
visuomotor skill learning and function learning. Our investigations is
structured into two main projects, each employing distinct experimental
paradigms and computational modeling frameworks to elucidate how and
when variability in training enhances or impedes subsequent
generalization.

In Project 1, we investigated the influence of varied practice in a
simple visuomotor projectile launching task. Experiments 1 and 2
compared the performance of constant and varied training groups to
assess potential benefits of variability on transfer to novel testing
conditions. To account for the observed empirical effects, we introduced
the Instance-based Generalization with Adaptive Similarity (IGAS) model.
IGAS provides a novel computational approach for quantifying the
similarity between training experiences and transfer conditions, while
also allowing for variability to influence the generalization gradient
itself.

Project 2 shifted focus to the domain of function learning by employing
a visuomotor extrapolation task. Across three experiments, we examined
how constant and varied training regimes affected learning,
discrimination between stimuli, and the ability to extrapolate to novel
regions of the function's input space. To model human performance in
this task, we fit the influential Associative Learning Model (ALM) and
the Extrapolation-Association Model (EXAM) to individual participant
data using advanced Bayesian parameter estimation techniques.

\printbibliography[title=References]

\end{document}