### 2024-04-12 Sections/pre_sections/cover.qmd

---
title: "Dissertation"
short-title: "Variability and Generalization"
date: 10-11-2023
author:
- name: Thomas Gorman
  affiliation: Indiana University
  url: www.tegorman13.github.io
  email: tegorman@iu.edu
  orcid: 0000-0001-5366-5442
---

<style>
  body {
    text-align: center;
  }
</style>

\
\
Thomas Gorman
\
\
\
\
\
\
\
\
\
\
\
\
\
\
\
\
\
\
\

Submitted to the faculty of the University Graduate School in partial fulfillment of the
requirements for the degree Doctor of Philosophy in the Department of Psychology and Brain
Sciences and the Cognitive Science Program, Indiana University
Indiana University
Fall 2023