### 2024-04-12 Sections/pre_sections/pre_docx.qmd

# The Role of Variability in Learning Transfer: A Similarity-Based Computational Approach

## Thomas E. Gorman

<!-- Submitted to the faculty of the University Graduate School in partial fulfillment of the
requirements for the degree Doctor of Philosophy in the Department of Psychology and Brain
Sciences and the Cognitive Science Program, Indiana University.

---

Accepted by the Graduate Faculty, Indiana University, in partial fulfillment of the
requirements for the degree of Doctor of Philosophy.

- Robert L. Goldstone, PhD
- Robert Nosofsky, PhD
- Peter Todd, PhD
- Mike Jones, PhD -->

---

## @2023

### Thomas E. Gorman

---

## Acknowledgements

---