### 2024-04-12 Sections/pre_sections/sigs.qmd

<style>
  body {
    text-align: center;
  }
</style>

Accepted by the Graduate Faculty, Indiana University, in partial fulfillment of the
requirements for the degree of Doctor of Philosophy.
\
\
\
\
\
\
\
Doctoral Committee \ \ 
___________________________ Robert L. Goldstone, Ph.D., Co-Chair \ \
___________________________ Robert Nosofsky, Ph.D., Co-Chair \ \
___________________________ Peter Todd, Ph.D.
___________________________ Mike Jones, Ph.D.

\ \ \ \ 
\ \ \ \ 
\ \ \ \

10/20/2023