### 2024-04-12 Sections/hold_temp.qmd
re-render

Schema theory and typical study design

Kerr & Booth pattern
Wrisberg - kerr & booth contrasts plus other comparisons
Green 1995 - partial support of Kerr & Booth - varied group gets less experience from testing scenario
Goode analogy study - reps Kerr & Booth

More common designs - lots of citations

catalano 1984 - description
Wiley 2018 - failed replication of Kerr & Booth

One such study that used a movement timing task, wherein subjects had to move their hand from a starting location, to a target location, attempting to arrive at the target location at specific time following the onset of a cue [@wrisbergVariabilityPracticeHypothesis1987]. This study utilized 4 different constant groups, and 3 varied groups, with one of the constant groups training under conditions identical to the testing conditions, and which were not trained on by any of the varied groups, e.g. the design of Kerr and Booth. However, in this case the varied group did not outperform the constant group.

A consequence of varied training is that the trial to trial sequences will tend to to consist of fewer repeats, and more switches between task condition (i.e. throwing locations; category exemplars; target velocitys), or in the case where variability is being manipulated by the dispersion of training items, the trial to trial sequences of varied participants will tend to consist of larger jumps in the stimulus space.

 The number of switches can vary greatly, with the two extremes being varied participants completing all of their training trials in one block before switching to the next condition (blocked sequencing), or if they alternate between conditions on a trial by trial basis (random/intermixed/interleaved sequencing).

  [@sheaContextualInterferenceEffects1979]. In this seminal study, both groups of training subjects trained with the same number of trials of three separate movement patterns. A blocked group that completed all of their trials with one sequence before beginning the next sequence, and a random group that trained with all three movement patterns interspersed throughout the course of training. Participants were also randomly assigned to retention testing under either blocked or random sequence conditions, thus resulting in all four training-testing combinations of blocked-blocked; blocked-random; random-blocked; random-random. There was some effect of sequence context, such that both groups performed better when the testing sequence matched their training sequence. However, the main finding of interest was the advantage of random-training, which resulted in superior testing performance than blocked training regardless of whether the testing stage had a blocked or random sequence, an effect observed both immediately after training, and in a follow up test ten days after the end of training. A more recent study [@chuaPracticeVariabilityPromotes2019], found a similar pattern of results in both a varied vs. constant training manipulation (expt 1, bean-bag throwing task), and of random training vs. blocked training manipulation (expt 2 & 3, bean-bag throwing & golf putting). The novelty of this study is that the experimenters queried subjects about their attentional focus throughout the training stage. In all three experiments varied or random trained-subjects reported significantly greater external attention (e.g. attending to the target distance), and constant or blocked subjects reported more internal attention (e.g. posture or hand position). The authors argue that the benefits of varied/random training may thusly be mediated by changes in attentional focus.

### Sequence Effects

A necessary consequence of varied training is that participants will have the experience of switching from one task condition to another. The number of switches can vary greatly, with the two extremes being varied participants completing all of their training trials in one before switching to the next condition (blocked sequencing), or if they alternate between conditions on a trial by trial basis (random/intermixed/interleaved sequencing). Not long after the initial influx of schema-theory inspired studies testing the benefits of variability hypothesis was shown that the influence of varied training might interact with the type of training sequence chosen by the experimenter  [@sheaContextualInterferenceEffects1979]. In this seminal study, both groups of training subjects trained with the same number of trials of three separate movement patterns. A blocked group that completed all of their trials with one sequence before beginning the next sequence, and a random group that trained with all three movement patterns interspersed throughout the course of training. Participants were also randomly assigned to retention testing under either blocked or random sequence conditions, thus resulting in all four training-testing combinations of blocked-blocked; blocked-random; random-blocked; random-random. There was some effect of sequence context, such that both groups performed better when the testing sequence matched their training sequence. However, the main finding of interest was the advantage of random-training, which resulted in superior testing performance than blocked training regardless of whether the testing stage had a blocked or random sequence, an effect observed both immediately after training, and in a follow up test ten days after the end of training.

Prior to the influential @sheaContextualInterferenceEffects1979 study, studies investigating the benefits of variability hypothesis had utilized both blocked and random training schedules, often without comment or justification. It was later observed [@leeInfluencePracticeSchedule1985] that positive evidence for benefits of varied training seemed more likely to occur for studies that utilized random schedules. The theoretical basis of such studies was invariably an appeal to Schmidt's schema theory; however schema theory made no clear predictions of an effect of study sequence on retention or generalization, thus prompting the need for alternate accounts. One such account, the elaborative processing account [@sheaContextEffectsMemory1983], draws on the earlier work [@battigFacilitationInterference1966] and argues that randomly sequencing conditions during training promotes comparison and contrastive processes between those conditions, which result in a deeper understanding of the training task than could arise via blocked sequencing. Supporting evidence for elaborative processing comes in the form of random-sequence trained subjects self-reporting more nuanced mental representations of movement patterns following training [@sheaContextEffectsMemory1983], and by manipulating whether subjects are able to perform comparisons during training [@wrightContributionElaborativeProcessing1992]. An alternative, though not incompatible account suggests that the benefits of random-sequencing are a result of such sequences forcing the learner to continually reconstruct the relevant motor task in working memory [@leeLocusContextualInterference]. Blocked training, on the other hand, allows the learner to maintain the same motor task in short term memory without decay for much of the training which facilitates training performance, but hinders ability to retrieve the appropriate motor memory in a later testing context. A much more recent study [@chuaPracticeVariabilityPromotes2019], replicates the standard findings of an advantage of varied training over constant training (expt 1, bean-bag throwing task), and of random training over blocked training (expt 2 & 3, bean-bag throwing & golf putting). The novelty of this study is that the experimenters queried subjects about their attentional focus throughout the training stage. In all three experiments varied or random trained-subjects reported significantly greater external attention (e.g. attending to the target distance), and constant or blocked subjects reported more internal attention (e.g. posture or hand position). The authors argue that the benefits of varied/random training may be mediated by changes in attentional focus, however the claims made in the paper seem to go far beyond what can be justified by the analyses reported -- e.g. the increased external focus could be a simple byproduct of varied training. A stronger form of evidence that was not provided may have been to use multiple regression analyses to show that the testing advantage of the varied/random groups over the constant/blocked groups could be accounted for by the differences in self-reported attentional focus.

In North 2019 - biggest effect for varied-random. 
North, J. S., Bezodis, N. E., Murphy, C. P., Runswick, O. R., Pocock, C., & Roca, A. (2019). The effect of consistent and varied follow-through practice schedules on learning a table tennis backhand. Journal of Sports Sciences, 37(6), 613–620. https://doi.org/10.1080/02640414.2018.1522683

## The study of variability

Studies investigating the "benefits of variability" hypothesis usually assign participants to either a constant or varied group for the training stage of the experiment. Then, subjects in both groups complete an identical testing stage which often consists items/conditions seen during training, and novel items/conditions.

If the varied group performs better in the testing stage, this is taken for evidence of the benefits of variability hypothesis. Even within this relatively straightforward between-groups design, researchers must navigate several crucial methodological choices, highlighted below:

1)  Variables Subject to Variation. In multidimensional tasks, researchers have the option to vary numerous variables. The experimenters must decide the specific dimension(s) across which variation will occur. For instance, in a projectile throwing accuracy task -- researchers might vary the distance from the target, the size of the target, the weight of the projectile. They might also vary a contextual variable not directly relevant to the task, but which will still be encoded by the subject on a trial by trial basis, e.g. the background color.

2)  Magnitude of Variation, relative to the control condition. The simplest comparison would be to compare a constant group who trains with 1 example/condition, against a varied group that trains from 2 examples/conditions. However, it is not uncommon in the literature for the varied condition to train from 3 or 4 conditions. For example, @catalanoDistantTransferCoincident1984a train varied subjects from 4 different velocities in their coincident timing task, and [@goodeSuperiorityVariableRepeated2008] have varied subjects' practice with 3 different variants (i.e. different letter scrambles of the same word) of an anagram for a given word, while their constant participants view the same variant 3 different times. Alternatively, rather than a constant vs. varied comparison, subjects in all conditions might experience a variety of training items, but with one group experiencing a greater number of unique items [@nosofskyModelguidedSearchOptimal2018].

3)  Locations within Task-Space. For tasks in which the stimuli or conditions fall within a continuous metric space, the experimenter must decide whether the varied instances are relatively close together (e.g. throwing a ball from a distance of 4 feet and 5 feet), or far apart (throwing from 4 feet and 20 feet). Spreading the varied training items further apart may be beneficial in terms of providing a more representative sample of the task space to the learner, however large distances may also result in significant differences in difficulty between the training examples, which can be a common confound in variability studies.

4)  Proximity of Testing to Training Conditions. Intuitively, the fairest form of comparison is to include testing conditions that are of an equivalent distance from both the varied and constant groups. However researchers might also attempt to demonstrate the benefits of variation as being sufficiently powerful to outperform constant training, even in cases where the constant group trained from a closer proximity to the testing conditions, or whose training conditions are identical to the testing conditions [@goodeSuperiorityVariableRepeated2008; @kerrSpecificVariedPractice1978].

Training Condtions - how many, how different, how far apart

(note that the studies described above manipulated the variability of the training examples, rather than the variability of the categories)

 Moreover, this effect was at odds with the predications of the baseline version of the GCM, thus providing some evidence that training variation may at least sometimes induce effects that cannot be entirely accounted for by exemplar-similarity accounts.

## Schmidt

A noteworthy early study that explored the benefits of variability, as postulated by Schmidt, was conducted by Kerr & Booth (1978). In their experiment, two groups of children, ages 8 and 12, were tasked with a bean bag throwing exercise. The constant training group practiced aiming at a target 3 feet away, while the varied training group practiced from distances of both 2 and 4 feet. During this training phase, participants were blindfolded to prevent visual targeting, though they could check where the beanbag landed after each attempt for feedback. Twelve weeks later, all participants were tested at the novel distance of 3 feet, which was familiar to the constant group but new to the varied group. Again blindfolded, they received feedback only after the first half of their throws during the final test. Notably, both age groups exhibited significantly better performance in the varied condition, with a more pronounced effect in the younger cohort. Although this study design indirectly supports the hypothesis that varied training promotes better generalization than constant training, it robustly demonstrates the efficacy of varied practice without directly testing novel positions for the constant group.

## Literature Review

An early and influential work on the influence of variability on category learning is that of Posner & Keele (1968). In an ambitious attempt to address the question of how category information is represented, the authors trained participants to categorize artificial dot patterns, manipulating whether learners were exposed to examples clustered close to the category prototypes (e.g. a low variability condition), or spread further away from the prototype (the varied-training group). It should be noted that both groups in this study were trained with the same number of unique instances and the manipulated difference was how spread out the instances were. The authors claim based on prior experiments using the same stimuli, that the training stimuli for the varied group were at least as far away from the testing stimuli as the training stimuli of the less-varied group. The authors interpreted their findings as evidence for the extraction of an abstraction or schema that is extracted and stored, and then over time becomes more likely to be the reference point from which generalization occurs, given that specific instances are thought to decay at a faster rate than prototypes or schema. The Posner and Keele study has been extremely influential and continues to be cited in contemporary research as clear evidence that schema abstraction underlies the benefits of varied training. It's also referenced as a key influence in the development of "Schema Theory of Motor Learning" (Schmidt, 1975), which in turn influenced decades of investigations on the potential benefits of varied training in motor skill learning. However, the classic Posner & Keele study despite being far more carefully designed than many subsequent studies, and despite being a relative rarity in explicitly discussing and attempting to control for potential confounds of similarity between groups, may nevertheless be emblematic of a common issue in many investigations of the effects of varied training on learning. The problem with Posner & Keele's conclusion was demonstrated clearly almost 3 decades later (Palmeri & Nosofsky, 2001), when researchers conducting a near replication of the original study also collected similarity judgements following training and performed multidimensional scaling analysis. Rather than being in the middle of the training stimuli as was the case in the physical stimuli space, the psychological representation of the prototype was shown reside at an extreme point, and generalization patterns by participants that would have seemed to warrant the learning of a prototype were then easily accounted for with only the assumption that the participants encoded instances. One of the primary concerns of the present paper is that many of the studies which purport to explain the benefits of variation via prototypes, schemas, or other abstractions, are often overlooking the potential of instance based similarity accounts.

Numerous researchers have attempted to provide coherent frameworks to account for the full range of influences of training variation and sequencing described above (along with many other effects not discussed). Such accounts are generally quite similar, invoking ideas of desirable levels of difficulty [@bjorkNewTheoryDisuse1992; @schmidtNewConceptualizationsPractice1992], or optimal challenge points [@guadagnoliChallengePointFramework2004]. They tend to start by describing the dissociation between acquisition performance (performance during training) and testing performance (delayed retention and/or transfer), most strikingly observed as varied/random training participants performing worse than their constant/blocked counterparts during the training stage of the study, but then outperforming the constant/blocked comparisons at a later retention or transfer stage. This observation is then used to justify the idea that the most enduring and generalizable learning occurs by training at an optimal level of training difficulty, with difficulty being some function of the experience of the learner, and the cognitive or visuomotor processing demands of the task. It then follows that the factors that tend to make training more difficult (i.e. increased variability or randomization), are more likely to be beneficial when the learner has some experience, or when the processing demands of the task are not too extreme (which may only occur after some experience with the task). Such frameworks may be helpful heuristics in some cases, but they also seem to be overly flexible such that any null result of some intervention might be accounted for by a suboptimal amount of training trials, or by suggesting the training task was too difficult. The development of computational models that can account for how changes in the parameters of the motor-skill task scale with difficulty, would be a great step forward.

; @wulfPrinciplesDerivedStudy2002

Through this combination of controlled experiments and computational modeling, the present work provides an extensive investigation into if, how, and when variability facilitates transfer

Together, these projects contribute to a deeper understanding of the role of training variability in cognitive development, offering insights that could inform educational and training practices across a variety of domains. The findings and methodological approaches detailed in this dissertation not only advance theoretical knowledge but also provide practical implications for designing more effective learning environments.

By taking a multi-pronged approach that combines empirical investigations across multiple experimental paradigms with computational modeling, this dissertation provides a comprehensive examination of the complex and sometimes opposing effects of training variability on learning and generalization. The contrasting results obtained from the HTT and HTW tasks underscore the importance of considering task characteristics when designing experiments to assess the influence of training manipulations. Furthermore, the computational modeling components of this work offer novel theoretical accounts for both the beneficial and detrimental effects of training variability observed in our experiments. Collectively, this dissertation highlights the value of integrating empirical and modeling approaches to uncover the cognitive mechanisms that support learning and generalization across a variety of domains.

```{md}
---
title: Full Dissertation
subtitle: "The Role of Variability in Learning Transfer: A Similarity-Based Computational Approach"
date: 04/04/24
author:
- name: Thomas Gorman
  affiliation: Indiana University
page-layout: full
code-fold: true
code-tools: true
toc: true
toc-location: body
toc-title: Contents
cache: true
execute: 
  warning: false
  eval: true
format:
  html: default
  hugo-md:
    echo: false
    #include: false
    html-math-method: mathjax
    output-file: diss_full_hugo.md
  gfm:
    echo: false
    output: false
    output-file: diss_full_gfm.md
  hikmah-manuscript-pdf:
    echo: false
    output-file: "tg_full_disser.pdf"
    mainfont: "Linux Libertine O"
    mainfontoptions:
      - "Numbers=Proportional"
      - "Numbers=OldStyle"
    mathfont: "Libertinus Math"
# editor: 
#   markdown: 
#     wrap: 72
prefer-html: true
---

::: {.content-visible when-format="html"}

{{< include pre_sections/pre_html.qmd >}}

:::

::: {.content-visible when-format="pdf"}

{{< include pre_sections/pre_pdf.qmd >}}

:::

::: {.content-visible when-format="docx"}
{{< include pre_sections/pre_docx.qmd >}}
:::

{{< pagebreak >}}

::: column-page-inset-right

# Abstract

The past century of research on human learning has produced ample evidence that although learners can improve at almost any task, such improvements are often specific to the trained task, with unreliable or even nonexistent transfer or generalization to novel tasks or conditions [@barnettWhenWhereWe2002; @dettermanCaseProsecutionTransfer1993]. Such generalization challenges are of noteworthy practical relevance, given that educators, trainers, and rehabilitators typically intend for their students to be able to apply what they have learned to new situations. It is therefore important to better understand the factors that influence generalization, and to develop cognitive models that can predict when generalization is likely to occur. Such characteristics have included training difficulty, spacing, temporal order, feedback schedules, and the primary focus of the current work - the variability of training examples.

This dissertation seeks to explore the cognitive underpinnings that govern the generalization of learning, focusing specifically on the role of variability during training in shaping subsequent transfer performance. A comprehensive review of the existing literature is presented, emphasizing the methodological complications associated with disentangling the confounding effects of similarity. Through a series of experiments involving several novel visuomotor tasks, this work investigates whether and how variability in training conditions affects performance in novel tasks. To theoretically account for the empirical outcomes, I employ both instance-based and connectionist computational models, both of which incorporate similarity-based mechanisms. These models serve to account for the extent to which variability influences the learners' generalization gradient, and also explain how training variation can produce both beneficial and deleterious outcomes.

{{< pagebreak >}}

::: {.content-visible when-format="html"}

{{< include pre_sections/toc_html.qmd >}}

:::

::: {.content-visible when-format="pdf"}

{{< include pre_sections/toc_pdf.qmd >}}

:::

{{< pagebreak >}}

\ \

::: {.content-visible unless-format="pdf"}
# References

::: {#refs}
:::

:::

# Appendix

[Apppendix](Appendix/Full_Appendix.qmd) available at tegorman13.github.io/Dissertation/Sections/Appendix.html

::: {.content-visible when-format="pdf"}

# References

:::

:::

<!-- ::: {.content-visible unless-format="markdown"}
### Appendix - Project 1

{{< include Appendix/IGAS-Supplemental.qmd >}}

{{< pagebreak >}}

### Appendix - Project 2 - Experiment 1

{{< include Appendix/E1_Appendix.qmd >}}
:::
 -->
```

Write a complete paragraph for the "Existing Theoretical Frameworks" sections at the end of the Introduction. Cite papers like on Bjork, 1994; Soderstrom & Bjork, 2015; Guadagnoli & Lee, 2004, and make the point that while these existing frameworks offer a conceptual framework for understanding the influence of variability, there is currently a dearth of computational modelling approaches being utilized to explain the effects of varied training in the visuomotor learning domain.

### Existing Theoretical Frameworks

Several theoretical frameworks have been proposed to account for the effects of training variability on learning and generalization. The contextual interference effect (Battig, 1972; Shea & Morgan, 1979) posits that varied practice leads to greater interference during acquisition but enhanced retention and transfer due to increased contextual discrimination and cognitive effort during learning. Similarly, the reconceptualized desirable difficulties framework (Bjork, 1994; Soderstrom & Bjork, 2015) suggests that varied practice introduces desirable difficulties that initially impair performance but promote long-term retention and transfer. The challenge point hypothesis (Guadagnoli & Lee, 2004) proposes that optimal learning occurs when the difficulty of the task is matched to the learner's skill level, with varied practice serving to increase task difficulty and facilitate learning for more skilled individuals. While these frameworks offer conceptual explanations for the influence of variability, there is currently a paucity of computational modeling approaches that explicitly capture the mechanisms underlying the effects of varied training, particularly in the domain of visuomotor learning. Existing computational models have largely focused on category learning (e.g., Nosofsky, 1986) and perceptual learning (e.g., Petrov et al., 2005), with limited attention to the role of variability in visuomotor skill acquisition and generalization. Given the importance of understanding the cognitive processes involved in visuomotor learning and transfer, there is a need for computational models that can provide a more precise and quantitative account of how varied practice influences the acquisition and generalization of visuomotor skills.

Existing theoretical frameworks, such as the desirable difficulties framework (Bjork, 1994) and the challenge point framework (Guadagnoli & Lee, 2004), provide valuable conceptual insights into the influence of training variability on learning and generalization. These frameworks suggest that introducing appropriate levels of challenge or difficulty during training, such as through increased variability, can enhance long-term retention and transfer (Soderstrom & Bjork, 2015). However, while these frameworks offer a high-level understanding of the mechanisms underlying the benefits of varied training, there is currently a scarcity of computational modeling approaches being applied to elucidate the specific cognitive processes and representations involved in the effects of training variability, particularly in the domain of visuomotor learning. Computational models, such as connectionist or Bayesian approaches, have the potential to provide a more precise and quantitative account of how variability during training influences the acquisition and generalization of visuomotor skills. By specifying the underlying mechanisms and parameters, these models can generate testable predictions and offer a deeper understanding of the cognitive processes that give rise to the observed effects of varied training on visuomotor learning and transfer.

The influence of training variability on learning and generalization has been a topic of considerable interest within cognitive science, particularly in the context of motor learning and skill acquisition. Theoretical frameworks such as Bjork's (1994) "desirable difficulties" and Soderstrom and Bjork's (2015) elaboration on the role of varied practice highlight the potential benefits of introducing variability during training to enhance learning outcomes. These theories suggest that varying the conditions of practice can improve the adaptability and robustness of learned skills, a principle also supported by Guadagnoli and Lee's (2004) challenge point framework, which posits that optimal learning occurs when the difficulty of practice tasks is appropriately matched to the learner's capabilities. While these frameworks provide valuable insights into the mechanisms through which variability might enhance learning, they predominantly focus on general psychological and pedagogical principles. Notably, there is a relative scarcity of computational modeling approaches that explicitly address how varied training influences learning processes in the visuomotor domain. Such models are crucial for quantitatively testing theories and providing a more detailed mechanistic understanding of how variability affects the acquisition, retention, and generalization of motor skills. As such, the current research seeks to fill this gap by developing and applying computational models to empirically investigate the effects of training variability in visuomotor tasks, thereby providing a more nuanced understanding of the conditions under which variability facilitates or impedes learning and transfer.