### 2024-04-15 Manuscript/pre_sections/pre_html.qmd

<div style="text-align: center; margin-top: 50px; font-size: 24px; font-weight: bold;">
 The Role of Variability in Learning Generalization: A Computational Modeling Approach
</div>

<div style="text-align: center; margin-top: 300px; font-size: 32px;">
  Thomas E. Gorman
</div>