### 2024-04-15 Manuscript/pre_sections/pre_pdf.qmd

\begin{centering}
\LARGE
{The Role of Variability in Learning Generalization: A Computational Modeling Approach}

\vspace*{1.5cm}

\LARGE
{Thomas E. Gorman}

\vspace{16.5cm}

\end{centering}

\vspace{6cm}

\pagenumbering{gobble}

\newpage