### 2024-04-16 Manuscript/pre_sections/pre_pdf.qmd
fix pdf title; re-render secondary pages; hide into and dis

\begin{centering}
\LARGE
{The Role of Variability in Learning Generalization: A Computational Modeling Approach}

\vspace*{1.5cm}

\LARGE
{Thomas E. Gorman}

\vspace{16.5cm}

\end{centering}

<!-- Submitted to the faculty of the University Graduate School in partial fulfillment of the
requirements for the degree Doctor of Philosophy in the Department of Psychology and Brain
Sciences and the Cognitive Science Program, Indiana University
Indiana University -->

\vspace{6cm}

\pagenumbering{gobble}

\newpage