### 2024-04-22 Manuscript/pre_sections/pre_docx.qmd
add docx version; fix igas figure text; re-render everything.

# The Role of Variability in Learning Generalization: A Computational Modeling Approach

## Thomas E. Gorman