### 2024-05-18 Presentation/pres1.qmd
start creating slides

---
title: "The Role of Variability in Learning Generalization"
format: 
  revealjs:
    theme: moon #white
    slideNumber: true
    center: true
    chalkboard: true
author: Thomas Gorman
institute: Indiana University
---

## Introduction
- Investigate effects of training variability on learning and generalization
- Focus areas: visuomotor skill learning and function learning tasks

## Background and Motivation
- Importance of training variability
- Research gaps in visuomotor skill learning
- Previous research in category learning, language learning, and function learning

## Theoretical Frameworks
- Schema Theory (Schmidt, 1975)
- Desirable Difficulties Framework (Bjork & Bjork, 2011)
- Challenge Point Framework (Guadagnoli & Lee, 2004)

## Overview of Current Work
- Project 1: Visuomotor projectile launching task
- Project 2: Visuomotor extrapolation task

## Project 1 - Hit The Target (HTT)
- Influence of varied practice in a projectile launching task
- Experiments 1 & 2:
  - Constant vs. varied training groups
  - Performance metrics and transfer testing

## Project 1 - Methodology
- Task description and experimental design
- Participants and procedures
- IGAS model introduction

## Project 1 - Results and Discussion
- Varied training group's superior performance
- Computational modeling results with IGAS
- Theoretical implications

## Project 2 - Hit The Wall (HTW)
- Influence of varied practice in a function learning task
- Experiments 1, 2, and 3:
  - Training regimes and testing conditions
  - Learning, discrimination, and extrapolation performance metrics

## Project 2 - Methodology
- Task description and experimental design
- Participants and procedures
- Application of ALM and EXAM models

## Project 2 - Results and Discussion
- Training variability's mixed results
- Modeling analysis with ALM and EXAM
- Theoretical implications

## General Discussion
- Compare HTT and HTW
- Empirical findings summary
- Computational modeling contributions

## Theoretical Contributions
- Insights into training variability effects on learning and generalization
- Computational models applications

## Practical Implications
- Training programs design implications
- Future research considerations

## Limitations and Future Directions
- Study limitations
- Future research suggestions

## Conclusion
- Recap of key findings
- Final thoughts on training variability's role in learning

## Questions and Discussion
- Open floor for questions from the committee and audience