### 2024-05-18 Presentation/slides2.qmd
Add embeded view of slides; re-render major pages.

---
title: "The Role of Variability in Learning Generalization: A Computational Modeling Approach"
subtitle: "Dissertation Defense"
format: 
  html: default
author:
  - name: Thomas Gorman
    orcid: 0000-0000-0000-0000
    affiliations: Indiana University
---

::: {.content-visible when-format="html"}

<div class="ratio ratio-16x9"> <iframe src="/Presentation/slides_revealjs.html"></iframe> </div>

test
:::