### 2024-05-22 Presentation/slides2.qmd
Re-add missing Procedure section; fix appendix link

---
title: "The Role of Variability in Learning Generalization: A Computational Modeling Approach"
subtitle: "Dissertation Defense"
format: 
  html: default
author:
  - name: Thomas Gorman
    orcid: 0000-0000-0000-0000
    affiliations: Indiana University
---