### 2024-05-23 Presentation/_fig_creation.qmd
updates to slides

```{r}
pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,here)
walk(c("Display_Functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

# Function to simulate exponential learning data
sim_exp <- function(N, alpha, beta, r, t0 = 0, sdlog = .1) {
  t0 + (alpha + (beta-alpha) * exp(-r * N)) + rlnorm(length(N), 0, sdlog)
}

# Parameters for simulation
N <- 1:50  # Number of trials
alpha <- 5  # Initial performance level
beta <- 15  # improvement

# Parameters for two groups
r_fast <- 0.1  # Faster learning rate
r_slow <- 0.05  # Slower learning rate

# Simulate data for two groups
set.seed(123)  # For reproducibility
data_fast <- sim_exp(N, alpha-3, beta, r_fast)
data_slow <- sim_exp(N, alpha, beta, r_slow)

# Combine data into a single dataframe
data <- data.frame(
  Trial = rep(N, 2),
  Performance = c(data_fast, data_slow),
  Group = factor(rep(c("Low Variability", "High Variability"), each = length(N)), levels=c("Low Variability", "High Variability"))
)

# Plot the data using ggplot2
tr_p <- ggplot(data, aes(x = Trial, y = Performance, color = Group)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "Learning Curves for Low and High Variability Training",
    x = "Training Trial",
    y = "Error",
    color = "Group"
  ) + ylim(0,20) +  theme(
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# ggsave(here::here("Presentation/assets/train_pattern.png"), tr_p)
```

```{r}

# Simulate testing data for two groups
set.seed(123)  # For reproducibility
test_data_fast <- rnorm(length(N), mean = 5, sd = 1)
test_data_slow <- rnorm(length(N), mean = 2.5, sd = 1)

# Combine testing data into a single dataframe
test_data <- data.frame(
  Trial = rep(N, 2),
  Performance = c(test_data_fast, test_data_slow),
  Group = factor(rep(c("Low Variability", "High Variability"), each = length(N)), levels=c("Low Variability", "High Variability"))
)

# Plot the testing data using ggplot2
test_p <- ggplot(test_data, aes(x = Group, y = Performance, fill = Group)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge", width = 0.7) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(0.7), width = 0.25) +
  labs(
    title = "Testing Performance for Low and High Variability Training",
    x = "Group",
    y = "Error",
    fill = "Group"
  ) + ylim(0, 10) + theme(
    panel.grid.minor = element_blank()
  )

# ggsave(here::here("Presentation/assets/test_pattern.png"), test_p)
```