### 2024-05-27 Presentation/_fig_creation.qmd
slide updates

# HTW Results

```{r}

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  data.table, stringr, here,conflicted, gt, ggh4x, patchwork, knitr)
#options(brms.backend="cmdstanr",mc.cores=4)
walk(c("brms","dplyr","bayestestR","here"), conflict_prefer_all, quiet = TRUE)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("brms","dplyr","bayestestR","here"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","deLosh_data","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

e1 <- readRDS(here("data/e1_08-21-23.rds")) 
e1Sbjs <- e1 |> group_by(id,condit) |> summarise(n=n())
testE1 <- e1 |> filter(expMode2 == "Test")
nbins=5
trainE1 <-  e1 |> filter(expMode2=="Train") |> group_by(id,condit, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE1_max <- trainE1 |> filter(Trial_Bin == nbins, bandInt==800)
trainE1_avg <- trainE1_max |> group_by(id,condit) |> summarise(avg = mean(dist))

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

}
```

## HTW E1

```{r}
#| label: fig-e1-test-dev
#| fig-cap: Experiment 1 Testing Accuracy. A) Empricial Deviations from target band during testing without feedback stage. B) Conditional effect of condition (Constant vs. Varied) and testing band type (trained bands vs. novel extrapolation bands) on testing accuracy. Error bars represent 95% credible intervals.
#| fig-width: 4
#| fig-height: 4

# pe1td <- testE1 |>  
#   ggplot(aes(x = vb, y = dist,fill=condit)) +
#     stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
#     stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
#   theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
#   labs(x="Band", y="Deviation From Target Band")

pe1td <- testE1 |> ggplot(aes(x = vb, y = dist,fill=condit,
                               col=ifelse(bandType=="Trained","black","lightgray"),
                               size=ifelse(bandType=="Trained",1.0,.5))) +
  stat_bar + 
  theme(legend.title.position="top",
        legend.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation", fill="Condition") +
     # scale_color_manual(values = c("black" = "black"), guide = "legend") +
  scale_color_manual(values=c("black","lightgray"), 
                     name= "Band Type",
                     labels = c("Trained","Extrapolation"),
                     guide = guide_legend(override.aes = list(fill = NA))
                     ) +
  scale_size_continuous(range = c(0.8, 1.0), guide="none")

bmtd <- readRDS(paste0(here::here("data/model_cache/"), "e1_dist_Cond_Type_RF_2.rds"))
pe1ce <- bmtd |> emmeans( ~condit + bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Deviation", x="Band Type")

#p2 <- (pe1td + pe1ce) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e1_test-dev.png"), p2, width=8, height=4, bg="white")
#p2

e1_vxBMM <- readRDS(paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k.rds")))
pe1vce <- e1_vxBMM |> emmeans( ~condit + bandInt,re_formula=NA, 
                     at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
gather_emmeans_draws() |> 
condEffects(bandInt) +
stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  geom_abline(slope = 1, intercept = 200, linetype = "dashed", color = "gray", size = .5,alpha=.5) +
scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                   labels = levels(testE1$vb), 
                   limits = c(100, 1200)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(400,1400,by=200),2)) +
theme(legend.position="none") + 
  #theme(legend.title=element_blank(),legend.position="right") + 
labs(y="Velcoity", x="Band")

fe <- fixef(e1_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e1_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e1_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e1_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e1Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den1 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density")

pid_slopes1 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
     theme(legend.title=element_blank(), 
           axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(~condit,axes="all",scales="free_y")

pe1td
pe1ce

pe1vce
pid_den1
pid_slopes1

ggsave(here::here("Presentation/assets/htw_e1_test1.png"), pe1td, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test2.png"), pe1ce, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test3.png"), pe1vce, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test4.png"), pid_den1, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test5.png"), pid_slopes1, width = 4.0, height = 3, units = "in", dpi = 300)
```

## HTW Experiment 2

```{r}
#| label: fig-e2-test-dev
#| fig-cap: Experiment 2 Testing Accuracy. A) Empricial Deviations from target band during testing without feedback stage. B) Conditional effect of condition (Constant vs. Varied) and testing band type (trained bands vs. novel extrapolation bands) on testing accuracy. Error bars represent 95% credible intervals.
#| fig-width: 5
#| fig-height: 5
e2 <- readRDS(here("data/e2_08-04-23.rds")) 
e2Sbjs <- e2 |> group_by(id,condit) |> summarise(n=n())
testE2 <- e2 |> filter(expMode2 == "Test")
nbins=5
trainE2 <-  e2 |> filter(expMode2=="Train") |> group_by(id,condit, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE2_max <- trainE2 |> filter(Trial_Bin == nbins, bandInt==600)

bmtd2 <- readRDS(paste0(here::here("data/model_cache/"), "e2_dist_Cond_Type_RF_2.rds"))

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

}

pe2td <- testE2 |> ggplot(aes(x = vb, y = dist,fill=condit,
                               col=ifelse(bandType=="Trained","black","lightgray"),
                               size=ifelse(bandType=="Trained",1.0,.5))) +
  stat_bar + 
  theme(legend.title.position="top",
        legend.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation", fill="Condition") +
  scale_color_manual(values=c("black","lightgray"), 
                     name= "Band Type",
                     labels = c("Trained","Extrapolation"),
                     guide = guide_legend(override.aes = list(fill = NA))
                     ) +
  scale_size_continuous(range = c(0.8, 1.0), guide="none")

pe2ce <- bmtd2 |> emmeans( ~condit + bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type")

e2_vxBMM <- readRDS(paste0(here::here("data/model_cache", "e2_testVxBand_RF_5k.rds")))

pe2vce <- e2_vxBMM |> emmeans( ~condit + bandInt,re_formula=NA, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  geom_abline(slope = 1, intercept = 200, linetype = "dashed", color = "gray", size = .5,alpha=.5) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(100, 1200)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(200,1400,by=200),2)) +
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e2_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e2_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e2_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e2_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e2Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den2 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density")

pid_slopes2 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
      theme(legend.title=element_blank(), 
        axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(~condit,axes="all",scales="free_y")

pe2td
pe2ce
pe2vce
pid_den2
pid_slopes2

ggsave(here::here("Presentation/assets/htw_e2_test1.png"), pe2td, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test2.png"), pe2ce, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test3.png"), pe2vce, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test4.png"), pid_den2, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test5.png"), pid_slopes2, width = 4.0, height = 3, units = "in", dpi = 600)

```

## HTW Exp 3

```{r}
#| fig-width: 4
#| fig-height: 6
e3 <- readRDS(here("data/e3_08-04-23.rds")) |> 
    mutate(trainCon=case_when(
    bandOrder=="Original" ~ "800",
    bandOrder=="Reverse" ~ "600",
    TRUE ~ NA_character_
    ), trainCon=as.numeric(trainCon)) 
e3Sbjs <- e3 |> group_by(id,condit,bandOrder) |> summarise(n=n())
testE3 <- e3 |> filter(expMode2 == "Test")
nbins=5
trainE3 <-  e3 |> filter(expMode2=="Train") |> group_by(id,condit,bandOrder, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE3_max <- trainE3 |> filter(Trial_Bin == nbins, bandInt==trainCon)

bmtd3 <- readRDS(paste0(here::here("data/model_cache/"), "e3_dist_Cond_Type_RF_2.rds"))

pe3td <- testE3 |> ggplot(aes(x = vb, y = dist,fill=condit,
                               col=ifelse(bandType=="Trained","black","lightgray"),
                               size=ifelse(bandType=="Trained",1.0,.5))) +
  stat_bar + 
  facet_wrap(~bandOrder,ncol=1) +
  theme(legend.title.position="top",
        legend.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation", fill="Condition") +
  scale_color_manual(values=c("black","lightgray"), 
                     name= "Band Type",
                     labels = c("Trained","Extrapolation"),
                     guide = guide_legend(override.aes = list(fill = NA))
                     ) +
  scale_size_continuous(range = c(0.8, 1.0), guide="none")

pe3ce <- bmtd3 |> emmeans( ~condit *bandOrder*bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Deviation", x="Band Type") + 
 facet_wrap(~bandOrder,ncol=1)

e3_vxBMM <- readRDS(paste0(here::here("data/model_cache", "e3_testVxBand_RF_5k.rds")))

pe3vce <- e3_vxBMM |> emmeans( ~condit* bandOrder* bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  facet_wrap(~bandOrder,ncol=1) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  geom_abline(slope = 1, intercept = 200, linetype = "dashed", color = "gray", size = .5,alpha=.5) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE3$vb), 
                     limits = c(100, 1200)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(300,1400,by=200),2)) +
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e3_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e3_vxBMM)[,1]["bandInt"]
fixed_effect_interaction1 <- fixef(e3_vxBMM)[,1]["conditVaried:bandInt"]
fixed_effect_interaction2 <- fixef(e3_vxBMM)[,1]["bandOrderReverse:bandInt"]
fixed_effect_interaction3 <- fixef(e3_vxBMM)[,1]["conditVaried:bandOrderReverse:bandInt"]

re <- data.frame(ranef(e3_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e3Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction1*(condit=="Varied") + 
           fixed_effect_interaction2*(bandOrder=="Reverse") + 
           fixed_effect_interaction3*(condit=="Varied" & bandOrder=="Reverse"),
  slope = Estimate + adjust )

pid_den3 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density") +
  facet_wrap(~bandOrder,ncol=1)

pid_slopes3 <- re |>  
    mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
    geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
    theme(legend.title=element_blank(), 
      axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_nested_wrap(bandOrder~condit,axes="all",scales="free_y")

pe3td
pe3ce
pe3vce
pid_den3
pid_slopes3

ggsave(here::here("Presentation/assets/htw_e3_test1.png"), pe3td, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test2.png"), pe3ce, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test3.png"), pe3vce, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test4.png"), pid_den3, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test5.png"), pid_slopes3, width = 4.0, height = 6, units = "in", dpi = 600)

```

# Sim Example

```{r}
pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,here)
walk(c("Display_Functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

# Function to simulate exponential learning data
sim_exp <- function(N, alpha, beta, r, t0 = 0, sdlog = .1) {
  t0 + (alpha + (beta-alpha) * exp(-r * N)) + rlnorm(length(N), 0, sdlog)
}

# Parameters for simulation
N <- 1:50  # Number of trials
alpha <- 5  # Initial performance level
beta <- 15  # improvement

# Parameters for two groups
r_fast <- 0.1  # Faster learning rate
r_slow <- 0.05  # Slower learning rate

# Simulate data for two groups
set.seed(123)  # For reproducibility
data_fast <- sim_exp(N, alpha-3, beta, r_fast)
data_slow <- sim_exp(N, alpha, beta, r_slow)

# Combine data into a single dataframe
data <- data.frame(
  Trial = rep(N, 2),
  Performance = c(data_fast, data_slow),
  Group = factor(rep(c("Low Variability", "High Variability"), each = length(N)), levels=c("Low Variability", "High Variability"))
)

# Plot the data using ggplot2
tr_p <- ggplot(data, aes(x = Trial, y = Performance, color = Group)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "Learning Curves for Low and High Variability Training",
    x = "Training Trial",
    y = "Error",
    color = "Group"
  ) + ylim(0,20) +  theme(
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# ggsave(here::here("Presentation/assets/train_pattern.png"), tr_p)
```

```{r}

# Simulate testing data for two groups
set.seed(123)  # For reproducibility
test_data_fast <- rnorm(length(N), mean = 5, sd = 1)
test_data_slow <- rnorm(length(N), mean = 2.5, sd = 1)

# Combine testing data into a single dataframe
test_data <- data.frame(
  Trial = rep(N, 2),
  Performance = c(test_data_fast, test_data_slow),
  Group = factor(rep(c("Low Variability", "High Variability"), each = length(N)), levels=c("Low Variability", "High Variability"))
)

# Plot the testing data using ggplot2
test_p <- ggplot(test_data, aes(x = Group, y = Performance, fill = Group)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge", width = 0.7) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(0.7), width = 0.25) +
  labs(
    title = "Testing Performance for Low and High Variability Training",
    x = "Group",
    y = "Error",
    fill = "Group"
  ) + ylim(0, 10) + theme(
    panel.grid.minor = element_blank()
  )

# ggsave(here::here("Presentation/assets/test_pattern.png"), test_p)
```

# IGAS Task Space

```{r}
#| label: fig-taskSpace
#| fig.cap: A) A visual representation of the combinations of throw parameters (x and y velocities applied to the ball at launch), which resulted in target hits during the testing phase. This empirical solution space was compiled from all of the participants in experiment 2. B) shows the solution space within the context of all of the throws made throughout the testing phase of the experiment.
#| fig-width: 11
#| fig-height: 9

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,here, RColorBrewer, patchwork)
walk(c("Display_Functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

e2<- readRDS(here::here('data/igas_e2_cleanedData-final.rds')) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity)

taskspace <- e2 %>% filter(AbsDistFromCenter<900)
taskspace$hitOrMiss <- ifelse(taskspace$trialType==11,"Hit Target","Missed Target")

solSpace <- e2 %>% filter(trialType==11)
#solSpace %>% ggplot(aes(x=X_Velocity,y=Y_Velocity)) + geom_point(aes(colour=ThrowPosition),alpha=0.58) + ggtitle("")

solSpace$Result = ifelse(solSpace$ThrowPosition==400,"400",solSpace$ThrowPosition)
solSpace$Result = ifelse(solSpace$ThrowPosition==500,"500",solSpace$Result)
solSpace$Result= ifelse(solSpace$ThrowPosition==625,"625",solSpace$Result)
solSpace$Result = ifelse(solSpace$ThrowPosition==675,"675",solSpace$Result)
solSpace$Result = ifelse(solSpace$ThrowPosition==800,"800",solSpace$Result)
solSpace$Result = ifelse(solSpace$ThrowPosition==900,"900",solSpace$Result)

missSpace <- e2 %>% filter(trialType !=11, abs(X_Velocity)<1000)
missSpace$Result = "Missed Target"
solSpace$Result <- solSpace$Result

# the usual method of changing the legend title does not seem to work after the colours are manually scaled. 
# multiplied velocoties by -1 to make the axes less confusing
ss= solSpace %>% ggplot(aes(x=X_Velocity*-1,y=Y_Velocity*-1)) + 
  geom_point(aes(colour=Result),alpha=0.6) + 
  scale_color_manual(values =brewer.pal(n=6,name="Set1")) +
  labs(color="Throw position:", 
  # title="Empirical Solution Space - Testing Positions",
  # subtitle="The XY velocity coordinates of all the successful throws", 
  x="X Release Velocity", 
  y="Y Release Velocity", )  
  theme(legend.position="top")

fullSpace <- rbind(missSpace,solSpace)

fs<- fullSpace %>% ggplot(aes(x=X_Velocity*-1,y=Y_Velocity*-1,colour=Result)) + 
  geom_point(aes(),alpha=0.6) + scale_color_manual(values =brewer.pal(n=7,name="Set1"))+
  labs(color=element_blank(),
  # title="Full Space of Solution Throws and Missed Throws",
  # subtitle="Colored points represent coordinates of target hits, brown points are coordinates of misses",
  x="X Release Velocity", 
  y="Y Release Velocity") + 
  theme(legend.title = element_blank(), legend.position="top")

ggsave(here::here("Presentation/assets/sol_space.png"),ss)
ggsave(here::here("Presentation/assets/full_space.png"),fs)

```