### 2024-05-27 Presentation/embed_slides.qmd

::::: {.content-visible when-format="html" unless-format="revealjs"}
\
[{{< fa file-lines >}} Click here](../paper.qmd) to open the dissertation manuscript.\

<br><br><br><br><br><br>

## Slides {{< fa person-chalkboard size=1.0xs >}} {.unnumbered .unlisted visibility="hidden"}

Presentation slides are available below. You can navigate these sides using the <kbd>←</kbd> and <kbd>→</kbd> keys. Press <kbd>m</kbd> to display all commands (e.g., press <kbd>o</kbd> to switch to slide overview)

```{=html}
<p class="buttons" style="text-align:center;">
<a class="btn btn-danger" target="_blank" href="/Presentation/slides_revealjs.html#{{< meta params.start_slide >}}"><i class="fa-solid fa-arrow-up-right-from-square"></i> Open slides in new window </a>
</p>
```

:::: column-body-outset-right
::: {.ratio .ratio-4x3}
```{=html}
<iframe src="/Presentation/slides_revealjs.html#{{< meta params.start_slide >}}">
</iframe>
```
:::
::::

<br><br><br><br><br><br>

:::::