### 2024-05-27 Presentation/slides.qmd

---
title: "The Role of Variability in Learning Generalization: A Computational Modeling Approach"
page-layout: full
aliases:
  - /slides.html
css: /Assets/Style/slides.css
format: 
  html: 
    subtitle: "Dissertation Defense"
  clean-revealjs:
    #theme: moon #white
    #output-ext: revealjs.html
    # template-partials:
    #   - assets/title-slide.html
    output-file: slides_revealjs
    date: 05/28/204
    slideNumber: true
    hash-type: number # Use slide numbers in the URL hash
    center: true
    scrollable: true
    chalkboard:
      theme: whiteboard
  pptx:
    incremental: false
  hugo-md:
    include: false
    html-math-method: mathjax
    output-file: slides_hugo.md
  gfm:
    output: false
    output-file: slides_gfm.md
author:
  - name: Thomas Gorman
    url: https://tegorman13.github.io
    orcid: 0000-0001-5366-5442
    affiliations:
      - name: Indiana University
        department: Psychological & Brain Sciences, Cognitive Science
# params:
#   start_slide: 
#engine: knitr
---

{{< include /Presentation/embed_slides.qmd >}}

# Introduction

-   Investigate effects of training variability on learning and generalization
-   Focus areas: visuomotor skill learning and function learning tasks

------------------------------------------------------------------------

## Background and Motivation

-   Importance of training variability
-   Research gaps in visuomotor skill learning
-   Previous research in category learning, language learning, and function learning

------------------------------------------------------------------------

## Common Empirical Patterns

:::::::::  columns
::: {.column width="60%"}
### Training
- Both training conditions complete the same number of training trials.
- Varied group has worse training performance.

![](assets/train_pattern.png)
:::

::: {.column width="40%"} 
::: {.fragment}
### Testing

- Tested from novel conditions.
- Varied group has better test performance

![](assets/test_pattern.png)
:::
:::
:::::::::

------------------------------------------------------------------------

## But also plenty of contradictory results and complications

- Cases where varied training makes no difference
- Cases where more training variation results in worse outcomes
- Cases where the influence of variation interacts with some other factor
    - difficulty
    - prior knowledge
    - Frequency effects, or amount of training/learning before testing

------------------------------------------------------------------------

## Theoretical Frameworks

-   Schema Theory (Schmidt, 1975)
-   Desirable Difficulties Framework (Bjork & Bjork, 2011)
-   Challenge Point Framework (Guadagnoli & Lee, 2004)

<!-- ![](assets/fig-toy-model1-1-1.png) -->

------------------------------------------------------------------------

## Overview of Current Work

-   Project 1: Visuomotor projectile launching task
-   Project 2: Visuomotor extrapolation task

------------------------------------------------------------------------

# Project 1

-   Influence of varied practice in a projectile launching task
-   Experiments 1 & 2:
    -   Constant vs. varied training groups
    -   Performance metrics and transfer testing

------------------------------------------------------------------------

## Hit The Target Task

:::::: column-page-right
::::: {.columns}
:::: {.column width="60%"}

::: {.content-visible when-format="html"}
![](assets/htt_vid.mov){height="540" loop="true" autoplay="true" style="max-width: 100%"}
:::

::: {.content-visible unless-format="html"}
![](assets/methodsFig1.png)
:::

::::

:::: {.column width="40%"}
- **Training Stage** - 200 training trials. Constant groups throws from single position. Varied group 60 trials from 2 locations. 
- **Transfer Stage** - All subjects tested from both positions they were trained, and the positions trained by other group
- **Data recorded** - For every throw, recorded the X velocity and Y velocity of ball at release
::::

:::::
::::::

------------------------------------------------------------------------

## Project 1 - Manipulations

:::: {.columns}
::: {.column width="50%"}

### Experiment 1
- Train a varied group from 2 positions, and a constant group from 1. Then test both groups, from the training positions and a novel position.

:::

::: {.column width="50%"}
### Experiment 2
- 6 constant conditions, each trained from a unique location, and a varied condition trained from 2 locations
:::

::::

------------------------------------------------------------------------

## Project 1 - Results

:::: {.columns}
::: {.column width="40%"}
### Experiment 1
![](assets/fig-IGAS_Testing1-1-2.png)
:::
::: {.column width="60%"}
### Experiment 2
![](assets/fig-e2testa-1-2.png)
:::
::::

------------------------------------------------------------------------

## Project 1 - Discussion

-   Varied training group's superior performance
-   Computational modeling results with IGAS
-   Theoretical implications

------------------------------------------------------------------------

## Project 1 Computational Model {.smaller .r-fit-text}

## Model of Similarity between training and solutions

:::{#fig-solspace}
![](assets/sol_space.png){fig-align="left" width="500px"}

Solution space
:::

## Model of Similarity between training and solutions

:::: {.columns}
::: {.column width="70%"}
:::{#fig-fullspace}
![](assets/full_space.png){fig-align="left" width="700px"}

Full Space of Solution Throws and Missed Throws. Colored points represent coordinates of target hits, brown points are coordinates of misses
:::
:::
::: {.column width="30%"}
some text
:::
::::
------------------------------------------------------------------------

## Project 1 Computational Model {.smaller .r-fit-text}

- For each subject, compute similarity between training throws and each of the 6 testing locations
- Test whether similarity explains the difference in performance between Constant and Varied conditions
  - Assuming equivalent generalization (1 c for both groups) - similarity improves model fit, but does not explain group difference
  - Assume training condition influences generalization gradient (fit c separately) - now similarity does explain the difference between Constant and Varied.

::: footer
**Equation 1:** $$ Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}}) $$
**Equation 2:** $$ d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 } $$
:::

## Project 1 General Discussion {.smaller .r-fit-text}

------------------------------------------------------------------------

# Project 2 - Variability and Extrapolation in a Function Learning Task

-   Influence of varied practice in a function learning task
-   Experiments 1, 2, and 3:
    -   Training regimes and testing conditions
    -   Learning, discrimination, and extrapolation performance metrics

------------------------------------------------------------------------

## Project 2 - Questions and Goals

**Empirical**
- Design a task-space large enough to assess multiple degrees of extrapolation 
- Compare varied and constant generalization from several distinct distances from their nearest training condition

**Model-based**
- If variation does influence extrapolation, can a connectionist model with similarity-based activation provide a good account?
- Can our modelling framework simultaneously account for both training and testing data?
- Accounting for the full distribution of responses

------------------------------------------------------------------------

## Hit The Wall Task

<!-- {{< video assets/htw_vid.mov width="720" height="540" >}} -->

::: {.content-visible when-format="html"}
![](assets/htw_vid.mov){height="400" loop="true" autoplay="true"}
:::
-   A total of 156 participants from IU's online subject pool
-   Varied group trains from 3 "velocity bands", constant group from 1

::: {.content-visible unless-format="html"}
![](assets/htw_task_fig.png)
:::

------------------------------------------------------------------------

## Project 2 - Methodology

-   A total of 156 participants from IU's online subject pool
-   Varied group trains from 3 "velocity bands", constant group from 1

![](assets/htw_task_fig.png)

## Project 2 - Methodology

::: {.content-visible unless-format="markdown"}
```{dot}
//| label: fig-design-e1
//| fig-cap: "Experiment 1 Design. Constant and Varied participants complete different training conditions."
//| fig-width: 9.5
//| fig-height: 3.5
//| echo: false  
digraph {
  graph [layout = dot, rankdir = LR]
  node [shape = rectangle, style = filled]
  data1 [label = " Varied Training \n800-1000\n1000-1200\n1200-1400", fillcolor = "#00A08A"]
  data2 [label = " Constant Training \n800-1000", fillcolor = "#FF0000"]
  Test3 [label = "    Final Test \n  Novel With Feedback  \n100-300\n350-550\n600-800", fillcolor = "#ECCBAE"]
  data1 -> Test1
  data2 -> Test1
  subgraph cluster {
    label = "Test Phase \n(Counterbalanced Order)"
    Test1 [label = "Test  \nNovel Bands \n100-300\n350-550\n600-800", fillcolor = "#ECCBAE"]
    Test2 [label = "  Test \n  Varied Training Bands  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
    Test1 -> Test2
  }
  Test2 -> Test3
}
```
:::

------------------------------------------------------------------------

## Project 2 - Experiment 1 Results {.nostretch}

::: {.columns}
::: {.column width="50%"}
![Plot 1](assets/htw_e1_test1.png)
![Plot 3](assets/htw_e1_test3.png)
:::

::: {.column width="50%"}
![Plot 2](assets/htw_e1_test2.png)
![Plot 4](assets/htw_e1_test4.png)
:::
:::

## Project 2 - Experiment 2 Design

::: {.content-visible unless-format="markdown"}
```{dot}
//| label: fig-design-e2
//| fig-cap: "Experiment 2 Design. Constant and Varied participants complete different training conditions. The training and testing bands are the reverse of Experiment 1. "
//| fig-width: 10.0
//| fig-height: 4
//| fig-responsive: false
//| echo: false 
//| 
/////| column: screen-inset-right
digraph {
  graph [layout = dot, rankdir = LR]
  node [shape = rectangle, style = filled]
  data1 [label = " Varied Training \n100-300\n350-550\n600-800", fillcolor = "#00A08A"]
  data2 [label = " Constant Training \n600-800", fillcolor = "#FF0000"]
  Test3 [label = "    Final Test \n  Novel With Feedback  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
  data1 -> Test1
  data2 -> Test1
  subgraph cluster {
    label = "Test Phase \n(Counterbalanced Order)"
    Test1 [label = "Test  \nNovel Bands  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
    Test2 [label = "  Test \n  Varied Training Bands  \n100-300\n350-550\n600-800", fillcolor = "#ECCBAE"]
    Test1 -> Test2
  }
  Test2 -> Test3
}
```
:::

## Project 2 - Experiment 2 Results {.nostretch}

::: {.columns}
::: {.column width="50%"}
![Plot 1](assets/htw_e2_test1.png)
![Plot 3](assets/htw_e2_test3.png)
:::

::: {.column width="50%"}
![Plot 2](assets/htw_e2_test2.png)
![Plot 4](assets/htw_e2_test4.png)
:::
:::

## Project 2 - Experiment 3

- Ordinal Feedback
- Both Reverse and Original Feedback Order

## Project 2 - Experiment 3 Results {.nostretch}

## Accuracy

:::: {.columns}

::: {.column width="50%"}
![Plot 1](assets/htw_e3_test1.png)
:::

::: {.column width="50%"}
![Plot 3](assets/htw_e3_test2.png)
:::

::::

## Discrimination

:::: {.columns}

::: {.column width="50%"}
![Plot 1](assets/htw_e3_test3.png)
:::

::: {.column width="50%"}
![Plot 3](assets/htw_e3_test4.png)
:::

::::

------------------------------------------------------------------------

# Project 2 - Computational Modeling

## Project 2 - ALM

![](assets/fig-alm-diagram-1-1.png)

------------------------------------------------------------------------

## ALM

|                    | **ALM Response Generation**                                        |                                                                                               |
|------------------|-----------------------------|-------------------------|
| Input Activation   | $a_i(X) = \frac{e^{-c(X-X_i)^2}}{\sum_{k=1}^M e^{-c(X-X_k)^2}}$    | Input nodes activate as a function of Gaussian similarity to stimulus                         |
| Output Activation  | $O_j(X) = \sum_{k=1}^M w_{ji} \cdot a_i(X)$                        | Output unit $O_j$ activation is the weighted sum of input activations and association weights |
| Output Probability | $P[Y_j|X] = \frac{O_j(X)}{\sum_{k=1}^M O_k(X)}$                    | The response, $Y_j$ probabilites computed via Luce's choice rule                              |
| Mean Output        | $m(X) = \sum_{j=1}^L Y_j \cdot \frac{O_j(x)}{\sum_{k=1}^M O_k(X)}$ | Weighted average of probabilities determines response to X                                    |
|                    | **ALM Learning**                                                   |                                                                                               |
| Feedback           | $f_j(Z) = e^{-c(Z-Y_j)^2}$                                         | feedback signal Z computed as similarity between ideal response and observed response         |
| magnitude of error | $\Delta_{ji}=(f_{j}(Z)-o_{j}(X))a_{i}(X)$                          | Delta rule to update weights.                                                                 |
| Update Weights     | $w_{ji}^{new}=w_{ji}+\eta\Delta_{ji}$                              | Updates scaled by learning rate parameter $\eta$.                                             |

------------------------------------------------------------------------

## EXAM

|                    | **EXAM Response Generation**                                                                                     |                                                                                                               |
|--------------------|-----------------------------------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------|
| Instance Retrieval | $P[X_i|X] = \frac{a_i(X)}{\sum_{k=1}^M a_k(X)}$                                                                 | Novel test stimulus $X$ activates input nodes $X_i$                                                           |
| Slope Computation  | $S = \frac{m(X_{1})-m(X_{2})}{X_{1}-X_{2}}$                                                                     | Slope value, $S$ computed from nearest training instances                                                     |
| Response           | $E[Y|X_i] = m(X_i) + S \cdot [X - X_i]$                                                                         | Final EXAM response is the ALM response for the nearest training stimulus, $m(X_i)$, adjusted by local slope $S$. |

------------------------------------------------------------------------

## Project 2 - Model Fitting Procedure {.r-fit-text}

### Approximate Bayesian Computation

:::: {.columns}

::: {.column width="60%"}
![](assets/abc_diagram1.png)
:::

::: {.column width="40%"}

- simulation based approach
- approximate likelihood
- uncertainty in parameter values
- full distribution of plausible model predictions for each participant

:::

::::

## Modelling Results

![](assets/fig-ee-e1-1.png)

# General Discussion

-   Compare HTT and HTW
-   Empirical findings summary
-   Computational modeling contributions

------------------------------------------------------------------------

## Comparison {.smaller .r-fit-text}

| Dimension | HTT (Project 1) | HTW (Project 2) |
|----------------|--------------------------------|-------------------------|
| Task Description | Projectile launching to hit a target | Projectile launching to hit wall at a specific velocity |
| Task Complexity | More complex parabolic trajectory, both x and y velocities relevant to outcome | Simpler 1D mapping of force to outcome. Only x velocity is relevant. |
| Task Space | More complex: xy velocity combinations closer to the solution space may still result in worse feedback due to striking the barrier. | Simpler: smooth, linear mapping between velocity and feedback. |
| Perceptual salience of Varied Conditions | Varied conditions (# of throwing distances) are perceptually distinct, i.e. salient differences in distance between launching box and target. | Varied conditions (# of velocity bands) are less salient - only difference is the numeral displayed on screen. |
| Testing Feedback | Testing always included feedback | Primary testing stage had no feedback. |
| Potential for Learning during Testing | Limited potential for learning during testing due to feedback. | Some potential for learning during no-feedback testing by observing ball trajectory. |
| Training Experience | Varied group gets half as much experience on any one position as the constant group. | Varied group gets 1/3 as much experience on any one velocity band as the constant group. |
| Testing Structure | Random interleaving of trained/transfer testing distances. | Blocked structure, separately testing trained vs extrapolation testing bands. |

------------------------------------------------------------------------

## Theoretical Contributions

-   Insights into training variability effects on learning and generalization
-   Computational models applications

------------------------------------------------------------------------

## Practical Implications

-   Training programs design implications
-   Future research considerations

------------------------------------------------------------------------

## Limitations and Future Directions

-   Study limitations
-   Future research suggestions

------------------------------------------------------------------------

## Conclusion

-   Recap of key findings
-   Final thoughts on training variability's role in learning

------------------------------------------------------------------------

## Questions and Discussion

-   Open floor for questions from the committee and audience

## Future Research Directions {.smaller .r-fit-text}

::::::::: columns
::::: {.column width="50%"}
::: fragment
### Theoretical Extensions

-   Investigate the role of individual differences in training variability effects.
-   Explore the impact of different types of feedback on learning and generalization.
-   Examine the influence of task complexity on the benefits of varied training.
:::

::: fragment
### Methodological Innovations

-   Develop new experimental paradigms to test training variability in more complex tasks.
-   Utilize advanced computational models to simulate learning processes under varied conditions.
-   Implement longitudinal studies to assess long-term effects of training variability.
:::
:::::

::::: {.column width="50%"}
::: fragment
### Practical Applications

-   Design adaptive training programs that adjust variability based on learner performance.
-   Apply findings to educational settings to enhance learning outcomes.
-   Use insights to improve training protocols in professional skill development.
:::

::: fragment
### Collaborative Opportunities

-   Partner with educational institutions to test training variability in classroom settings.
-   Collaborate with industry professionals to apply research findings in real-world training programs.
-   Engage with interdisciplinary teams to explore the broader implications of training variability.
:::
:::::
:::::::::

------------------------------------------------------------------------