### 2024-05-28 Presentation/_hold_slides.qmd
fix slide image displays

---
title: My Slide Show
format:
  revealjs: default
---

## Project 1 - Design  {.nostretch .smaller}

:::: {.columns}
::: {.column width="40%"}

### Experiment 1222
- Train varied group from 2 positions, and a constant group from 1.
- Test both groups from the trained positions - and 1 novel position

![](assets/igas_e1_manip.png){.absolute bottom=30 width="50%"}

:::

::: {.column width="60%"}
### Experiment 2

- 6 distinct constant conditions

![](assets/igas_e2_manip.png){.absolute bottom=0 width="90%"}
:::

::::

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="60%"}
### Solution Space
![](assets/sol_space2.png)

### Full Space
:::{#fig-fullspace23}
![](assets/full_space2.png)

Full Space of Solution Throws and Missed Throws. Colored points represent coordinates of target hits, brown points are coordinates of misses
:::

:::

::: {.column width="40%"}

### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$
:::
::::