### 2024-05-28 Presentation/slides2.qmd

---
title: "The Role of Variability in Learning Generalization: A Computational Modeling Approach"
page-layout: full
aliases:
  - /slides.html
css: /Assets/Style/slides.css
format: 
  # html: 
  #   subtitle: "Dissertation Defense"
  # clean-revealjs:
  #   #theme: moon #white
  #   #output-ext: revealjs.html
  #   # template-partials:
  #   #   - assets/title-slide.html
  #   output-file: slides_revealjs
  #   date: 05/28/204
  #   slideNumber: true
  #   hash-type: number # Use slide numbers in the URL hash
  #   center: true
  #   scrollable: true
  #   chalkboard:
  #     theme: whiteboard
  # # pptx:
  #   incremental: false
  # hugo-md:
  #   include: false
  #   html-math-method: mathjax
  #   output-file: slides_hugo.md
  # gfm:
  #   output: false
  #   output-file: slides_gfm.md
  beamer: default
    # incremental: false
    # colortheme: lily 
author:
  - name: Thomas Gorman
    url: https://tegorman13.github.io
    orcid: 0000-0001-5366-5442
    affiliations:
      - name: Indiana University
        department: Psychological & Brain Sciences, Cognitive Science
# params:
#   start_slide: 5
#engine: knitr
---

# Learning Generalization

- Learning is often specific
- Longstanding scientific interest in how to improve generalization or transfer

------------------------------------------------------------------------

## Variability

- Variation during training linked to improved transfer in numerous domains 
- What does "variability mean in the context of learning interventions?"

## Types of Variability

- The number of unique items/problems experienced
- How spread out examples are

## My Dissertation Focus

- Number of unique examples
- Visuomotor skill learning and function learning

- Addressing methodological shortcomings of previous work
- Adapting cognitive models from other domains to account for results

------------------------------------------------------------------------

## Common Experimental Manipulations

- Varied vs. Constant

## Common Empirical Patterns

:::::::::  columns
::: {.column width="55%"}
### Training
- Both training conditions complete the same number of training trials.
- Varied group has worse training performance.

![](assets/train_pattern.png)
:::

::: {.column width="45%"} 
::: {.fragment}
### Testing

- Tested from novel conditions.
- Varied group has better test performance

![](assets/test_pattern.png)
:::
:::
:::::::::

------------------------------------------------------------------------

## But also plenty of contradictory results and complications

- Cases where varied training doesn't benefit generalization
- Cases where more training variation results in worse outcomes
- Cases where the influence of variation interacts with some other factor
    - difficulty
    - prior knowledge
    - Frequency effects, or amount of training/learning before testing

------------------------------------------------------------------------

## Theoretical Frameworks

-   Schema Theory (Schmidt, 1975)
-   Desirable Difficulties Framework (Bjork & Bjork, 2011)
-   Challenge Point Framework (Guadagnoli & Lee, 2004)

::: footer
@schmidtSchemaTheoryDiscrete1975 @bjorkMakingThingsHard2011 @guadagnoliChallengePointFramework2004
:::

<!-- ![](assets/fig-toy-model1-1-1.png) -->

------------------------------------------------------------------------

## Overview of Current Work

### Project 1[^1]

- Visuomotor projectile launching task
- two experiments
- Beneficial effect of variability
- Instance-based similarity model

### Project 2

- Visuomotor extrapolation task
- Three experiments
- Effect of variability is null or negative
- Connectionist model (ALM) and hybrid associative & rule model (EXAM)

--

# Project 1  - Issues with previous research

- Assumptions about what is encoded
- Assumptions about the formation of abstractions 
- Aggregation issues and similarity confounds

---

## designs that seemingly overcome these issues

- Kerr & Booth 1978

## Experiment 1

- Conceptual replication of Kerr & Booth design 
- Testing classic pattern as well

------------------------------------------------------------------------

## Hit The Target Task

:::::: {column-page-right}
::::: {.columns}
:::: {.column width="60%"}

::: {.content-visible when-format="html"}
![](assets/htt_vid.mov){height="540" loop="true" autoplay="true" style="max-width: 100%"}
:::

::: {.content-visible unless-format="html"}
![](assets/methodsFig1.png)
:::

::::

:::: {.column width="40%"}
- **Training Stage** - 200 training trials with feedback. Constant groups trains from single position. Varied group practices from two positions. 
- **Transfer Stage** - All subjects tested from both positions they were trained, and the positions trained by other group
- **Data recorded** - For every throw, recorded the X velocity and Y velocity of ball at release
::::

:::::
::::::

------------------------------------------------------------------------

------------------------------------------------------------------------

## Project 1 - Experiment 1 Results {.nostretch .smaller}

:::: {.columns}
::: {.column width="40%"}
### Experiment 1
![](assets/igas_e1_manip.png)
:::
::: {.column width="60%"}
::: fragment
### Training
![](assets/htt_e1_train.png)
:::
::: fragment
### Testing
![](assets/fig-IGAS_Testing1-1-2.png)
:::
:::
::::

## Project 1 - Experiment 2 Results

:::: {.columns}
::: {.column width="40%"}
### Experiment 1
![](assets/igas_e2_manip.png)
:::
::: {.column width="60%"}
::: fragment
### Training
![](assets/htt_e2_train.png)
:::
::: fragment
### Testing
![](assets/htt_e2_test.png)
:::
:::
::::

------------------------------------------------------------------------

## Project 1 - Discussion

-   Varied training group's superior performance
-   Computational modeling results with IGAS
-   Theoretical implications

------------------------------------------------------------------------

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="50%"}

![](assets/sol_space2.png){width="65%"}
![](assets/full_space2.png){width="50%"}
:::

::: {.column width="50%"}

### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$

:::
::::

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="70%"}

![](assets/sol_space.png)

![](assets/full_space.png)
:::
::: {.column width="30%"}
### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$
:::
::::

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="60%"}

### Solution Space
![](assets/sol_space2.png)

### Full Space

![](assets/full_space2.png)

:::

::: {.column width="40%"}

### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$
:::
::::

## Computational Model - Does Similarity Account for Effect of Variability?

- Test whether similarity explains the difference in performance between Constant and Varied conditions
  - Assuming equivalent generalization (1 c for both groups) - similarity improves model fit, but does not explain group difference
  - Assume training condition influences generalization gradient (fit c separately) - now similarity does explain the difference between Constant and Varied.

::: footer
**Equation 1:** $$ Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}}) $$
**Equation 2:** $$ d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 } $$
:::

## Project 1 General Discussion {.smaller .r-fit-text}

------------------------------------------------------------------------

# Project 2 - Variability and Extrapolation in a Function Learning Task

-   Influence of varied practice in a function learning task
-   Experiments 1, 2, and 3:
    -   Training regimes and testing conditions
    -   Learning, discrimination, and extrapolation performance metrics

------------------------------------------------------------------------

## Project 2 - Questions and Goals

**Empirical**
- Design a task-space large enough to assess multiple degrees of extrapolation 
- Compare varied and constant generalization from several distinct distances from their nearest training condition

**Model-based**
- If variation does influence extrapolation, can an associative learning model with similarity-based activation provide a good account?
- Can our modelling framework simultaneously account for both training and testing data?
- Accounting for the full distribution of responses

------------------------------------------------------------------------

## Hit The Wall Task

<!-- {{< video assets/htw_vid.mov width="720" height="540" >}} -->

::: {.content-visible when-format="html"}
![](assets/htw_vid.mov){height="400" loop="true" autoplay="true"}
:::

::: {.content-visible unless-format="html"}
![](assets/htw_task_fig.png)
:::

- Target velocity presented at top of screen
- Participants attempt to "hit the wall" at the correct velocity
- Feedback during training - how many units above or below the target-band
- Only the "x velocity" component of the throw is relevant to the task

------------------------------------------------------------------------

## Project 2 - Experiment 1 Design {.nostretch .smaller}

:::: {.columns}
::: {.column width="60%"}
![](assets/htw_e1_manip.png)

![](assets/e1_design.png)
:::
::: {.column width="40%"}

- 156 participants included in final analysis
- Varied group trains from 3 "velocity bands", constant group from 1
:::
::::

------------------------------------------------------------------------

## Project 2 - Experiment 1 Results {.nostretch}

::: {.columns}
::: {.column width="50%"}
![Plot 1](assets/htw_e1_test1.png)
![Plot 3](assets/htw_e1_test3.png)
:::

::: {.column width="50%"}
![Plot 2](assets/htw_e1_test2.png)
![Plot 4](assets/htw_e1_test4.png)
:::
:::

## Project 2 - Experiment 2 Design {.smaller}

:::: {.columns}
::: {.column width="60%"}
![](assets/htw_e2_manip.png)

![](assets/e2_design.png)
:::
::: {.column width="40%"}

<br><br><br><br>

- Training and Testing bands are in reversed order, relative to Experiment 1
- 110 participants included in final analysis

:::
::::

## Project 2 - Experiment 2 Results {.nostretch}

::: {.columns}
::: {.column width="50%"}
![Plot 1](assets/htw_e2_test1.png)
![Plot 3](assets/htw_e2_test3.png)
:::

::: {.column width="50%"}
![Plot 2](assets/htw_e2_test2.png)
![Plot 4](assets/htw_e2_test4.png)
:::
:::

## Project 2 - Experiment 3

- Ordinal Feedback
  - directional feedback indicating too high, too low, or correct
- Both Original (Experiment 1), and reverse (Experiment 2) orders included
- 195 participants included in final analysis

## Project 2 - Experiment 3 Results {.nostretch}

## Accuracy

:::: {.columns}

::: {.column width="50%"}
![Plot 1](assets/htw_e3_test1.png)
:::

::: {.column width="50%"}
![Plot 3](assets/htw_e3_test2.png)
:::

::::

## Discrimination

:::: {.columns}

::: {.column width="50%"}
![Plot 1](assets/htw_e3_test3.png)
:::

::: {.column width="50%"}
![Plot 3](assets/htw_e3_test4.png)
:::

::::

------------------------------------------------------------------------

# Project 2 - Computational Modeling

## Project 2 - ALM

![](assets/fig-alm-diagram-1-1.png)

------------------------------------------------------------------------

## ALM

|                    | **ALM Response Generation**                                        |                                                                                               |
|------------------|-----------------------------|-------------------------|
| Input Activation   | $a_i(X) = \frac{e^{-c(X-X_i)^2}}{\sum_{k=1}^M e^{-c(X-X_k)^2}}$    | Input nodes activate as a function of Gaussian similarity to stimulus                         |
| Output Activation  | $O_j(X) = \sum_{k=1}^M w_{ji} \cdot a_i(X)$                        | Output unit $O_j$ activation is the weighted sum of input activations and association weights |
| Output Probability | $P[Y_j|X] = \frac{O_j(X)}{\sum_{k=1}^M O_k(X)}$                    | The response, $Y_j$ probabilites computed via Luce's choice rule                              |
| Mean Output        | $m(X) = \sum_{j=1}^L Y_j \cdot \frac{O_j(x)}{\sum_{k=1}^M O_k(X)}$ | Weighted average of probabilities determines response to X                                    |
|                    | **ALM Learning**                                                   |                                                                                               |
| Feedback           | $f_j(Z) = e^{-c(Z-Y_j)^2}$                                         | feedback signal Z computed as similarity between ideal response and observed response         |
| magnitude of error | $\Delta_{ji}=(f_{j}(Z)-o_{j}(X))a_{i}(X)$                          | Delta rule to update weights.                                                                 |
| Update Weights     | $w_{ji}^{new}=w_{ji}+\eta\Delta_{ji}$                              | Updates scaled by learning rate parameter $\eta$.                                             |

------------------------------------------------------------------------

## EXAM

|                    | **EXAM Response Generation**                                                                                     |                                                                                                               |
|--------------------|-----------------------------------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------|
| Instance Retrieval | $P[X_i|X] = \frac{a_i(X)}{\sum_{k=1}^M a_k(X)}$                                                                 | Novel test stimulus $X$ activates input nodes $X_i$                                                           |
| Slope Computation  | $S = \frac{m(X_{1})-m(X_{2})}{X_{1}-X_{2}}$                                                                     | Slope value, $S$ computed from nearest training instances                                                     |
| Response           | $E[Y|X_i] = m(X_i) + S \cdot [X - X_i]$                                                                         | Final EXAM response is the ALM response for the nearest training stimulus, $m(X_i)$, adjusted by local slope $S$. |

------------------------------------------------------------------------

## Project 2 - Model Fitting Procedure {.r-fit-text}

### Approximate Bayesian Computation

:::: {.columns}

::: {.column width="60%"}
![](assets/abc_diagram1.png)
:::

::: {.column width="40%"}

- simulation based approach
- approximate likelihood
- uncertainty in parameter values
- full distribution of plausible model predictions for each participant

:::

::::

----

## Modelling Results

![](assets/fig-ee-e1-1.png)

# General Discussion

-   Compare HTT and HTW
-   Empirical findings summary
-   Computational modeling contributions

------------------------------------------------------------------------

## Comparison {.smaller .r-fit-text}

| Dimension | HTT (Project 1) | HTW (Project 2) |
|----------------|--------------------------------|-------------------------|
| Task Description | Projectile launching to hit a target | Projectile launching to hit wall at a specific velocity |
| Task Complexity | More complex parabolic trajectory, both x and y velocities relevant to outcome | Simpler 1D mapping of force to outcome. Only x velocity is relevant. |
| Task Space | More complex: xy velocity combinations closer to the solution space may still result in worse feedback due to striking the barrier. | Simpler: smooth, linear mapping between velocity and feedback. |
| Perceptual salience of Varied Conditions | Varied conditions (# of throwing distances) are perceptually distinct, i.e. salient differences in distance between launching box and target. | Varied conditions (# of velocity bands) are less salient - only difference is the numeral displayed on screen. |
| Testing Feedback | Testing always included feedback | Primary testing stage had no feedback. |
| Potential for Learning during Testing | Limited potential for learning during testing due to feedback. | Some potential for learning during no-feedback testing by observing ball trajectory. |
| Training Experience | Varied group gets half as much experience on any one position as the constant group. | Varied group gets 1/3 as much experience on any one velocity band as the constant group. |
| Testing Structure | Random interleaving of trained/transfer testing distances. | Blocked structure, separately testing trained vs extrapolation testing bands. |

------------------------------------------------------------------------

## Conclusion

## References
::: {#refs}
:::