### 2024-05-28 Presentation/_hold_slides.qmd

---
title: My Slide Show
format:
  revealjs: default
---

## designs that seemingly overcome these issues

- Kerr & Booth 1978

::: footer
@kerrSpecificVariedPractice1978
:::

## Project 1 - Design  {.nostretch .smaller}

:::: {.columns}
::: {.column width="40%"}

### Experiment 1222
- Train varied group from 2 positions, and a constant group from 1.
- Test both groups from the trained positions - and 1 novel position

![](assets/igas_e1_manip.png){.absolute bottom=30 width="50%"}

:::

::: {.column width="60%"}
### Experiment 2

- 6 distinct constant conditions

![](assets/igas_e2_manip.png){.absolute bottom=0 width="90%"}
:::

::::

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="60%"}
### Solution Space
![](assets/sol_space2.png)

### Full Space
:::{#fig-fullspace23}
![](assets/full_space2.png)

Full Space of Solution Throws and Missed Throws. Colored points represent coordinates of target hits, brown points are coordinates of misses
:::

:::

::: {.column width="40%"}

### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$
:::
::::

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="70%"}

![](assets/sol_space.png)

![](assets/full_space.png)
:::
::: {.column width="30%"}
### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$
:::
::::

## Computational Model - Does Similarity Account for Effect of Variability?

- Test whether similarity explains the difference in performance between Constant and Varied conditions
  - Assuming equivalent generalization (1 c for both groups) - similarity improves model fit, but does not explain group difference
  - Assume training condition influences generalization gradient (fit c separately) - now similarity does explain the difference between Constant and Varied.

::: footer
**Equation 1:** $$ Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}}) $$
**Equation 2:** $$ d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 } $$
:::

## Project 1 General Discussion {.smaller .r-fit-text}

------------------------------------------------------------------------