### 2024-05-28 Presentation/slides.qmd
update slides

---
title: "The Role of Variability in Learning Generalization: A Computational Modeling Approach"
page-layout: full
aliases:
  - /slides.html
css: /Assets/Style/slides.css
title-slide-attributes:
  data-background-image: /Presentation/assets/sol_blank.png
  data-background-size: 300px #contain
  data-background-opacity: "0.15"
  data-background-position: "0% 0%"
format-links: [clean-revealjs, pptx,gfm]
format: 
  html: 
    subtitle: "Dissertation Defense"
  clean-revealjs:
    #theme: moon #white
    #output-ext: revealjs.html
    # template-partials:
    #   - assets/title-slide.html
    output-file: slides_revealjs
    date: 05/28/2024
    slideNumber: true
    hash-type: number # Use slide numbers in the URL hash
    center: true
    scrollable: true
    chalkboard:
      theme: whiteboard
  pptx:
    incremental: false
  hugo-md:
    include: false
    html-math-method: mathjax
    output-file: slides_hugo.md
  gfm:
    output: false
    output-file: slides_gfm.md
  # beamer:
  #   incremental: false
  #   colortheme: lily 
author:
  - name: Thomas Gorman
    url: https://tegorman13.github.io
    orcid: 0000-0001-5366-5442
    affiliations:
      - name: Indiana University
        department: Psychological & Brain Sciences, Cognitive Science
# params:
#   start_slide: 
#engine: knitr
---

{{< include /Presentation/embed_slides.qmd >}}

# Learning Generalization

- Learning is often specific
- Longstanding scientific interest in how to improve generalization or transfer

------------------------------------------------------------------------

## Variability

- Variation during training linked to improved transfer in numerous domains 
- What does "variability mean in the context of learning interventions?"

## Types of Variability

- The number of unique items/problems experienced
- How spread out examples are

## My Dissertation Focus

- Number of unique examples
- Visuomotor skill learning and function learning

- Addressing methodological shortcomings of previous work
- Adapting cognitive models from other domains to account for results

------------------------------------------------------------------------

## Common Experimental Manipulations

- Varied vs. Constant

## Common Empirical Patterns

:::::::::  columns
::: {.column width="55%"}
### Training
- Both training conditions complete the same number of training trials.
- Varied group has worse training performance.

![](assets/train_pattern.png)
:::

::: {.column width="45%"} 
::: {.fragment}
### Testing

- Tested from novel conditions.
- Varied group has better test performance

![](assets/test_pattern.png)
:::
:::
:::::::::

------------------------------------------------------------------------

## But also plenty of contradictory results and complications

- Cases where varied training doesn't benefit generalization
- Cases where more training variation results in worse outcomes
- Cases where the influence of variation interacts with some other factor
    - difficulty
    - prior knowledge
    - Frequency effects, or amount of training/learning before testing

------------------------------------------------------------------------

## Overview {.smaller}

### Project 1

- Visuomotor projectile launching task
- two experiments
- Beneficial effect of variability
- Instance-based similarity model

### Project 2

- Visuomotor extrapolation task
- Three experiments
- Effect of variability is null or negative
- Connectionist model (ALM) and hybrid associative & rule model (EXAM)

------------------------------------------------------------------------

---

## Project 1 
<br>

### An instance-based model account of the benefits of varied practice in visuomotor skill[^1]

<br><br><br>

::: footer
[^1]: Now published in Cognitive Psychology -  [@gormanInstancebasedModelAccount2022]
:::

---

## Theoretical Frameworks

<br>

::: {.fragment .semi-fade-out}
- Desirable Difficulties Framework (Bjork & Bjork, 2011)
<br>  
- Challenge Point Framework (Guadagnoli & Lee, 2004)
::: 
- Schema Theory (Schmidt, 1975)

::: footer
@schmidtSchemaTheoryDiscrete1975 @bjorkMakingThingsHard2011 @guadagnoliChallengePointFramework2004
:::

::: notes
Variety of theoretical explanations
Schema theory in paticular has been extremely influential 
:::

___

## Issues with previous research {.smaller}

- Assumptions about what is encoded
- Assumptions about the formation of abstractions 
- Aggregation issues and similarity confounds

::: fragment
:::: columns
::: {.column}

![](assets/agg_issues.png)
:::
::: {.column}
<br>

-  If generalization decays exponentially with distance, constant groups 1 and 2 may have far less generalization potential to the testing condition. 
- performance of individual constant groups often not reported in analyses.

:::

::::
:::
::: footnote
e.g. [@mccrackenTestSchemaTheory1977; @chuaPracticeVariabilityPromotes2019; @delreyEffectsContextualInterference1982]
:::

---

## Designs that avoid similarity issue

### Kerr & Booth 1978

:::: columns
::: {.column}
![](assets/kerr_booth.png)
:::
::: {.column}
- Bean bag tossing task
- Constant and varied conditions train from distinct positions
- Both groups are tested from the position where the constant group trained
- Impressive demonstration of varied training outperforming constant training 
:::
::::
::: footer
@kerrSpecificVariedPractice1978
:::

---

## Experiment 1

- Conceptual replication of Kerr & Booth design 
- Also test positions novel to both groups
- Validate computerized visuomotor learning task

## Experiment 1 Design {.smaller}

![](assets/igas_e1_manip.png){fig-align="left" fig-align="top"}

- Constant trains from one position (760)
  - 200 trials
- Varied trains from two positions (610 and 910)
  - 100 trials per position
- Both groups are tested from all three training positions, and a new position novel to both groups (835)

------------------------------------------------------------------------

## Hit The Target Task {.smaller}

:::::: column-page-right
::::: {.columns}
:::: {.column width="60%"}

::: {.content-visible when-format="html"}
![](assets/htt_vid.mov){height="540" loop="true" autoplay="true" style="max-width: 100%"}
:::

::: {.content-visible unless-format="html"}
![](assets/methodsFig1.png)
:::

::::

:::: {.column width="40%"}
- **Training Stage** - 200 training trials with feedback. Constant groups trains from single position. Varied group practices from two positions. 
- **Transfer Stage** - All subjects tested from both positions they were trained, and the positions trained by other group
- **Data recorded** - For every throw, recorded the X velocity and Y velocity of ball at release
::::

:::::
::::::

::: notes
- Use mouse to flick the ball towards the target
- Must release from orange square
- Subjects have flexibility in terms of trajectory and release point inside the square
- Nicety of computerized task - we record the precise details easily
:::

------------------------------------------------------------------------

------------------------------------------------------------------------

## Project 1 - Experiment 1 Results {.smaller}

:::: {.columns}
::: {.column width="50%"}
::: fragment
#### Training
![](assets/htt_e1_train.png){width="95%"}
:::
:::
::: {.column width="50%"}
::: fragment
#### Testing
![](assets/fig-IGAS_Testing1-1-2.png){width="95%"}
:::
:::
::::

::: notes
Varied training significantly better at the position that was novel for all subjects. Also better even at the position that the constant group trained at.

4 different positions, the varied group trained from the easiest and the hardest position, they also interpolated the constant group. 
:::

___

## Experiment 2 Design

![](assets/igas_e2_manip.png)

::: notes
task mostly the same - minor adjustments to barrier
6 different constant groups
varied group doesn't train on either the easiest or the hardest position

larger design to give us more data to work with for modelling
:::

## Project 1 - Experiment 2 Results

:::: {.columns}
::: {.column width="50%"}
::: fragment
#### Training {.smaller}
![](assets/htt_e2_train.png){width="95%"}
:::
:::
::: {.column width="50%"}

::: fragment
#### Testing {.smaller}
![](assets/htt_e2_test.png){width="95%"}
:::
:::
::::

::: notes

- This time we can compare varied and constant training - by just comparing the varied group to
the 2 constant groups that trained at the same locations
- Here we have all 6 of the constant groups collapsed together
- also did more fine grained grained comparisons

:::

------------------------------------------------------------------------

## Project 1 Computational Model {.smaller .r-fit-text}

### Instance encoding, and similarity-based generalization assumptions

- Instances refers to individual throws (x and y velocity)
- Similarity refers to the distance between training throws and the solution space of each of the eventual testing positions.

___

## Project 1 Computational Model {.smaller}

:::: {.columns}
::: {.column width="50%"}

![](assets/sol_space2.png){width="65%"}
![](assets/full_space2.png){width="50%"}
:::

::: {.column width="50%"}

#### Computing Similarity

- Euclidean distance between each training throw, and each solution space
  - separately for each participant, and each of the testing positions
- Similarity computed as a Gaussian decay function of distance, i.e. larger distances result in lower similarity
- Each participant gets their own similarity score for each of the 6 testing positions
  - similarity score for a given testing position is the sum of the similarities between each training throw and the entire empirical solution space

#### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$

:::
::::

## Non-linear similarity

![](assets/c_example.png)

___

## Does this similarity metric work?

- Our similarity measure does is a significant predictor of testing performance. 
  - Remains significant when controlling for training accuracy, and training condition.

___

## Accounting for group-level effect {.smaller}

::: fragment
- The base version of the model fit a single $c$ parameter 
  - thus assuming that everyone generalizes to the same degree
:::

::: fragment
- We next fit a 2-c version, separately optimizing for varied and constant groups, resulting in new 2-c similarity scores for each participant. 
  - The optimal c value for the varied group is smaller than that of the constant group, indicative of broader generalization. 
:::

::: fragment
- When 2-c similarity is added to the linear model predicting testing performance as a function of condition, the effect of training condition is no longer significant. 
  - not the case with the original 1-c similarities. 
:::

::: fragment  
- Thus the influence of varied training can be explained by an instance-based similarity model, IF one assumes flexibility in the generalization gradient. 
:::

## Conclusions

::: fragment 
- New empirical support for a benefit of variability in visuomotor skill tasks
:::

::: fragment 
- Some support for the less common pattern observed by @kerrSpecificVariedPractice1978
:::

::: fragment 
- Demonstrated the utility of a similarity based approach in this domain
:::

---

# Project 2

---

## Project 2

### Impact of Training Variability on Visuomotor Function Learning and Extrapolation

-   Influence of varied practice in a function learning task
-   Experiments 1, 2, and 3:
    -   Training regimes and testing conditions
    -   Learning, discrimination, and extrapolation performance metrics

------------------------------------------------------------------------

## Project 2 - Questions and Goals

**Empirical**
- Design a task-space large enough to assess multiple degrees of extrapolation 
- Compare varied and constant generalization from several distinct distances from their nearest training condition

**Model-based**
- If variation does influence extrapolation, can an associative learning model with similarity-based activation provide a good account?
- Can our modelling framework simultaneously account for both training and testing data?
- Accounting for the full distribution of responses

------------------------------------------------------------------------

## Hit The Wall Task {.smaller}

<!-- {{< video assets/htw_vid.mov width="720" height="540" >}} -->

::: {.content-visible when-format="html"}
![](assets/htw_vid.mov){height="400" loop="true" autoplay="true"}
:::

::: {.content-visible unless-format="html"}
![](assets/htw_task_fig.png)
:::

- Target velocity presented at top of screen
- Participants attempt to "hit the wall" at the correct velocity
- Feedback during training - how many units above or below the target-band
- Only the "x velocity" component of the throw is relevant to the task

------------------------------------------------------------------------

## Project 2 - Experiment 1 Design {.nostretch .smaller}

:::: {.columns}
::: {.column width="60%"}
![](assets/htw_e1_manip.png){fig-align="left" fig-align="top"}

![](assets/e1_design.png){fig-align="left" fig-align="top"}
:::
::: {.column width="40%"}

<br><br><br>

- 156 participants included in final analysis
- Varied group trains from 3 "velocity bands", constant group from 1
- Both groups complete same total number of training trials
:::
::::

------------------------------------------------------------------------

## Project 2 - Experiment 1 Results

::: {.columns}
::: {.column width="50%"}
![](assets/htw_e1_test1.png){width="70%"}
![](assets/htw_e1_test3.png){width="70%"}
:::

::: {.column width="50%"}
![](assets/htw_e1_test2.png){width="70%"}
![](assets/htw_e1_test4.png){width="70%"}
:::
:::

## Project 2 - Experiment 2 Design {.smaller}

:::: {.columns}
::: {.column width="60%"}
![](assets/htw_e2_manip.png){fig-align="left" fig-align="top"}

![](assets/e2_design.png){fig-align="left" fig-align="top"}
:::
::: {.column width="40%"}

<br><br><br><br>

- Training and Testing bands are in reversed order, relative to Experiment 1
- 110 participants included in final analysis

:::
::::

## Project 2 - Experiment 2 Results

::: {.columns}
::: {.column width="50%"}
![](assets/htw_e2_test1.png){width="75%" .fragment fragment-index=1}
![](assets/htw_e2_test3.png){width="75%" .fragment fragment-index=2}
:::

::: {.column width="50%"}
![](assets/htw_e2_test2.png){width="75%" .fragment fragment-index=1}
![](assets/htw_e2_test4.png){width="75%" .fragment fragment-index=2}
:::
:::

## Project 2 - Experiment 3

- Ordinal Feedback
  - directional feedback indicating too high, too low, or correct
- Both Original (Experiment 1), and reverse (Experiment 2) orders included
- 195 participants included in final analysis

## Project 2 - Experiment 3 Results {.nostretch}

## Experiment 3 - Accuracy

:::: {.columns}

::: {.column width="50%"}
![](assets/htw_e3_test1.png){width="80%"}
:::

::: {.column width="50%"}
![](assets/htw_e3_test2.png){width="80%"}
:::

::::

## Experiment 3 - Discrimination

:::: {.columns}

::: {.column width="50%"}
![](assets/htw_e3_test3.png){width="80%"}
:::

::: {.column width="50%"}
![](assets/htw_e3_test4.png){width="80%"}
:::

::::

------------------------------------------------------------------------

# Project 2 - Computational Modeling

## The Associative Learning Model ALM {.smaller}

:::: {.columns}

::: {.column width="70%"}
![](assets/fig-alm-diagram-1-1.png)
:::
::: {.column width="30%"}
- Two layer network - adapted from ALCOVE (@kruschkeALCOVEExemplarbasedConnectionist1992)
- Input layer node for each stimulus, output node for each response 
- Input nodes activate as a function of their Gaussian similarity to the stimulus
- Weights udpated via delta rule - error driven learning
- Provides good account of human learning data, and interpolation performance, but struggles with extrapolation [@deloshExtrapolationSineQua1997]
:::
::::
------------------------------------------------------------------------

## Extrapolation - Association Model (EXAM) {.smaller}

- Extension to ALM to account for human extrapolation behavior 
- When a novel stimulus is presented, EXAM assumes the nearest 2-3 prior examples are retrieved, and used to compute a slope
- The slope is used to adjust the ALM response

|                    | **EXAM Response Generation**                                                                                     |                                                                                                               |
|--------------------|-----------------------------------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------|
| Slope Computation  | $S = \frac{m(X_{1})-m(X_{2})}{X_{1}-X_{2}}$                                                                     | Slope value, $S$ computed from nearest training instances                                                     |
| Response           | $E[Y|X_i] = m(X_i) + S \cdot [X - X_i]$                                                                         | Final EXAM response is the ALM response for the nearest training stimulus, $m(X_i)$, adjusted by local slope $S$. |

------------------------------------------------------------------------

## Project 2 - Model Fitting Procedure {.smaller}

### Approximate Bayesian Computation (ABC)

:::: {.columns}

::: {.column width="60%"}
![](assets/abc_diagram1.png){width="80%"}
:::

::: {.column width="40%"}

- simulation based parameter estimation [@turnerTutorialApproximateBayesian2012; @kangasraasioParameterInferenceComputational2019]
- Useful for models with unknown likelihood functions (e.g. many neural network and drift diffusion models)
- full distribution of plausible model predictions for each participant

:::

::::

## Project 2 - Model Fitting Procedure {.smaller}

### Model Fitting Approach

- Two parameters for both ALM and EXAM
  - generalization parameter: $c$
  - learning rate parameter: $lr$

- ALM and EXAM fit separately to each individual participant
- Rejection based ABC used to obtain 200 posterior samples, per participant, per model 
  - i.e. 200 plausible values of $c$ and $lr$

- All models fit with three different approaches
    1) Fit to only the training data
    2) Fit to both training and testing data
    3) Fit to only testing data

---

## Modelling Results {.smaller}

#### Posterior Predictive Distribution

![](assets/post_pred_e1.png)

::: notes
 Empirical data and Model predictions for mean velocity across target bands. Fitting methods (Test Only, Test & Train, Train Only) - are separated across rows, and Training Condition (Constant vs. Varied) are separated by columns. Each facet contains the predictions of ALM and EXAM, alongside the observed data.
:::

## Model Comparison - Experiment 1

![](assets/model_reg_e1.png)

::: notes
Figure 21: A-C) Conditional effects of Model (ALM vs EXAM) and Condition (Constant vs. Varied). Lower values on the y axis indicate better model fit. D) Specific contrasts of model performance comparing 1) EXAM fits between constant and varied training; 2) ALM vs. EXAM for the varied group; 3) ALM fits between constant and varied. Negative error differences indicate that the term on the left side (e.g., EXAM Constant) tended to have smaller model residuals.
:::

## Model Comparison - Experiment 2 and 3

![](assets/fig-e2_e3_ae-1.png)

::: notes
Conditional effects of Model (ALM vs EXAM) and Condition (Constant vs. Varied) on Model Error for Experiment 2 and 3 data. Experiment 3 also includes a control for the order of training vs. testing bands (original order vs. reverse order).
:::

___

## Modelling Results {.smaller}

### Best fitting models per participant

![](assets/best_models.png)

::: notes
Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM
:::

___

# General Discussion

-   Compare HTT and HTW
-   Empirical findings summary
-   Computational modeling contributions

------------------------------------------------------------------------

## Comparison {.smaller .r-fit-text}

| Dimension | HTT (Project 1) | HTW (Project 2) |
|----------------|--------------------------------|-------------------------|
| Task Description | Projectile launching to hit a target | Projectile launching to hit wall at a specific velocity |
| Task Complexity | More complex parabolic trajectory, both x and y velocities relevant to outcome | Simpler 1D mapping of force to outcome. Only x velocity is relevant. |
| Task Space | More complex: xy velocity combinations closer to the solution space may still result in worse feedback due to striking the barrier. | Simpler: smooth, linear mapping between velocity and feedback. |
| Perceptual salience of Varied Conditions | Varied conditions (# of throwing distances) are perceptually distinct, i.e. salient differences in distance between launching box and target. | Varied conditions (# of velocity bands) are less salient - only difference is the numeral displayed on screen. |
| Testing Feedback | Testing always included feedback | Primary testing stage had no feedback. |
| Potential for Learning during Testing | Limited potential for learning during testing due to feedback. | Some potential for learning during no-feedback testing by observing ball trajectory. |
| Training Experience | Varied group gets half as much experience on any one position as the constant group. | Varied group gets 1/3 as much experience on any one velocity band as the constant group. |
| Testing Structure | Random interleaving of trained/transfer testing distances. | Blocked structure, separately testing trained vs extrapolation testing bands. |

------------------------------------------------------------------------

## Conclusion

## References
::: {#refs}
:::

___

## Extras

___

::: {.content-visible when-format="html"}
## Manuscript {.scrollable .unnumbered .unlisted}

```{=html}
<iframe width="800" height="750" src="https://tegorman13.github.io/Dissertation/Manuscript/output/manuscript.html" title="Webpage example"></iframe>
```
:::

------------------------------------------------------------------------

## ALM {.scrollable .unnumbered .unlisted}

|                    | **ALM Response Generation**                                        |                                                                                               |
|------------------|-----------------------------|-------------------------|
| Input Activation   | $a_i(X) = \frac{e^{-c(X-X_i)^2}}{\sum_{k=1}^M e^{-c(X-X_k)^2}}$    | Input nodes activate as a function of Gaussian similarity to stimulus                         |
| Output Activation  | $O_j(X) = \sum_{k=1}^M w_{ji} \cdot a_i(X)$                        | Output unit $O_j$ activation is the weighted sum of input activations and association weights |
| Output Probability | $P[Y_j|X] = \frac{O_j(X)}{\sum_{k=1}^M O_k(X)}$                    | The response, $Y_j$ probabilites computed via Luce's choice rule                              |
| Mean Output        | $m(X) = \sum_{j=1}^L Y_j \cdot \frac{O_j(x)}{\sum_{k=1}^M O_k(X)}$ | Weighted average of probabilities determines response to X                                    |
|                    | **ALM Learning**                                                   |                                                                                               |
| Feedback           | $f_j(Z) = e^{-c(Z-Y_j)^2}$                                         | feedback signal Z computed as similarity between ideal response and observed response         |
| magnitude of error | $\Delta_{ji}=(f_{j}(Z)-o_{j}(X))a_{i}(X)$                          | Delta rule to update weights.                                                                 |
| Update Weights     | $w_{ji}^{new}=w_{ji}+\eta\Delta_{ji}$                              | Updates scaled by learning rate parameter $\eta$.                                             |

------------------------------------------------------------------------

## EXAM {.scrollable .unnumbered .unlisted}

|                    | **EXAM Response Generation**                                                                                     |                                                                                                               |
|--------------------|-----------------------------------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------|
| Instance Retrieval | $P[X_i|X] = \frac{a_i(X)}{\sum_{k=1}^M a_k(X)}$                                                                 | Novel test stimulus $X$ activates input nodes $X_i$                                                           |
| Slope Computation  | $S = \frac{m(X_{1})-m(X_{2})}{X_{1}-X_{2}}$                                                                     | Slope value, $S$ computed from nearest training instances                                                     |
| Response           | $E[Y|X_i] = m(X_i) + S \cdot [X - X_i]$                                                                         | Final EXAM response is the ALM response for the nearest training stimulus, $m(X_i)$, adjusted by local slope $S$. |

------------------------------------------------------------------------