### 2024-05-28 Presentation/_hold_slides.qmd
update slides again

---
title: My Slide Show
format:
  revealjs: default
---

## designs that seemingly overcome these issues

- Kerr & Booth 1978

::: footer
@kerrSpecificVariedPractice1978
:::

## Project 1 - Design  {.nostretch .smaller}

:::: {.columns}
::: {.column width="40%"}

### Experiment 1222
- Train varied group from 2 positions, and a constant group from 1.
- Test both groups from the trained positions - and 1 novel position

![](assets/igas_e1_manip.png){.absolute bottom=30 width="50%"}

:::

::: {.column width="60%"}
### Experiment 2

- 6 distinct constant conditions

![](assets/igas_e2_manip.png){.absolute bottom=0 width="90%"}
:::

::::

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="60%"}
### Solution Space
![](assets/sol_space2.png)

### Full Space
:::{#fig-fullspace23}
![](assets/full_space2.png)

Full Space of Solution Throws and Missed Throws. Colored points represent coordinates of target hits, brown points are coordinates of misses
:::

:::

::: {.column width="40%"}

### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$
:::
::::

## Project 1 Computational Model {.smaller .r-fit-text}

:::: {.columns}
::: {.column width="70%"}

![](assets/sol_space.png)

![](assets/full_space.png)
:::
::: {.column width="30%"}
### Model Steps

- each of the 6 positions has an empirical solution space
- Compute the summed similarity between training throws, and solution spaces
  - separately for each participant, and each of the testing positions
- We now have a measure of how similar training behavior was to testing solutions

### Model Definition

- $d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 }$
- $Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}})$
:::
::::

## Computational Model - Does Similarity Account for Effect of Variability?

- Test whether similarity explains the difference in performance between Constant and Varied conditions
  - Assuming equivalent generalization (1 c for both groups) - similarity improves model fit, but does not explain group difference
  - Assume training condition influences generalization gradient (fit c separately) - now similarity does explain the difference between Constant and Varied.

::: footer
**Equation 1:** $$ Similarity_{I,J} = \sum_{i=I}\sum_{j=J} (e^{-c^\cdot d^{p}_{i,j}}) $$
**Equation 2:** $$ d_{i,j} = \sqrt{(x_{Train_i}-x_{Solution_j})^2 + (y_{Train_i}-y_{Solution_j})^2 } $$
:::

## Project 1 General Discussion {.smaller .r-fit-text}

---------------------------------------------------------

---------------

## Project 2 - Experiment 2 Results

::: {.columns}
::: {.column width="50%"}
![](assets/htw_e2_test1.png){width="75%"}
![](assets/htw_e2_test3.png){width="75%"}
:::

::: {.column width="50%"}
![](assets/htw_e2_test2.png){width="75%"}
![](assets/htw_e2_test4.png){width="75%"}
:::
:::

## Project 2 - Experiment 2 Results {.smaller margin="0"}

::: {.columns}
::: {.column width="50%"}
![](assets/htw_e2_test1.png){width=65%"}
:::
::: {.column width="50%"}
![](assets/htw_e2_test2.png){width="65%"}
:::
:::

::: fragment
::: {.columns}
::: {.column width="50%"}

![](assets/htw_e2_test3.png){width="65%"}
:::
::: {.column width="50%"}
![](assets/htw_e2_test4.png){width="65%"}
:::
:::
:::

## Issues with previous research {.smaller}

- Assumptions about what is encoded
- Assumptions about the formation of abstractions 
- Aggregation issues and similarity confounds

::: fragment
:::: columns
::: {.column}

![](assets/agg_issues.png)
:::
::: {.column}
<br>

-  If generalization decays exponentially with distance, constant groups 1 and 2 may have far less generalization potential to the testing condition. 
- performance of individual constant groups often not reported in analyses.

:::

::::
:::
::: footnote
e.g. [@mccrackenTestSchemaTheory1977; @chuaPracticeVariabilityPromotes2019; @delreyEffectsContextualInterference1982]
:::

---