### 2024-06-21 index.qmd
minor changes

---
title: "Dissertation"
author: Thomas E. Gorman
comments: false
date-format: 'MMM YYYY'
format:
  html:
    grid:
      sidebar-width: 250px
      body-width: 1400px
      margin-width: 170px
      gutter-width: 1.5rem
    self-contained: false
---

## Project 1

[{{< fa file-lines >}} Project 1 page](https://tegorman13.github.io/Dissertation/Sections/IGAS.html)

[{{< fa file-pdf >}} pdf of the journal
article](Assets/Gorman_Goldstone_2022_Instance-based_model_varied_practice.pdf){target="_blank"}\
[{{< fa scroll >}} Link to online version of journal
article](https://www.sciencedirect.com/science/article/abs/pii/S0010028522000299){target="_blank"}

## Project 2

[{{< fa file-lines >}} Link to project 2 page](https://tegorman13.github.io/htw/){target="_blank"} \

[{{< fa file-pdf >}} Working Draft of Manuscript](https://tegorman13.github.io/htw/paper.html){target="_blank"}