### 2024-09-07 Manuscript/pre_sections/diss_cv.qmd

\pagenumbering{gobble}

# Curriculum Vitae

## Education {.unnumbered .unlisted}

::: {.content-visible when-format="pdf"}
```{=tex}
\begin{tabular}{p{0.3\textwidth} p{0.7\textwidth}}
2017 - May 2024 & \textbf{Indiana University - Bloomington} \\
& PhD in Psychology and Cognitive Science \\ 
& Dissertation: \textit{The Role of Variability in Learning Generalization: A Computational Modeling Approach}\\
2011 - 2015 & \textbf{University of Wisconsin - Madison} \\
& B.Sc. in Psychology \\
& Thesis: \textit{Short-term mindfulness intervention reduces the negative attentional effects associated with heavy media multitasking} \\
\end{tabular}
```
:::

## Journal Articles and Conference Proceedings {.unnumbered .unlisted}

-   **Gorman**, T. E., & Goldstone, R. L. (2022). An instance-based model account of the benefits of varied practice in visuomotor skill. *Cognitive Psychology*, *137*, 101491.

-   Kattner, F., Cochrane, A., Cox, C. R., **Gorman**, T. E., & Green, C. S. (2017). Perceptual learning generalization from sequential perceptual training as a change in learning rate. *Current Biology*, *27*(6), 840-846.

-   **Gorman**, T.E., & Green, C.S. (2016). Short-term mindfulness intervention reduces the negative attentional effects associated with heavy media multitasking. *Scientific Reports*, *6*.

## Reviews and Book Chapters {.unnumbered .unlisted}

-   **Gorman**, T.E., Gentile, D.A., & Green, C.S. Problem gaming: a short primer (2018). *American Journal of Play*, *10* (3), 309-327

-   **Gorman**, T.E., & Green, C.S. (2017). Young minds on video games. In *Cognitive development in digital contexts*. 121-143. Academic Press.

-   Green, C. S., **Gorman**, T., & Bavelier, D. (2016). Action Video-Game Training and Its Effects on Perception and Attentional Control. In *Cognitive Training* (pp. 107-116). Springer International Publishing.

## Presentations {.unnumbered .unlisted}

-   Half Day Tutorial on Measuring Mindfulness Behaviorally: Onsite/Online Data Col-lection with jsPsych
    -   Cognitive Science Conference – 2018
-   Does interleaving go the distance? Exploring the effect of dissimilarity on interleaved category learning
    -   Math Psych/ICCM 2018
-   Short term mindfulness intervention reduces cognitive deficits in heavy media multi-taskers
    -   Undergraduate Research Symposium – UW-Madison – April 16th, 2015

## Awards {.unnumbered .unlisted}

IU-Bloomington – Development Training Grant – Fall 2019

IU-Bloomington - Steinmetz Summer Research Award - 2018

UW-Madison - Undergraduate Research Scholar Award – 2015

UW-Madison - Hilldale Undergraduate/Faculty Research Fellowship – 2014

## Ad-hoc Reviewer {.unnumbered .unlisted}

Cognitive Science

Journal of Experimental Psychology: General

Journal of Experimental Psychology: Human Perception and Performance

Contemporary Educational Psychology

## Conferences {.unnumbered .unlisted}

Cognitive Science Conference - 2021

Cognitive Science Conference - 2018

Mathematical Psychology & ICCM 2018

Boston Meeting on Methods in Cognitive Training (NSF Sponsored) – May 2017