### 2024-10-11 Manuscript/pre_sections/pre_pdf_full.qmd
cv additions and re-render

\begin{centering}

{THE ROLE OF VARIABILITY IN LEARNING GENERALIZATION: A COMPUTATIONAL MODELING APPROACH}

\vspace{5\baselineskip}

{Thomas E. Gorman}

\vspace{10.5cm}

Submitted to the faculty of the University Graduate School \\
in partial fulfillment of the requirements \\
for the degree \\
Doctor of Philosophy \\
in the Cognitive Science Program and \\
Department of Psychological and Brain Sciences, \\
Indiana University \\
September 2024

\vspace{6cm}

\pagenumbering{gobble}
\end{centering}

\newpage
\pagenumbering{roman}
\setcounter{page}{2}
\begin{center}

Accepted by the Graduate Faculty, Indiana University, in partial fulfillment of the
requirements for the degree of Doctor of Philosophy.
\end{center}
\vspace{1cm}
\
Doctoral Committee
\vspace{1cm}

\begin{flushright}
\underline{\hspace{5cm}} \\
Robert L. Goldstone, Ph.D.
\vspace{2cm}

\underline{\hspace{5cm}} \\
Robert M. Nosofsky, Ph.D.
\vspace{2cm}

\underline{\hspace{5cm}} \\
Peter M. Todd, Ph.D.
\vspace{2cm}

\underline{\hspace{5cm}} \\
Michael N. Jones, Ph.D.
\end{flushright}

\
May 28th, 2024

\newpage

\begin{centering}

\vspace*{6.5cm}

@2024 \\
\vspace{1cm}

Thomas E. Gorman
\vspace{.2cm}

\vspace{5cm}

\end{centering}

\newpage
\begin{center}
\textbf{Acknowledgements}
\end{center}
\newpage