### 2024-10-12 Manuscript/pre_sections/abstract.qmd

\begin{centering}

{Thomas E. Gorman}

{THE ROLE OF VARIABILITY IN LEARNING GENERALIZATION: A COMPUTATIONAL MODELING APPROACH}

\end{centering}

Researchers and educators have long been interested in how the variability of training examples experienced
during learning influences subsequent generalization. However, previous studies have reported both beneficial
and detrimental effects, leaving the nature of this relationship unclear. Furthermore, a formal computational
modeling approach to account for the effects of training variability has been largely lacking within the domain of
visuomotor skill learning. The current work presents two projects which shed light on this issue, each consisting of
several experiments which compare varied training (i.e., multiple training items) in comparison to constant
training (training with a single item).
The empirical results, much like those of the broader literature, provide evidence both for and against a benefit of
variable training in visuomotor skill learning, highlighting the complex interplay between training conditions and
generalization performance. To provide a theoretical account of these findings, this dissertation employs both
instance-based and connectionist computational modeling approaches. The instance-based modeling approach
introduced in project 1 provides a theoretically justifiable method of quantifying/controlling for similarity
between training and testing conditions, while also demonstrating that varied training may induce broader
generalization in the similarity function relating training and test items. In project 2, the Extrapolation-
Association Model (EXAM) provided the best account of the testing data across all experiments, capturing the
constant groups' ability to extrapolate linearly to novel regions despite limited training experience, while also
revealing potential detriments of varied training for simple extrapolation tasks.
By integrating empirical findings with computational modeling approaches, this dissertation sheds light on the
complex relationship between training variability and generalization performance.

\newpage