### 2024-10-12 index.qmd
initial re-render with inclusion of acknowledgement and abstract

---
title: "Dissertation"
author:
- name: Thomas E. Gorman
  affiliations: 
  - name:  Department of Psychological and Brain Sciences, Indiana University-Bloomington, USA
    affiliation-url: https://psych.indiana.edu/
  - name:  Program in Cognitive Science, Indiana University-Bloomington, USA
    affiliation-url: https://cogs.indiana.edu/index.html
  url: https://tegorman13.github.io/
comments: false
date: today
#date-format: 'MMM YYYY'
format:
  html:
    grid:
      sidebar-width: 250px
      body-width: 1400px
      margin-width: 170px
      gutter-width: 1.5rem
    self-contained: false
---

## Project 1

[{{< fa file-lines >}} Project 1 page](https://tegorman13.github.io/Dissertation/Sections/IGAS.html)

[{{< fa file-pdf >}} pdf of the journal article](https://tegorman13.github.io/pdf/Gorman_Goldstone_2022_Instance-based_model_varied_practice.pdf){target="_blank"}\

[{{< fa scroll >}} Link to online version of journal article](https://www.sciencedirect.com/science/article/abs/pii/S0010028522000299){target="_blank"}

[{{< fa brands github >}} repo](https://github.com/tegorman13/IGAS_Variability) \

## Project 2

[{{< fa file-lines >}} Link to project 2 page](https://tegorman13.github.io/htw/){target="_blank"} \

[{{< fa file-pdf >}} Working Draft of Manuscript](https://tegorman13.github.io/htw/paper.html){target="_blank"}

[{{< fa brands github >}} repo](https://github.com/tegorman13/htw) \

## Full Manuscript

[{{< fa file-lines >}} Web page version](https://tegorman13.github.io/Dissertation/Manuscript/output/manuscript.html)

[{{< fa file-lines >}} Version with code](https://tegorman13.github.io/Dissertation/Sections/full.html)

[{{< fa file-pdf >}} pdf version](https://tegorman13.github.io/Dissertation/paper.html){target="_blank"}\

[{{< fa brands github >}} repo](https://github.com/tegorman13/Dissertation) \

## Slides

[{{< fa file-lines >}} Slides overview](https://tegorman13.github.io/Dissertation/Presentation/slides.html)

[{{< fa person-chalkboard size=1.0xs >}} Defense slides](https://tegorman13.github.io/Dissertation/Presentation/slides_revealjs.html#/)

##

-----

[Revisions](revisions.qmd)