### 2024-10-12 Manuscript/manuscript_full.qmd

---
code-repo: "Access the code, data, and analysis at <https://github.com/tegorman13/Dissertation>"
toc: false
toc-depth: 4
toc-location: body
#cache: true
output-dir: output
execute: 
  warning: false
  include: true
  echo: false
format:
  docx:
    prefer-html: true
    output-file: "TG_Full_Dissertation2.docx"
    toc: true
  html:
    toc: false
    language: 
      title-block-published: "Latest version"
    shift-heading-level-by: 0
    toc-depth: 4
    link-citations: true
    page-layout: full
    format-links: [pdf, gfm, docx]
  pdf:
    cite-method: citeproc
    #date: today
    output-file: "TG_Full_Dissertation2.pdf"
    mathfont: "Libertinus Math"
    documentclass: article
    papersize: letter
    fontsize: 11pt
    linestretch: 1.5
          # Document styling
    geometry:
      - top=1in     
      - bottom=1in  
      - left=1in    
      - right=1in  
      - heightrounded

      # Citations and links
    # lot: true
    # lof: true
    colorlinks: true
    linkcolor: Black
    urlcolor: Black
    citecolor: Black
    link-citations: true
    # linestretch: 1.05
    include-in-header:
      # The \everydisplay thing makes all math blocks use \small font
      text: |
        \usepackage{mathtools}
        \everydisplay\expandafter{\the\everydisplay\small}
        \usepackage{float}
        \floatplacement{figure}{H}
        \floatplacement{table}{H}
        \usepackage{setspace}
        \pagenumbering{arabic}
        \usepackage{titlesec}
        \titleformat{\section}{\fontsize{12}{14}\bfseries}{\thesection}{1em}{}
        \titleformat{\subsection}{\fontsize{11}{14}\bfseries}{\thesubsection}{1em}{}
---

::: {.content-visible when-format="html"}

{{< include pre_sections/pre_html2.qmd >}}

:::

::: {.content-visible when-format="pdf"}

{{< include pre_sections/pre_pdf_full.qmd >}}
{{< include pre_sections/acknowledgements.qmd >}}
{{< include pre_sections/abstract.qmd >}}

:::

::: {.content-visible when-format="docx"}
{{< include pre_sections/pre_docx.qmd >}}
{{< include pre_sections/acknowledgements.qmd >}}
{{< include pre_sections/abstract.qmd >}}
:::

{{< pagebreak >}}

::: {.content-visible when-format="html"}

{{< include pre_sections/toc_html.qmd >}}

:::

::: {.content-visible when-format="pdf"}

{{< include pre_sections/toc_pdf.qmd >}}

:::

{{< pagebreak >}}

# Introduction

## Varied Training and Generalization

Varied training has been shown to influence learning in a wide array of different tasks and domains, including categorization [@hahnEffectsCategoryDiversity2005; @maddoxStimulusRangeDiscontinuity2011; @posnerGenesisAbstractIdeas1968; @nosofskyModelguidedSearchOptimal2019; @morgensternOneshotCategorizationNovel2019; @plebanekEffectsFrequencyVariability2021], language learning [@jonesDensityDistinctivenessEarly2020; @perryLearnLocallyThink2010; @twomeyAllRightNoises2018; @wonnacottInputEffectsAcquisition2012; @brekelmansDoesHighVariability2022], anagram completion [@goodeSuperiorityVariableRepeated2008], perceptual learning [@manentiVariabilityTrainingUnlocks2023; @lovibondStimulusDiscriminabilityInduction2020; @robsonSpecificVariedPractice2022a; @zamanPerceptualVariabilityImplications2021], trajectory extrapolation [@fulvioTaskSpecificResponseStrategy2014], cognitive control tasks [@sabahWhenLessMore2019; @moshon-cohenStimulusVariabilityImproves2024], associative learning [@leeEvidentialDiversityIncreases2019; @reichmannVariabilityAbstractionEvaluative2023; @fanStimulusDiversityIncreases2022; @pradaExperiencedCategoryVariability2020; @liveseyRevisitingPeakShift2019], visual search [@georgeStimulusVariabilityTask2021; @gonzalezDiversityTrainingEnhances2011; @kelleyLearningAttendEffects2009], voice identity learning [@lavanEffectsHighVariability2019], face recognition [@honigPerceptualSimilarityModulates2022; @burtonIdentityVariationRepresentations2016; @menonVariationPhotosSame2015],  the perception of social group heterogeneity [@linvilleExemplarAbstractionModels1993; @konovalovaInformationSamplingExplanation2020; @parkPerceptionVariabilityCategory1987; @gershmanStructureLearningPrinciples2023] ,  simple motor learning [@braunMotorTaskVariation2009; @kerrSpecificVariedPractice1978; @rollerVariablePracticeLenses2001; @willeyLimitedGeneralizationVaried2018], sports training  [@greenPracticeVariabilityTransfer1995a; @northEffectConsistentVaried2019; @breslinConstantVariablePractice2012], and  complex skill learning [@seowTransferEffectsVaried2019; @huetEducationAttentionExplanation2011; @hacquesVisualControlClimbing2022]. See @czyzVariabilityPracticeInformation2021 or @ravivHowVariabilityShapes2022 for more detailed reviews.

{{< pagebreak >}}

# References
::: {#refs}
:::

{{< pagebreak >}}

::: {.content-visible when-format="pdf"}

{{< include pre_sections/diss_cv.qmd >}}

:::